### 2015-02-20 test.tex
!

\documentclass{scrartcl}
\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath,amssymb,amstext}
\usepackage[automark]{scrpage2}
\usepackage[bottom]{footmisc}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage[a4paper,vmargin={30mm,40mm},hmargin={28mm,28mm}]{geometry}
\usepackage[hidelinks]{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
\title{PDBF Testdokument}
\author{
 Patrick Bender\\
  \texttt{s9pabend@stud.uni-saarland.de}
  \and
 Christian Goldstein\\
  \texttt{c.goldstein@goldstein-systemberatung.de}
}
\date{\today{}, Saarbrücken}
\pagestyle{plain}
\clearscrheadfoot
\cfoot[]{\pagemark}
\lstset{
basicstyle=\ttfamily,
commentstyle=\itshape,
keywordstyle=\bfseries,
    frame=single,
    breaklines=true,
    postbreak=,
keepspaces=true
}
\sloppy

\makeatletter
\def\@maketitle{%
  \vspace*{-\topskip}      % remove the initial space
  \begingroup\centering    % instead of \begin{center}
  \let \footnote \thanks
  \hrule height \z@        % to avoid the insertion of lineskip glue
    {\LARGE \@title \par}%
    \vskip 1.5em 
    {\large
      \lineskip .5em 
      \begin{tabular}[t]{c}%
        \@author
      \end{tabular}\par}%
    \vskip 1em 
    {\large \@date}%
  \par\endgroup            % instead of \end{center}
  \vskip 1.5em             % <--- modify this to adjust the separation
}
\makeatother

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{pred}{rgb}{0.9,0,0}
\definecolor{pgrey}{rgb}{0.46,0.45,0.48}
\lstset{language=Java,
  showspaces=false,
  showtabs=false,
  breaklines=true,
  showstringspaces=false,
  breakatwhitespace=true,
  commentstyle=\color{pgreen},
  keywordstyle=\color{pblue},
  stringstyle=\color{pred},
  basicstyle=\ttfamily\scriptsize,
  moredelim=[il][\textcolor{pgrey}]{$$},
  moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%}
}

\begin{document}
\maketitle
\begin{wrapfigure}{r}{0.47\textwidth}
  \begin{center}
    \includegraphics[width=0.4\textwidth]{Tabelle1.png}
  \end{center}
  \caption{TestGraph}
\end{wrapfigure} ~\\
Test Queries:\\
\color{blue}
SELECT * FROM test;\\
SELECT col2 FROM test\\
WHERE col1=1;\\
INSERT INTO test\\
VALUES (1, 123);\color{black}\\[8pt]
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\[8pt]
Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.   \\[8pt]
Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.   \\[8pt]
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat. Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.

\end{document}

### 2015-02-24 test.tex
latex compilation is now in alpha stage - generating lineCharts is possible - position calculation works TODO: - options for lineCharts - texts

@@
 \usepackage[T1]{fontenc}
-\usepackage{graphicx}
 \usepackage{amsfonts}
@@
 \usepackage{xcolor}
+\usepackage{PDBF}
 \usepackage{tabularx}
@@
 \begin{wrapfigure}{r}{0.47\textwidth}
-  \begin{center}
-    \includegraphics[width=0.4\textwidth]{Tabelle1.png}
-  \end{center}
-  \caption{TestGraph}
+     \centering
+     \lineChart
+     \caption{Graph1}
 \end{wrapfigure} ~\\
@@
 Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\[8pt]
+
+\begin{wrapfigure}{r}{0.47\textwidth}
+     \centering
+     \lineChart
+     \caption{Graph2}
+\end{wrapfigure} ~\\
 Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
@@
 Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
-
+\\[8pt]
+\begin{figure}[h]
+    \centering
+    \lineChart
+    \caption{Graph3}
+\end{figure}
 \end{document}

### 2015-02-24 src/pdbf/common/UnitTypeAdapter.java

@@ -1,45 +1,0 @@
-package pdbf.common;
-
-import java.lang.reflect.Type;
-
-import com.google.gson.JsonDeserializationContext;
-import com.google.gson.JsonDeserializer;
-import com.google.gson.JsonElement;
-import com.google.gson.JsonObject;
-import com.google.gson.JsonParseException;
-import com.google.gson.JsonPrimitive;
-import com.google.gson.JsonSerializationContext;
-import com.google.gson.JsonSerializer;
-
-public class UnitTypeAdapter implements JsonSerializer<Unit>,
-	JsonDeserializer<Unit> {
-    private static final String CLASSNAME = "C";
-    private static final String INSTANCE = "I";
-
-    @Override
-    public JsonElement serialize(Unit src, Type typeOfSrc,
-	    JsonSerializationContext context) {
-	JsonObject retValue = new JsonObject();
-	String className = src.getClass().getCanonicalName();
-	retValue.addProperty(CLASSNAME, className);
-	JsonElement elem = context.serialize(src);
-	retValue.add(INSTANCE, elem);
-	return retValue;
-    }
-
-    @Override
-    public Unit deserialize(JsonElement json, Type typeOfT,
-	    JsonDeserializationContext context) throws JsonParseException {
-	JsonObject jsonObject = json.getAsJsonObject();
-	JsonPrimitive prim = (JsonPrimitive) jsonObject.get(CLASSNAME);
-	String className = prim.getAsString();
-	Class<?> klass = null;
-	try {
-	    klass = Class.forName(className);
-	} catch (ClassNotFoundException e) {
-	    e.printStackTrace();
-	    throw new JsonParseException(e.getMessage());
-	}
-	return context.deserialize(jsonObject.get(INSTANCE), klass);
-    }
-}

### 2015-02-24 src/pdbf/html/CreateTestLargeDB.java

@@ -4,12 +4,9 @@ import java.io.FileWriter;
 import java.util.ArrayList;
 import java.util.Random;
 
-import pdbf.common.CustomUnit;
 import pdbf.common.LineChart;
 import pdbf.common.Overlay;
 import pdbf.common.Text;
-import pdbf.common.Unit;
-import pdbf.common.UnitTypeAdapter;
 import pdbf.common.Visualization;
 import pdbf.common.VisualizationTypeAdapter;
 
@@ -23,18 +20,13 @@ public class CreateTestLargeDB {
 	builder.setPrettyPrinting();
 	builder.registerTypeAdapter(Visualization.class,
 		new VisualizationTypeAdapter());
-	builder.registerTypeAdapter(Unit.class, new UnitTypeAdapter());
 	Gson gson = builder.create();
 
 	ArrayList<Overlay> overlays = new ArrayList<Overlay>();
 
-	//Date date = new Date();
-	//CustomUnit temp = new CustomUnit("Temperatur");
-	CustomUnit z1 = new CustomUnit("Zahl1");
-	CustomUnit z2 = new CustomUnit("Zahl2");
 	LineChart chart = new LineChart("SELECT * FROM data2 WHERE Zahl2 > 0;",
 		0.5150353982300885, 0.8407079646017699, 0.2496875, 0.3601875,
-		false, z1, z2, 1);
+		1, false, "Zahl1", "Zahl2");
 	Overlay cur = new Overlay("Tabelle1", chart);
 	overlays.add(cur);
 

### 2015-02-24 src/pdbf/latex/CompleteRun.java

@@ -1,0 +1,15 @@
+package pdbf.latex;
+
+import pdbf.html.HTML_Compiler;
+
+public class CompleteRun {
+
+    public static void main(String[] args) {
+	System.out.println("Compile LaTeX...");
+	LaTeX_Compiler.main(args);
+	System.out.println("Compile HTML...");
+	HTML_Compiler.main(args);
+	System.out.println("Finished!");
+    }
+    
+}

### 2015-02-24 test.tex
automatic image generation works TODO: fix scale of image

@@
      \centering
-     \lineChart
+     \lineChart{width=0.4\textwidth, height=0.3\textwidth}
      \caption{Graph1}
@@
 At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\[8pt]
+Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
 
-\begin{wrapfigure}{r}{0.47\textwidth}
+\begin{wrapfigure}{r}{0.37\textwidth}
      \centering
-     \lineChart
+     \lineChart{width=0.3\textwidth, height=0.4\textwidth}
      \caption{Graph2}
-\end{wrapfigure} ~\\
+\end{wrapfigure}
 Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
@@
 Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
+Lorem ipsum dolor sit amet, consectetuer adipiscing \sql{Testquery test test test} elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
 Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
 \\[8pt]
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \lineChart
+    \lineChart{width=0.4\textwidth, height=0.4\textwidth}
     \caption{Graph3}

### 2015-02-24 src/pdbf/java/Java_Compiler.java

@@ -1,11 +1,13 @@
 package pdbf.java;
 
 import java.io.File;
-import java.io.FileReader;
+import org.apache.commons.io.FileUtils;
 
 import pdbf.common.Overlay;
+import pdbf.common.Tools;
 import pdbf.common.Visualization;
 import pdbf.common.VisualizationTypeAdapter;
+
 import com.google.gson.Gson;
 import com.google.gson.GsonBuilder;
 
@@ -36,9 +38,8 @@ public class Java_Compiler {
 	Gson gson = builder.create();
 	Overlay[] overlays = null;
 	try {
-	    FileReader reader = new FileReader("config.json");
-	    overlays = gson.fromJson(reader, Overlay[].class);
-	    reader.close();
+	    String json = FileUtils.readFileToString(new File("config.json"), Tools.utf8);
+	    overlays = gson.fromJson(json, Overlay[].class);
 	} catch (Exception e) {
 	    e.printStackTrace();
 	}

### 2015-02-25 test.tex
implemented sql commands and improved inline positioning fixed coordinates for sql and chart

@@
 \usepackage{tabularx}
+\usepackage{wrapfig}
 \usepackage[a4paper,vmargin={30mm,40mm},hmargin={28mm,28mm}]{geometry}
@@
 \begin{wrapfigure}{r}{0.47\textwidth}
-     \centering
-     \lineChart{width=0.4\textwidth, height=0.3\textwidth}
+{ \centering %
+\lineChart{0.4\textwidth}{0.3\textwidth} %
+\\ }
+\vspace{-5pt}
 \caption{Graph1}
+\vspace{-20pt}
 \end{wrapfigure} ~\\
@@
 At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
+Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\
 
-\begin{wrapfigure}{r}{0.37\textwidth}
-     \centering
-     \lineChart{width=0.3\textwidth, height=0.4\textwidth}
+\noindent \begin{wrapfigure}{r!}{0.37\textwidth}
+     \vspace{-12pt}
+     { \centering %
+     \lineChart{0.3\textwidth}{0.4\textwidth} %
+     \\ }
+     \vspace{-5pt}
      \caption{Graph2}
+     \vspace{-20pt}
 \end{wrapfigure}
@@
 Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
-\\[8pt]
+Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait facilisi.
+\newpage
+\noindent %
 Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing \sql{Testquery test test test} elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
 Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
 \\[8pt]
-\begin{figure}[h!]
+A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A \sql{Test multiline!} A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A
+\begin{figure}[h]
     \centering
-    \lineChart{width=0.4\textwidth, height=0.4\textwidth}
+    \lineChart{0.4\textwidth}{0.4\textwidth} \\
     \caption{Graph3}
 \end{figure}
+
 \end{document}

### 2015-02-25 capture.js

@@ -1,0 +1,29 @@
+var system = require('system');
+if (system.args.length !== 2) {
+console.log('Usage: ' + system.args[0] + ' webpagefile');
+phantom.exit();
+} 
+
+var page = require('webpage').create();
+page.open(system.args[1], function() {
+	var w = page.evaluate(function() {
+		return outw;
+	});
+	var h = page.evaluate(function() {
+		return outh;
+	});
+	page.zoomFactor = 1.75;
+	page.viewportSize = {
+		width: w,
+		height: h
+	};
+	window.setTimeout(function () {
+		var name = page.evaluate(function() {
+			return json.name;
+		});
+        page.render(name +'.png');
+        phantom.exit();
+    }, 500);
+});
+
+

### 2015-02-25 test.tex
another gitignore update
hopefully last gitignore fix
tracked bug with page numbers and fixed it

@@
 Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-\\[8pt]
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.\\[8pt]
 Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
@@
 \\[8pt]
-A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A \sql{Test multiline!} A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A
+A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A \sql{Test multiline!} A A A A A A A A A A A A A A A A A A A A A A \sql{Testquery 2} A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A \sql{Testquery 3} A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A
 \begin{figure}[h]

### 2015-02-25 test.tex
fixed bug with sql command

--- 2015-02-25 test.tex
@@
 \end{wrapfigure} ~\\
-Test Queries:\\
-\color{blue}
-SELECT * FROM test;\\
-SELECT col2 FROM test\\
-WHERE col1=1;\\
-INSERT INTO test\\
-VALUES (1, 123);\color{black}\\[8pt]
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
+Test Queries:\\[3pt]
+\sql{SELECT * FROM test;}{SELECT * FROM test;}\\[3pt]
+\sql{SELECT col2 FROM test WHERE col1=1;}{SELECT col2 FROM test WHERE col1=1;}\\[3pt]
+\sql{INSERT INTO test}{INSERT INTO test VALUES (1, 123);}\\
+\sql{VALUES (1, 123);}{INSERT INTO test VALUES (1, 123);}\\[8pt]
+Lorem ipsum \sql{SELECT * FROM test;}{SELECT * FROM test;} dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
 At vero eos et accusam et justo duo dolores et ea rebum.
@@
 \\[8pt]
-A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A \sql{Test multiline!} A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A A
 \begin{figure}[h]

### 2015-02-25 test.tex
- removed some unused stuff from test.tex - adjusted scale of chart images to html size - chart legend position fixed

@@
 \documentclass{scrartcl}
-\usepackage{ucs}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{amsfonts}
-\usepackage{amsmath,amssymb,amstext}
-\usepackage[automark]{scrpage2}
-\usepackage[bottom]{footmisc}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{wrapfig}
-\usepackage{listings}
-\usepackage{setspace}
-\usepackage{xcolor}
 \usepackage{PDBF}
-\usepackage{tabularx}
 \usepackage{wrapfig}
 \usepackage[a4paper,vmargin={30mm,40mm},hmargin={28mm,28mm}]{geometry}
-\usepackage[hidelinks]{hyperref}
-\hypersetup{
-    colorlinks,
-    linkcolor={red!50!black},
-    citecolor={blue!50!black},
-    urlcolor={blue!80!black}
-}
+
 \title{PDBF Testdokument}
@@
 \date{\today{}, Saarbrücken}
-\pagestyle{plain}
-\clearscrheadfoot
-\cfoot[]{\pagemark}
-\lstset{
-basicstyle=\ttfamily,
-commentstyle=\itshape,
-keywordstyle=\bfseries,
-    frame=single,
-    breaklines=true,
-    postbreak=,
-keepspaces=true
-}
-\sloppy
 
@@
 \def\@maketitle{%
-  \vspace*{-\topskip}      % remove the initial space
-  \begingroup\centering    % instead of \begin{center}
+  \vspace*{-\topskip}%
+  \begingroup\centering%
   \let \footnote \thanks
-  \hrule height \z@        % to avoid the insertion of lineskip glue
+  \hrule height \z@%
     {\LARGE \@title \par}%
@@
     {\large \@date}%
-  \par\endgroup            % instead of \end{center}
-  \vskip 1.5em             % <--- modify this to adjust the separation
+  \par\endgroup%
+  \vskip 1.5em%
 }
 \makeatother
-
-\definecolor{pblue}{rgb}{0.13,0.13,1}
-\definecolor{pgreen}{rgb}{0,0.5,0}
-\definecolor{pred}{rgb}{0.9,0,0}
-\definecolor{pgrey}{rgb}{0.46,0.45,0.48}
-\lstset{language=Java,
-  showspaces=false,
-  showtabs=false,
-  breaklines=true,
-  showstringspaces=false,
-  breakatwhitespace=true,
-  commentstyle=\color{pgreen},
-  keywordstyle=\color{pblue},
-  stringstyle=\color{pred},
-  basicstyle=\ttfamily\scriptsize,
-  moredelim=[il][\textcolor{pgrey}]{$$},
-  moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%}
-}
 

### 2015-02-26 test.tex
- added test data - fixed lots of bugs - added parameters to latex commands

@@
 { \centering %
-\lineChart{0.4\textwidth}{0.3\textwidth} %
+\lineChart{0.4\textwidth}{0.3\textwidth}{SELECT * FROM data1;}{Money (in €)}{Date}{false}{title: "Title"} %
 \\ }
@@
      { \centering %
-     \lineChart{0.3\textwidth}{0.4\textwidth} %
+     \lineChart{0.3\textwidth}{0.4\textwidth}{SELECT * FROM data1 WHERE date >= '2008/05/08';}{Money (in €)}{Date}{true}{}%
      \\ }
@@
     \centering
-    \lineChart{0.4\textwidth}{0.4\textwidth} \\
+    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}{Runtime (in seconds)}{Date}{false}{} \\
     \caption{Graph3}

### 2015-02-26 src/pdbf/common/BarChart.java

@@ -2,15 +2,11 @@ package pdbf.common;
 
 public class BarChart extends Chart {
 
-    public BarChart(String query, double x1, double x2, double y1, double y2,
-	    long page, boolean logScale, String xUnitName, String yUnitName,
-	    long yUnitFrom, long yUnitTo, String options) {
-	super(query, x1, x2, y1, y2, page, logScale, xUnitName, yUnitName,
-		yUnitFrom, yUnitTo, options);
+    public BarChart(String query, double x1, double x2, double y1, double y2, long page, boolean logScale, String xUnitName, String yUnitName, String options) {
+	super(query, x1, x2, y1, y2, page, logScale, xUnitName, yUnitName, options);
     }
 
-    public BarChart(String query, double x1, double x2, double y1, double y2,
-	    long page, boolean logScale, String xUnitName, String yUnitName) {
+    public BarChart(String query, double x1, double x2, double y1, double y2, long page, boolean logScale, String xUnitName, String yUnitName) {
 	super(query, x1, x2, y1, y2, page, logScale, xUnitName, yUnitName);
     }
 

### 2015-02-26 src/pdbf/common/Tools.java

@@ -10,7 +10,7 @@ import org.apache.commons.io.FileUtils;
 public class Tools {
 
     public static Charset utf8 = Charset.forName("UTF-8");
-    
+
     public static String encodeFileToBase64Binary(File file) {
 	String encodedfile = null;
 	try {
@@ -21,7 +21,7 @@ public class Tools {
 	}
 	return encodedfile;
     }
-    
+
     public static String encodeStringToBase64Binary(String string) throws UnsupportedEncodingException {
 	return Base64.encodeBase64String(string.getBytes(utf8));
     }

### 2015-02-26 src/pdbf/common/VisualizationTypeAdapter.java

@@ -11,14 +11,12 @@ import com.google.gson.JsonPrimitive;
 import com.google.gson.JsonSerializationContext;
 import com.google.gson.JsonSerializer;
 
-public class VisualizationTypeAdapter implements JsonSerializer<Visualization>,
-	JsonDeserializer<Visualization> {
+public class VisualizationTypeAdapter implements JsonSerializer<Visualization>, JsonDeserializer<Visualization> {
     private static final String CLASSNAME = "C";
     private static final String INSTANCE = "I";
 
     @Override
-    public JsonElement serialize(Visualization src, Type typeOfSrc,
-	    JsonSerializationContext context) {
+    public JsonElement serialize(Visualization src, Type typeOfSrc, JsonSerializationContext context) {
 	JsonObject retValue = new JsonObject();
 	String className = src.getClass().getCanonicalName();
 	retValue.addProperty(CLASSNAME, className);
@@ -28,8 +26,7 @@ public class VisualizationTypeAdapter implements JsonSerializer<Visualization>,
     }
 
     @Override
-    public Visualization deserialize(JsonElement json, Type typeOfT,
-	    JsonDeserializationContext context) throws JsonParseException {
+    public Visualization deserialize(JsonElement json, Type typeOfT, JsonDeserializationContext context) throws JsonParseException {
 	JsonObject jsonObject = json.getAsJsonObject();
 	JsonPrimitive prim = (JsonPrimitive) jsonObject.get(CLASSNAME);
 	String className = prim.getAsString();

### 2015-02-26 src/pdbf/html/CreateTest.java

@@ -1,68 +1,0 @@
-package pdbf.html;
-
-import java.io.FileWriter;
-import java.util.ArrayList;
-import pdbf.common.LineChart;
-import pdbf.common.Overlay;
-import pdbf.common.Text;
-import pdbf.common.Visualization;
-import pdbf.common.VisualizationTypeAdapter;
-
-import com.google.gson.Gson;
-import com.google.gson.GsonBuilder;
-
-public class CreateTest {
-
-    public static void main(String[] args) {
-	GsonBuilder builder = new GsonBuilder();
-	builder.setPrettyPrinting();
-	builder.registerTypeAdapter(Visualization.class,
-		new VisualizationTypeAdapter());
-	Gson gson = builder.create();
-
-	ArrayList<Overlay> overlays = new ArrayList<Overlay>();
-
-	LineChart chart = new LineChart("SELECT * FROM data1",
-		0.5150353982300885, 0.8407079646017699, 0.2496875, 0.3601875, 1, false, "Datum", "Temperatur"); 
-	Overlay cur = new Overlay("Tabelle1", chart);
-	overlays.add(cur);
-
-	Text text = new Text("SELECT * FROM test;",
-		0.12831858407079646017699115044248,
-		0.33185840707964601769911504424779, 0.2515625, 0.2671875, 1);
-	cur = new Overlay("SQL1", text);
-	overlays.add(cur);
-
-	text = new Text("SELECT col2 FROM test WHERE col1=1;",
-		0.12831858407079646017699115044248,
-		0.34955752212389380530973451327434, 0.2671875, 0.2984375, 1);
-	cur = new Overlay("SQL2", text);
-	overlays.add(cur);
-
-	text = new Text("INSERT INTO test VALUES (1, 123);",
-		0.12831858407079646017699115044248,
-		0.30088495575221238938053097345133, 0.2984375, 0.334375, 1);
-	cur = new Overlay("SQL3", text);
-	overlays.add(cur);
-
-	try {
-	    FileWriter writer = new FileWriter("config.json");
-	    gson.toJson(overlays, writer);
-	    writer.close();
-	} catch (Exception e) {
-	    e.printStackTrace();
-	}
-	
-	try {
-	    FileWriter writer = new FileWriter("db.sql");
-	    //Random r = new Random(System.currentTimeMillis());
[…]

### 2015-02-26 src/pdbf/html/CreateTestLargeDB.java

@@ -1,71 +1,0 @@
-package pdbf.html;
-
-import java.io.FileWriter;
-import java.util.ArrayList;
-import java.util.Random;
-
-import pdbf.common.LineChart;
-import pdbf.common.Overlay;
-import pdbf.common.Text;
-import pdbf.common.Visualization;
-import pdbf.common.VisualizationTypeAdapter;
-
-import com.google.gson.Gson;
-import com.google.gson.GsonBuilder;
-
-public class CreateTestLargeDB {
-
-    public static void main(String[] args) {
-	GsonBuilder builder = new GsonBuilder();
-	builder.setPrettyPrinting();
-	builder.registerTypeAdapter(Visualization.class,
-		new VisualizationTypeAdapter());
-	Gson gson = builder.create();
-
-	ArrayList<Overlay> overlays = new ArrayList<Overlay>();
-
-	LineChart chart = new LineChart("SELECT * FROM data2 WHERE Zahl2 > 0;",
-		0.5150353982300885, 0.8407079646017699, 0.2496875, 0.3601875,
-		1, false, "Zahl1", "Zahl2");
-	Overlay cur = new Overlay("Tabelle1", chart);
-	overlays.add(cur);
-
-	Text text = new Text("SELECT * FROM test;",
-		0.12831858407079646017699115044248,
-		0.33185840707964601769911504424779, 0.2515625, 0.2671875, 1);
-	cur = new Overlay("SQL1", text);
-	overlays.add(cur);
-
-	text = new Text("SELECT col2 FROM test WHERE col1=1;",
-		0.12831858407079646017699115044248,
-		0.34955752212389380530973451327434, 0.2671875, 0.2984375, 1);
-	cur = new Overlay("SQL2", text);
-	overlays.add(cur);
-
-	text = new Text("INSERT INTO test VALUES (1, 123);",
-		0.12831858407079646017699115044248,
-		0.30088495575221238938053097345133, 0.2984375, 0.334375, 1);
-	cur = new Overlay("SQL3", text);
-	overlays.add(cur);
-
-	try {
-	    FileWriter writer = new FileWriter("config.json");
-	    gson.toJson(overlays, writer);
-	    writer.close();
-	} catch (Exception e) {
-	    e.printStackTrace();
-	}
-	
-	try {
-	    FileWriter writer = new FileWriter("db.sql");
[…]

### 2015-03-02 test.tex
cleanup main directory

@@
 
+\DBSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
+\DBSQLFile{somedb.sql}
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{dim_compiler}
+
 \begin{document}
@@
 { \centering %
-\lineChart{0.4\textwidth}{0.3\textwidth}{SELECT * FROM data1;}{Money (in €)}{Date}{false}{title: "Title"} %
+\lineChart{0.4\textwidth}{0.3\textwidth}{SELECT * FROM data1;}{Money (in €)}{Date}{false}{"title": "Title", "titleHeight": 25} %
 \\ }
@@
      { \centering %
-     \lineChart{0.3\textwidth}{0.4\textwidth}{SELECT * FROM data1 WHERE date >= '2008/05/08';}{Money (in €)}{Date}{true}{}%
+     \lineChart{0.3\textwidth}{0.4\textwidth}{SELECT * FROM data1 WHERE date >= '2008/05/11';}{Money (in €)}{Date}{true}{}%
      \\ }

### 2015-03-02 src/pdbf/common/Database.java

@@ -1,0 +1,22 @@
+package pdbf.common;
+
+public class Database extends Visualization {
+    public int type;
+    public String value1;
+    public String value2;
+    public String value3;
+    public String value4;
+
+    public Database(int type, String value1, String value2, String value3, String value4) {
+	super();
+	this.type = type;
+	this.value1 = value1;
+	this.value2 = value2;
+	this.value3 = value3;
+	this.value4 = value4;
+    }
+    
+    public Database() {
+    }
+    
+}

### 2015-03-02 src/pdbf/html/HTML_Compiler.java

@@ -11,8 +11,8 @@ public class HTML_Compiler {
 	System.out.println("Compiling HTML...");
 	try {
 	    String viewer;
-	    String viewerHEAD = FileUtils.readFileToString(new File("templateHEADsqlite.html"), Tools.utf8);
-	    String viewerTAIL = FileUtils.readFileToString(new File("templateTAILsqlite.html"), Tools.utf8);
+	    String viewerHEAD = FileUtils.readFileToString(new File("out/web/templateHEADsqlite.html"), Tools.utf8);
+	    String viewerTAIL = FileUtils.readFileToString(new File("out/web/templateTAILsqlite.html"), Tools.utf8);
 	    viewer = viewerHEAD + 
 		    "pdf_base64 = \"" + Tools.encodeFileToBase64Binary(new File("test.pdf")) + "\";\r\n" + 
 		    "db_base64 = \"" + Tools.encodeFileToBase64Binary(new File("db.sql")) + "\";\r\n" + 
@@ -25,8 +25,8 @@ public class HTML_Compiler {
 	
 	try {
 	    String viewer;
-	    String viewerHEAD = FileUtils.readFileToString(new File("templateHEADalasql.html"), Tools.utf8);
-	    String viewerTAIL = FileUtils.readFileToString(new File("templateTAILalasql.html"), Tools.utf8);
+	    String viewerHEAD = FileUtils.readFileToString(new File("out/web/templateHEADalasql.html"), Tools.utf8);
+	    String viewerTAIL = FileUtils.readFileToString(new File("out/web/templateTAILalasql.html"), Tools.utf8);
 	    viewer = viewerHEAD + 
 		    "pdf_base64 = \"" + Tools.encodeFileToBase64Binary(new File("test.pdf")) + "\";\r\n" + 
 		    "db_base64 = \"" + Tools.encodeFileToBase64Binary(new File("db.sql")) + "\";\r\n" + 

### 2015-03-02 src/pdbf/java/Java_Compiler.java

@@ -1,6 +1,7 @@
 package pdbf.java;
 
 import java.io.File;
+
 import org.apache.commons.io.FileUtils;
 
 import pdbf.common.Overlay;

### 2015-03-02 test.tex
Added .project and .settings to gitignore
end of db.sql

@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{dim_compiler}
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time} %dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
 
@@
 \end{figure}
+\begin{figure}[h]
+    \centering
+    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT * FROM dim_time;}{Runtime}{Row ID}{false}{} \\
+    \caption{Graph4}
+\end{figure}
 

### 2015-03-03 test.tex
!

@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time} %dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
 

### 2015-03-03 src/pdbf/html/CompleteRun_HTML.java

@@ -9,10 +9,11 @@ public class CompleteRun_HTML {
     public static void main(String[] args) {
 	LaTeX_Compiler.main(args);
 	HTML_Compiler.main(args);
-	//new File("dim.json").delete();
-	//new File("config.json").delete();
-	//new File("db.sql").delete();
-	//new File("null.png").delete();
+	new File("dim.json").delete();
+	new File("config.json").delete();
+	new File("db.sql").delete();
+	new File("db.json").delete();
+	new File("null.png").delete();
 	System.out.println("Finished!");
     }
     

### 2015-03-03 src/pdbf/latex/Alasql.java

@@ -10,8 +10,8 @@ import com.google.gson.JsonSerializer;
 
 public class Alasql implements JsonSerializer<Alasql> {
     String databaseid = "alasql";
-    ArrayList<Table> tables = new ArrayList<Table>();
-    ArrayList<String> tableNames = new ArrayList<String>();
+    private ArrayList<Table> tables = new ArrayList<Table>();
+    private ArrayList<String> tableNames = new ArrayList<String>();
     Emptyclass views = new Emptyclass();
     Emptyclass indices = new Emptyclass();
     Emptyclass sqlCache = new Emptyclass();
@@ -21,16 +21,23 @@ public class Alasql implements JsonSerializer<Alasql> {
     @Override
     public JsonElement serialize(Alasql src, Type typeOfSrc, JsonSerializationContext context) {
 	JsonObject retValue = new JsonObject();
-	retValue.addProperty("databaseid", databaseid);
-	for (int i = 0; i < tables.size(); ++i) {
-	    retValue.add(tableNames.get(i), context.serialize(tables.get(i)));
+	retValue.addProperty("databaseid", src.databaseid);
+	JsonObject tables = new JsonObject();
+	for (int i = 0; i < src.tables.size(); ++i) {
+	    tables.add(src.tableNames.get(i), context.serialize(src.tables.get(i)));
 	}
-	retValue.add("views", context.serialize(views));
-	retValue.add("indices", context.serialize(indices));
-	retValue.add("sqlCache", context.serialize(sqlCache));
-	retValue.addProperty("sqlCacheSize", sqlCacheSize);
-	retValue.addProperty("dbversion", dbversion);
+	retValue.add("tables", tables);
+	retValue.add("views", context.serialize(src.views));
+	retValue.add("indices", context.serialize(src.indices));
+	retValue.add("sqlCache", context.serialize(src.sqlCache));
+	retValue.addProperty("sqlCacheSize", src.sqlCacheSize);
+	retValue.addProperty("dbversion", src.dbversion);
 	return retValue;
+    }
+    
+    public void addTable(Table table, String tableName) {
+	tables.add(table);
+	tableNames.add(tableName);
     }
 
 }

### 2015-03-03 src/pdbf/latex/Column.java

@@ -3,4 +3,11 @@ package pdbf.latex;
 public class Column {
     String columnid;
     String dbtypeid;
+    
+    public Column(String columnid, String dbtypeid) {
+	super();
+	this.columnid = columnid;
+	this.dbtypeid = dbtypeid;
+    }
+    
 }

### 2015-03-03 src/pdbf/latex/Database.java

@@ -1,5 +1,0 @@
-package pdbf.latex;
-
-public class Database {
-    Alasql alasql = new Alasql();
-}

### 2015-03-03 src/pdbf/latex/LaTeX_Compiler.java

@@ -72,7 +72,8 @@ public class LaTeX_Compiler {
 	// Read JSON
 	GsonBuilder builder = new GsonBuilder();
 	builder.registerTypeAdapter(Visualization.class, new VisualizationTypeAdapter());
-	builder.setPrettyPrinting();
+	builder.registerTypeAdapter(Alasql.class, new Alasql());
+	builder.registerTypeAdapter(Table.class, new Table());
 	gson = builder.create();
 	Overlay[] overlays = null;
 	try {
@@ -87,14 +88,22 @@ public class LaTeX_Compiler {
 	}
 	// Split
 	File f = new File("db.sql");
+	File f2 = new File("db.json");
 	if (f.exists()) {
 	    if (!f.delete()) {
 		System.out.println("db.sql could not be deleted! Exiting...");
 		System.exit(-1);
 	    }
 	}
+	if (f2.exists()) {
+	    if (!f2.delete()) {
+		System.out.println("db.json could not be deleted! Exiting...");
+		System.exit(-1);
+	    }
+	}
 	try {
 	    f.createNewFile();
+	    f2.createNewFile();
 	} catch (IOException e1) {
 	    e1.printStackTrace();
 	}
@@ -128,10 +137,10 @@ public class LaTeX_Compiler {
 	} catch (Exception e) {
 	    e.printStackTrace();
 	}
-	
+
 	// Remove database entries from config
 	ArrayList<Overlay> olist = new ArrayList<Overlay>(Arrays.asList(overlays));
-	for (int i = olist.size()-1; i >= 0; --i) {
+	for (int i = olist.size() - 1; i >= 0; --i) {
 	    if (olist.get(i).type instanceof Database) {
 		olist.remove(i);
 	    }
@@ -153,7 +162,8 @@ public class LaTeX_Compiler {
     private static void processDatabase(Overlay overlay) {
 	try {
 	    Database db = (Database) overlay.type;
-	    File f = new File("db.json");
+	    File f = new File("db.sql");
+	    File f2 = new File("db.json");
 	    switch (db.type) {
 	    case 1:
 		FileUtils.writeStringToFile(f, db.value1 + System.lineSeparator(), Tools.utf8, true);
@@ -163,52 +173,38 @@ public class LaTeX_Compiler {
 		FileUtils.writeStringToFile(f, in + System.lineSeparator(), Tools.utf8, true);
 		break;
 	    case 3:
-		try {
[…]

### 2015-03-03 src/pdbf/latex/Table.java

@@ -2,44 +2,48 @@ package pdbf.latex;
 
 import java.lang.reflect.Type;
 import java.util.ArrayList;
+import java.util.Collection;
 
 import com.google.gson.JsonElement;
 import com.google.gson.JsonObject;
 import com.google.gson.JsonSerializationContext;
 import com.google.gson.JsonSerializer;
+import com.google.gson.reflect.TypeToken;
 
 public class Table implements JsonSerializer<Table> {
-    ArrayList<Column> columns;
-
-    // "xcolumns":{ },
-    // "data":[ ],
-    // "inddefs":{ },
-    // "indices":{ },
-    // "uniqs":{ },
-    // "uniqdefs":{ },
-    // "defaultfns":""
+    public ArrayList<Column> columns = new ArrayList<Column>();
+    public ArrayList<Data> data = new ArrayList<Data>();
+    Emptyclass inddefs = new Emptyclass();
+    Emptyclass indices = new Emptyclass();
+    Emptyclass uniqs = new Emptyclass();
+    Emptyclass uniqdefs = new Emptyclass();
+    String defaultfns = "";
 
     @Override
     public JsonElement serialize(Table src, Type typeOfSrc, JsonSerializationContext context) {
-	Column[] columns = new Column[this.columns.size()];
-	columns = this.columns.toArray(columns);
 	JsonObject retValue = new JsonObject();
-	retValue.add("columns", context.serialize(columns));
-	JsonObject xolumns = new JsonObject();
-	for (Column c : columns) {
-	    JsonObject col = conte
-		    //TODO: check db.sql mit vielen daten (100.000)
+	Type columnListType = new TypeToken<Collection<Column>>() {}.getType();
+	retValue.add("columns", context.serialize(src.columns, columnListType));
+	JsonObject xcolumns = new JsonObject();
+	for (Column c : src.columns) {
+	    xcolumns.add(c.columnid, context.serialize(c));
 	}
-	retValue.add("xcolumns", cols);
-	retValue.addProperty("databaseid", databaseid);
-	for (int i = 0; i < tables.size(); ++i) {
-	    retValue.add(tableNames.get(i), context.serialize(tables.get(i)));
+	retValue.add("xcolumns", xcolumns);
+	JsonObject[] data = new JsonObject[src.data.size()];
+	for (int i = 0; i < src.data.size(); ++i) {
+	    JsonObject cur = new JsonObject();
[…]

### 2015-03-03 test.tex
fixed bug with parsing of strings for charts seperated tests

--- 2015-03-02 test.tex
@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{dim_compiler}
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment,
 
@@
 { \centering %
-\lineChart{0.4\textwidth}{0.3\textwidth}{SELECT * FROM data1;}{Money (in €)}{Date}{false}{"title": "Title", "titleHeight": 25} %
+~%\lineChart{0.4\textwidth}{0.3\textwidth}{SELECT * FROM data1;}{Money (in €)}{Date}{false}{"title": "Title", "titleHeight": 25} %
 \\ }
@@
      { \centering %
-     \lineChart{0.3\textwidth}{0.4\textwidth}{SELECT * FROM data1 WHERE date >= '2008/05/11';}{Money (in €)}{Date}{true}{}%
+     ~%\lineChart{0.3\textwidth}{0.4\textwidth}{SELECT * FROM data1 WHERE date >= '2008/05/11';}{Money (in €)}{Date}{true}{}%
      \\ }
@@
     \centering
-    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}{Runtime (in seconds)}{Date}{false}{} \\
+    ~%\lineChart{0.4\textwidth}{0.4\textwidth}{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}{Runtime (in seconds)}{Date}{false}{} \\
     \caption{Graph3}
+\end{figure}
+\begin{figure}[h]
+    \centering
+    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT Zahl1, Zahl2 as some_number FROM pgtest;}{Number}{Row ID}{false}{} \\
+    \caption{Graph4}
 \end{figure}

### 2015-03-09 test.tex
hybrid pdf/html is working fixed some bugs
TODO: komische rahmen im adobe reader

--- 2015-03-02 test.tex
@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time} %dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment,
 
@@
     \centering
-    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT * FROM dim_time;}{Runtime}{Row ID}{false}{} \\
+    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}{Number}{Row ID}{false}{} \\
     \caption{Graph4}

### 2015-03-09 src/pdbf/common/Tools.java

@@ -1,6 +1,7 @@
 package pdbf.common;
 
 import java.io.File;
+import java.io.IOException;
 import java.io.UnsupportedEncodingException;
 import java.nio.charset.Charset;
 
@@ -24,6 +25,12 @@ public class Tools {
 
     public static String encodeStringToBase64Binary(String string) throws UnsupportedEncodingException {
 	return Base64.encodeBase64String(string.getBytes(utf8));
+    }
+    
+    public static String escapeQuotes(File file) throws IOException {
+	String tmp = FileUtils.readFileToString(file);
+	tmp = tmp.replace("\"", "\\\"");
+	return tmp;
     }
 
 }

### 2015-03-10 test.tex
switched to xkeyval package

--- 2015-03-03 test.tex
@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
 
@@
 { \centering %
-\lineChart{0.4\textwidth}{0.3\textwidth}{SELECT * FROM data1;}{Money (in €)}{Date}{false}{"title": "Title", "titleHeight": 25} %
+\lineChart[width=0.4\textwidth, height=0.3\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
 \\ }
@@
      { \centering %
-     \lineChart{0.3\textwidth}{0.4\textwidth}{SELECT * FROM data1 WHERE date >= '2008/05/11';}{Money (in €)}{Date}{true}{}%
+     \lineChart[width=0.3\textwidth, height=0.4\textwidth, xunit=Money (in €), yunit=Date, logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
      \\ }
@@
     \centering
-    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}{Runtime (in seconds)}{Date}{false}{} \\
+    \lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Runtime (in seconds), yunit=Date]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}\\%
     \caption{Graph3}
@@
     \centering
-    \lineChart{0.4\textwidth}{0.4\textwidth}{SELECT * FROM dim_time;}{Runtime}{Row ID}{false}{} \\
+    \lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}\\%
     \caption{Graph4}

### 2015-03-10 src/pdbf/latex/LaTeX_Compiler.java

@@ -71,6 +71,7 @@ public class LaTeX_Compiler {
 
 	// Read JSON
 	GsonBuilder builder = new GsonBuilder();
+	builder.disableHtmlEscaping();
 	builder.registerTypeAdapter(Visualization.class, new VisualizationTypeAdapter());
 	builder.registerTypeAdapter(Alasql.class, new Alasql());
 	builder.registerTypeAdapter(Table.class, new Table());
@@ -148,7 +149,8 @@ public class LaTeX_Compiler {
 	overlays = new Overlay[olist.size()];
 	overlays = olist.toArray(overlays);
 	try {
-	    FileUtils.writeStringToFile(new File("config.json"), gson.toJson(overlays), Tools.utf8);
+	    String json = gson.toJson(overlays);
+	    FileUtils.writeStringToFile(new File("config.json"), json, Tools.utf8);
 	} catch (IOException e1) {
 	    e1.printStackTrace();
 	}

### 2015-03-11 test.tex
added initial support for inputenc utf-8

@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
+%\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
 
@@
      { \centering %
-     \lineChart[width=0.3\textwidth, height=0.4\textwidth, xunit=Money (in €), yunit=Date, logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
+     ~%\lineChart[width=0.3\textwidth, height=0.4\textwidth, xunit=Money (in €), yunit=Date, logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
      \\ }
@@
     \centering
-    \lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Runtime (in seconds), yunit=Date]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}\\%
+    ~%\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Runtime (in seconds), yunit=Date]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}\\%
     \caption{Graph3}
@@
     \centering
-    \lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}\\%
+    ~%\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}\\%
     \caption{Graph4}

### 2015-03-11 out/web/capture.js

@@ -4,7 +4,7 @@ console.log('Usage: ' + system.args[0] + ' webpagefile');
 phantom.exit();
 } 
 
-//try {
+try {
 	var page = require('webpage').create();
 
 	page.open(system.args[1], function() {
@@ -17,7 +17,7 @@ phantom.exit();
 		page.evaluate(function() {
 			document.body.bgColor = 'white';
 		});
-		page.zoomFactor = 2.2;
+		page.zoomFactor = 2.0;
 		page.viewportSize = {
 			width: w,
 			height: h
@@ -30,9 +30,9 @@ phantom.exit();
 			phantom.exit();
 		}, 500);
 	});
-//} catch(e) {
-//	console.log(e);
-//	phantom.exit();
-//}
+} catch(e) {
+	console.log(e);
+	phantom.exit();
+}
 
 

### 2015-03-12 beamer_test.tex
added beamer testfile

\documentclass{beamer}
\usetheme{Frankfurt}
\usepackage[ngerman]{babel}
\usepackage{PDBF}
%\setbeameroption{show only notes}

\mdqoff

\addtobeamertemplate{note page}{\setlength{\parskip}{12pt}}
\makeatletter
\beamertemplatenavigationsymbolsempty
\setbeamerfont{footline}{size=\normalsize}
\setbeamertemplate{footline}{{\par\hspace*{\fill}\vspace*{0.25cm}\insertframenumber~~~}}
\setbeamertemplate{items}[default]
\setbeamertemplate{sections/subsections in toc}[sections numbered]
\setbeamertemplate{subsection in toc}{}
\setbeamercolor{title}{fg=white,bg=red!50!green!70}
\setbeamercolor{frametitle}{fg=white,bg=red!50!green!70}
\setbeamercolor{subsection in toc}{fg=black}
\setbeamercolor{section in toc}{fg=black}
\setbeamercolor{item}{fg=black}
\makeatother

\title{Beamer and PDBF}   
\author{
 Patrick Bender\\
  \texttt{s9pabend@stud.uni-saarland.de}
  \\[5pt]
 Christian Goldstein\\
  \texttt{c.goldstein@goldstein-systemberatung.de}
}
\date{}

\DBSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
\DBSQLFile{somedb.sql}
\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment

\begin{document}

\begin{frame}
\titlepage
\vspace{-2.3cm}
\begin{center}
\includegraphics[scale=0.25]{C:/Users/User/Downloads/UdS-Logo.jpg}
\end{center}
Presenter: Patrick Bender
\end{frame}
\note{bla}

\section{SQL Queries} 
\subsection*{}
\begin{frame}\frametitle{Simple Query}

\end{frame}
\note{bla}

\subsection*{}
\begin{frame}\frametitle{Advanced Query}

\end{frame}
\note{bla}

\section{Charts} 
\subsection*{}
\begin{frame}\frametitle{Line Chart with little data}
\def\test{testtest}
	{\centering
	\lineChart[width=0.4\textwidth, height=0.3\textwidth, xunit=\test Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
	\\}
\end{frame}
\note{bla}

\subsection*{}
\begin{frame}\frametitle{Line Chart with much data}
	{\centering
	\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
	\\}
\end{frame}
\note{bla}

\section*{}
\begin{frame}
	Thank you for your attention!
\end{frame}
\note{bla}

\begin{frame}
	Questions?
\end{frame}
\note{bla}

\end{document}

### 2015-03-12 src/pdbf/html/CompleteRun_HTML.java

@@ -9,11 +9,25 @@ public class CompleteRun_HTML {
     public static void main(String[] args) {
 	LaTeX_Compiler.main(args);
 	HTML_Compiler.main(args);
+	
 	new File("dim.json").delete();
 	new File("config.json").delete();
 	new File("db.sql").delete();
 	new File("db.json").delete();
 	new File("null.png").delete();
+	
+	new File("test.aux").delete();
+	new File("test.log").delete();
+	new File("test.synctex.gz").delete();
+	
+	new File("beamer_test.aux").delete();
+	new File("beamer_test.log").delete();
+	new File("beamer_test.nav").delete();
+	new File("beamer_test.out").delete();
+	new File("beamer_test.snm").delete();
+	new File("beamer_test.toc").delete();
+	new File("beamer_test.synctex.gz").delete();
+	
 	System.out.println("Finished!");
     }
     

### 2015-03-12 beamer_test.tex
moved floating point calculation from latex to java because of performance reasons.

@@
 
-\mdqoff
-
 \addtobeamertemplate{note page}{\setlength{\parskip}{12pt}}
@@
 \setbeamerfont{footline}{size=\normalsize}
-\setbeamertemplate{footline}{{\par\hspace*{\fill}\vspace*{0.25cm}\insertframenumber~~~}}
+\setbeamertemplate{footline}{{\par\vspace*{-2cm}\hspace*{\fill}\vspace*{0.25cm}\insertframenumber~~~}}
 \setbeamertemplate{items}[default]
@@
 \setbeamercolor{item}{fg=black}
+\def\beamer@framenotesbegin{% at beginning of slide
+  \gdef\beamer@noteitems{}%
+  \gdef\beamer@notes{{}}% used to be totally empty.
+}
+\addtobeamertemplate{frametitle}{}{\vspace{0.4cm}}
 \makeatother
@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
 
@@
 \begin{frame}
+\vspace{-1.0cm}
 \titlepage
-\vspace{-2.3cm}
+\vspace{-2.5cm}
 \begin{center}
@@
 \subsection*{}
-\begin{frame}\frametitle{Simple Query}
-
+\begin{frame}\frametitle{Simple Queries} 
+	\sql{SELECT * FROM test;}\\[12pt]
+	\sql{SELECT col2 FROM test WHERE col1=1;}\\[12pt]
+	\sqlmult{INSERT INTO test}{INSERT INTO test VALUES (1, 123);}\\
+	\sqlmult{VALUES (1, 123);}{INSERT INTO test VALUES (1, 123);}
 \end{frame}
@@
 \subsection*{}
-\begin{frame}\frametitle{Advanced Query}
-
+\begin{frame}\frametitle{Advanced Queries}
+\vspace{-0.4cm}
+	\def\longquery{SELECT
+m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
+d.tableschema   AS [Table schema],
+c.compiler      AS Compiler,
+c.optlevel      AS [Opt. level],
+COALESCE(cast(l.chunksize AS VARCHAR), 'n/a') AS [Chunk size (tuples)],
+IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
+IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
+t.timestamp 	AS [Time of experiment],
+ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
+FROM fact_experiment AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_compiler c ON c.id = v.compilerid
+    JOIN dim_layout l ON l.id = v.layoutid
+    JOIN dim_implementation i ON i.id = v.implementationid
+    JOIN dim_query q ON q.id = v.queryid
+    JOIN dim_time t ON t.id = v.timeid
+ORDER BY m.id DESC, d.tableschema, q.querytext,
+    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize
+LIMIT 100;}
+	\begin{scriptsize}
+	\sqlmult{SELECT}{\longquery}\\
+	\sqlmult{m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,}{\longquery}\\
+	\sqlmult{d.tableschema   AS [Table schema],}{\longquery}\\
+	\sqlmult{c.compiler      AS Compiler,}{\longquery}\\
+	\sqlmult{c.optlevel      AS [Opt. level],}{\longquery}\\
+	\sqlmult{COALESCE(cast(l.chunksize AS VARCHAR), 'n/a') AS [Chunk size (tuples)],}{\longquery}\\
+	\sqlmult{IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],}{\longquery}\\
+	\sqlmult{IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],}{\longquery}\\
+	\sqlmult{t.timestamp 	AS [Time of experiment],}{\longquery}\\
+	\sqlmult{ROUND(v.runtime, 3) 		AS [Query time (in Sec)]}{\longquery}\\
+	\sqlmult{FROM fact\_experiment AS v}{\longquery}\\
+	\sqlmult{    JOIN dim\_machine m ON m.id = v.machineid}{\longquery}\\
+	\sqlmult{    JOIN dim\_dataset d ON d.id = v.datasetid}{\longquery}\\
+	\sqlmult{    JOIN dim\_compiler c ON c.id = v.compilerid}{\longquery}\\
+	\sqlmult{    JOIN dim\_layout l ON l.id = v.layoutid}{\longquery}\\
+	\sqlmult{    JOIN dim\_implementation i ON i.id = v.implementationid}{\longquery}\\
+	\sqlmult{    JOIN dim\_query q ON q.id = v.queryid}{\longquery}\\
+	\sqlmult{    JOIN dim\_time t ON t.id = v.timeid}{\longquery}\\
+	\sqlmult{ORDER BY m.id DESC, d.tableschema, q.querytext,}{\longquery}\\
+	\sqlmult{    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize}{\longquery}\\
+	\sqlmult{LIMIT 100;}{\longquery}\\
+	\end{scriptsize}
 \end{frame}
@@
 \begin{frame}\frametitle{Line Chart with little data}
-\def\test{testtest}
 	{\centering
-	\lineChart[width=0.4\textwidth, height=0.3\textwidth, xunit=\test Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
+	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
 	\\}
@@
 	{\centering
-	\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
+	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
 	\\}
@@
 \begin{frame}
-	Thank you for your attention!
+\vspace{1.0cm}
+	\begin{center}Thank you for your attention!\end{center}
 \end{frame}
@@
 \begin{frame}
-	Questions?
+\vspace{1.0cm}
+	\begin{center}Questions?\end{center}
 \end{frame}

### 2015-03-12 test.tex
moved floating point calculation from latex to java because of performance reasons.

@@
 Test Queries:\\[3pt]
-\sql{SELECT * FROM test;}{SELECT * FROM test;}\\[3pt]
-\sql{SELECT col2 FROM test WHERE col1=1;}{SELECT col2 FROM test WHERE col1=1;}\\[3pt]
-\sql{INSERT INTO test}{INSERT INTO test VALUES (1, 123);}\\
-\sql{VALUES (1, 123);}{INSERT INTO test VALUES (1, 123);}\\[8pt]
-Lorem ipsum \sql{SELECT * FROM test;}{SELECT * FROM test;} dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
+\sql{SELECT * FROM test;}\\[3pt]
+\sql{SELECT col2 FROM test WHERE col1=1;}\\[3pt]
+\sqlmult{INSERT INTO test}{INSERT INTO test VALUES (1, 123);}\\
+\sqlmult{VALUES (1, 123);}{INSERT INTO test VALUES (1, 123);}\\[8pt]
+Lorem ipsum \sql{SELECT * FROM test;} dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
 At vero eos et accusam et justo duo dolores et ea rebum.

### 2015-03-13 beamer_test.tex
scaling is working

--- 2015-03-12 beamer_test.tex
@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
 
@@
 \subsection*{}
-\begin{frame}\frametitle{Advanced Query}
-
+\begin{frame}\frametitle{Advanced Queries}
+%\vspace{-0.4cm}
+%	\def\longquery{SELECT
+%m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
+%d.tableschema   AS [Table schema],
+%c.compiler      AS Compiler,
+%c.optlevel      AS [Opt. level],
+%COALESCE(cast(l.chunksize AS VARCHAR), 'n/a') AS [Chunk size (tuples)],
+%IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
+%IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
+%t.timestamp 	AS [Time of experiment],
+%ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
+%FROM fact_experiment AS v
+%    JOIN dim_machine m ON m.id = v.machineid
+%    JOIN dim_dataset d ON d.id = v.datasetid
+%    JOIN dim_compiler c ON c.id = v.compilerid
+%    JOIN dim_layout l ON l.id = v.layoutid
+%    JOIN dim_implementation i ON i.id = v.implementationid
+%    JOIN dim_query q ON q.id = v.queryid
+%    JOIN dim_time t ON t.id = v.timeid
+%ORDER BY m.id DESC, d.tableschema, q.querytext,
+%    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize
+%LIMIT 100;}
+%	\begin{scriptsize}
+%	\sqlmult{SELECT}{\longquery}\\
+%	\sqlmult{m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,}{\longquery}\\
+%	\sqlmult{d.tableschema   AS [Table schema],}{\longquery}\\
+%	\sqlmult{c.compiler      AS Compiler,}{\longquery}\\
+%	\sqlmult{c.optlevel      AS [Opt. level],}{\longquery}\\
+%	\sqlmult{COALESCE(cast(l.chunksize AS VARCHAR), 'n/a') AS [Chunk size (tuples)],}{\longquery}\\
+%	\sqlmult{IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],}{\longquery}\\
+%	\sqlmult{IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],}{\longquery}\\
+%	\sqlmult{t.timestamp 	AS [Time of experiment],}{\longquery}\\
+%	\sqlmult{ROUND(v.runtime, 3) 		AS [Query time (in Sec)]}{\longquery}\\
+%	\sqlmult{FROM fact\_experiment AS v}{\longquery}\\
+%	\sqlmult{    JOIN dim\_machine m ON m.id = v.machineid}{\longquery}\\
+%	\sqlmult{    JOIN dim\_dataset d ON d.id = v.datasetid}{\longquery}\\
+%	\sqlmult{    JOIN dim\_compiler c ON c.id = v.compilerid}{\longquery}\\
+%	\sqlmult{    JOIN dim\_layout l ON l.id = v.layoutid}{\longquery}\\
+%	\sqlmult{    JOIN dim\_implementation i ON i.id = v.implementationid}{\longquery}\\
+%	\sqlmult{    JOIN dim\_query q ON q.id = v.queryid}{\longquery}\\
+%	\sqlmult{    JOIN dim\_time t ON t.id = v.timeid}{\longquery}\\
+%	\sqlmult{ORDER BY m.id DESC, d.tableschema, q.querytext,}{\longquery}\\
+%	\sqlmult{    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize}{\longquery}\\
+%	\sqlmult{LIMIT 100;}{\longquery}\\
+%	\end{scriptsize}
 \end{frame}
@@
 \subsection*{}
+\mdqoff
 \begin{frame}\frametitle{Line Chart with little data}
-\def\test{testtest}
-	{\centering
-	\lineChart[width=0.4\textwidth, height=0.3\textwidth, xunit=\test Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
+	{ \centering 
+	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 20}]{SELECT * FROM data1;}%
 	\\}
 \end{frame}
+\mdqon
 \note{bla}
@@
 	{\centering
-	\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
+	~%\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
 	\\}
@@
 \begin{frame}
-	Thank you for your attention!
+\vspace{1.0cm}
+	\begin{center}\sql{SELECT * FROM test;}\\Thank you for your attention!\end{center}
 \end{frame}

### 2015-03-13 out/web/capture.js

@@ -4,7 +4,7 @@ console.log('Usage: ' + system.args[0] + ' webpagefile');
 phantom.exit();
 } 
 
-try {
+//try {
 	var page = require('webpage').create();
 
 	page.open(system.args[1], function() {
@@ -30,9 +30,9 @@ try {
 			phantom.exit();
 		}, 500);
 	});
-} catch(e) {
-	console.log(e);
-	phantom.exit();
-}
+//} catch(e) {
+//	console.log(e);
+//	phantom.exit();
+//}
 
 

### 2015-03-16 test.tex
scaling is working
!

@@
-\documentclass{scrartcl}
+\documentclass[10pt]{scrartcl}
 \usepackage{PDBF}
@@
 \DBSQLFile{somedb.sql}
-%\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest}% , dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
+
+\renewcommand\pdbfScale{0.833}
 
@@
 { \centering %
-\lineChart[width=0.4\textwidth, height=0.3\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
+\lineChart[width=0.4\textwidth, height=0.3\textwidth,  options={"title": "Title"}]{SELECT * FROM data1;}%
 \\ }
@@
      { \centering %
-     ~%\lineChart[width=0.3\textwidth, height=0.4\textwidth, xunit=Money (in €), yunit=Date, logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
+     \lineChart[width=0.3\textwidth, height=0.4\textwidth, xunit=Date, yunit=Money (in €), logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
      \\ }
@@
     \centering
-    ~%\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Runtime (in seconds), yunit=Date]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}\\%
+    \lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}\\%
     \caption{Graph3}
@@
     \centering
-    ~%\lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}\\%
+    \lineChart[width=1.0\textwidth, height=0.5\textwidth, xunit=Row ID, yunit=Number]{SELECT Zahl1, Zahl2 as some_number FROM pgtest;}\\%
     \caption{Graph4}

### 2015-03-16 out/web/capture.js

@@ -17,7 +17,6 @@ phantom.exit();
 		page.evaluate(function() {
 			document.body.bgColor = 'white';
 		});
-		page.zoomFactor = 3.9;
 		page.viewportSize = {
 			width: w,
 			height: h
@@ -26,7 +25,7 @@ phantom.exit();
 			var name = page.evaluate(function() {
 				return json.name;
 			});
-			page.render(name +'.png');
+			page.render(name +'.png', {format: 'png', quality: '100'});
 			phantom.exit();
 		}, 500);
 	});

### 2015-03-16 src/pdbf/latex/LaTeX_Compiler.java

@@ -34,6 +34,7 @@ public class LaTeX_Compiler {
     private static ArrayList<String> cleanupfiles = new ArrayList<String>();
     private static Gson gson;
     private static String suffix;
+    private static Dimension dimOrg;
 
     public static void main(String[] args) {
 	if (OS.contains("win")) {
@@ -83,28 +84,23 @@ public class LaTeX_Compiler {
 	    String json = FileUtils.readFileToString(new File("config.json"), Tools.utf8);
 	    overlays = gson.fromJson(json, Overlay[].class);
 	    String json2 = FileUtils.readFileToString(new File("dim.json"), Tools.utf8);
-	    Dimension dim = gson.fromJson(json2, Dimension.class);
-	    //TODO: do the calculations to percent from latex
-	    //TODO: chrome not working? why?
-		for (int i = 0; i < overlays.length; ++i) {
-		    if (overlays[i].type instanceof Chart) {
-			Chart c = (Chart)overlays[i].type;
-			c.x1 = c.x1 / dim.width;
-			c.x2 = c.x2 / dim.width;
-			c.y1 = c.y1 / dim.height;
-			c.y2 = c.y2 / dim.height;
-		    }
-		    if (overlays[i].type instanceof Text) {
-			Text t = (Text)overlays[i].type;
-			t.x1 = t.x1 / dim.width;
-			t.x2 = t.x2 / dim.width;
-			t.y1 = (t.y1 + 65536 * t.fontsize) / dim.height;
-			t.y2 = t.y2 / dim.height;
-		    }
+	    dimOrg = gson.fromJson(json2, Dimension.class);
+	    for (int i = 0; i < overlays.length; ++i) {
+		if (overlays[i].type instanceof Chart) {
+		    Chart c = (Chart) overlays[i].type;
+		    c.x1 = c.x1 / dimOrg.width;
+		    c.x2 = c.x2 / dimOrg.width;
+		    c.y1 = c.y1 / dimOrg.height;
+		    c.y2 = c.y2 / dimOrg.height;
 		}
-
-	    
-
+		if (overlays[i].type instanceof Text) {
+		    Text t = (Text) overlays[i].type;
+		    t.x1 = t.x1 / dimOrg.width;
+		    t.x2 = t.x2 / dimOrg.width;
+		    t.y1 = (t.y1 + 65536 * t.fontsize) / dimOrg.height;
+		    t.y2 = t.y2 / dimOrg.height;
+		}
+	    }
[…]

### 2015-03-23 beamer_test.tex
!

--- 2015-03-12 beamer_test.tex
@@
-\documentclass{beamer}
+\documentclass[10pt]{beamer}
 \usetheme{Frankfurt}
-\usepackage[ngerman]{babel}
 \usepackage{PDBF}
-%\setbeameroption{show only notes}
+\setbeameroption{hide notes}
 
@@
 \setbeamercolor{item}{fg=black}
+\setbeamertemplate{caption}[numbered]
+\setbeamercolor{caption name}{fg=black}
 \def\beamer@framenotesbegin{% at beginning of slide
@@
 
-\title{Beamer and PDBF}   
+\title{PDBF}   
 \author{
@@
 
+\renewcommand\pdbfScale{0.833}
+
 \DBSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
 
@@
 \end{frame}
-\note{bla}
+
+\section{Goals} 
+\subsection*{}
+\vspace{1.0cm}
+\begin{frame}
+	\begin{Large}\begin{center}What are the goals?\end{center}\end{Large}
+\end{frame}
 
@@
 \end{frame}
-\note{bla}
 
@@
 \begin{frame}\frametitle{Line Chart with little data}
-	{\centering
-	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title", "titleHeight": 25}]{SELECT * FROM data1;}%
+	{ \centering 
+	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title"}]{SELECT * FROM data1;}%
 	\\}
 \end{frame}
-\note{bla}
 
@@
 	{\centering
-	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Number, yunit=Row ID]{SELECT Zahl1, Zahl2 as some_number FROM pgtest WHERE Zahl1 > 45000;}%
+	    \lineChart[width=1.0\textwidth, height=0.5\textwidth, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}%
 	\\}
 \end{frame}
-\note{bla}
+
+\subsection*{}
+\begin{frame}\frametitle{Other Chart types}
+	\begin{figure}[h]
+    \centering
+    \begin{minipage}{.48\textwidth}
+    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}%
+    \caption{Multi-column Line Chart}
+    \end{minipage}
+    \hspace{5pt}
+    \begin{minipage}{.48\textwidth}
+    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}%
+    \caption{Multi-column Bar Chart}
+    \end{minipage}\\[8pt]
+    \end{figure}
+\end{frame}
+
+\subsection*{}
+\begin{frame}\frametitle{Other Chart types}
+	\begin{figure}[h]
+    \centering
+    \begin{minipage}{.48\textwidth}
+    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}, overlap=2]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \caption{Multi-column overlapping Bar Chart}
+    \end{minipage}
+    \hspace{5pt}
+    \begin{minipage}{.48\textwidth}
+    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}, options={"strokeWidth": 0, "drawPoints": true, "pointSize": 3, "highlightCircleSize": 4, "xRangePad": 10}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \caption{Multi-column Dot Chart}
+    \end{minipage}
+    \end{figure}
+\end{frame}
 
@@
 \vspace{1.0cm}
-	\begin{center}Thank you for your attention!\end{center}
+	\begin{Large}\begin{center}Thank you for your attention!\end{center}\end{Large}
 \end{frame}
-\note{bla}
 
@@
 \vspace{1.0cm}
-	\begin{center}Questions?\end{center}
+	\begin{Large}\begin{center}Questions?\end{center}\end{Large}
 \end{frame}
-\note{bla}
 

### 2015-03-23 test.tex
!

@@
   \texttt{s9pabend@stud.uni-saarland.de}
-  \and
- Christian Goldstein\\
-  \texttt{c.goldstein@goldstein-systemberatung.de}
 }
@@
 
+\renewcommand\pdbfScale{0.833}
+
 \DBSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, pgtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
-
-\renewcommand\pdbfScale{0.833}
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
 
@@
 At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
 Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.\\
@@
 Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait facilisi.
+Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
+
 \newpage
-\noindent %
-Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
-\\[8pt]
-\begin{figure}[h]
+
+\begin{figure}[h!]
     \centering
-    \lineChart[width=0.4\textwidth, height=0.4\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}\\%
-    \caption{Graph3}
-\end{figure}
-\begin{figure}[h]
-    \centering
-    \lineChart[width=1.0\textwidth, height=0.5\textwidth, xunit=Row ID, yunit=Number]{SELECT Zahl1, Zahl2 as some_number FROM pgtest;}\\%
-    \caption{Graph4}
+    \begin{minipage}{.48\textwidth}
+    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}
+    \caption{Multi-column Line Chart}
+    \end{minipage}
+    \hspace{11pt}
+    \begin{minipage}{.48\textwidth}
+    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \caption{Multi-column Bar Chart}
+    \end{minipage} \\[8pt]
+
+    \begin{minipage}{.48\textwidth}
+    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}, overlap=2]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \caption{Multi-column overlapping Bar Chart}
+    \end{minipage}
+    \hspace{11pt}
+    \begin{minipage}{.48\textwidth}
+    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}, options={"strokeWidth": 0, "drawPoints": true, "pointSize": 3, "highlightCircleSize": 4, "xRangePad": 10}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \caption{Multi-column Dot Chart}
+    \end{minipage} \\[8pt]
+
+    \begin{minipage}{\textwidth}
+    \lineChart[width=1.0\textwidth, height=0.45\textwidth, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}\\%
+    \caption{Lot of points}
+    \end{minipage}
 \end{figure}
 
+\newpage
+\begin{wrapfigure}{r}{0.52\textwidth}
+{ \centering %
+\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}]
+{SELECT
+m.cpuarchitecture AS Machine,
+IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
+IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
+c.compiler      AS Compiler,
+c.optlevel      AS [Opt. level],
+l.name          AS [Layout name],
+CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provi- ded at],
+ROUND(v.runtime, 3) 		AS [Query time]
+FROM fact_experiment AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_compiler c ON c.id = v.compilerid
+    JOIN dim_layout l ON l.id = v.layoutid
+    JOIN dim_implementation i ON i.id = v.implementationid
+    JOIN dim_query q ON q.id = v.queryid
+ORDER BY m.id DESC, d.tableschema, q.querytext,
+    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+{SELECT
+m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
+d.tableschema   AS [Table schema],
+q.querytext     AS [Query],
+c.compiler      AS Compiler,
+c.optlevel      AS [Opt. level],
+CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) AS [Chunk size (tuples)],
+IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
+IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
+t.timestamp 	AS [Time of experiment],
+ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
+FROM fact_experiment AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_compiler c ON c.id = v.compilerid
+    JOIN dim_layout l ON l.id = v.layoutid
+    JOIN dim_implementation i ON i.id = v.implementationid
+    JOIN dim_query q ON q.id = v.queryid
+    JOIN dim_time t ON t.id = v.timeid
+ORDER BY m.id DESC, d.tableschema, q.querytext,
+    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+\\ }
+\vspace{-5pt}
+\caption{Pivot table}
+\vspace{-23pt}
+\end{wrapfigure}
+Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.\\[8pt]
+Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
+Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait facilisi.\\[8pt]
+Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
+At vero eos et accusam et justo duo dolores et ea rebum.
+Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
+Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
+At vero eos et accusam et justo duo dolores et ea rebum.
+Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
+Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper suscipit lobortis nisl ut aliquip ex ea commodo consequat.
+Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto odio dignissim qui blandit praesent luptatum zzril delenit augue duis dolore te feugait facilisi.
+
 \end{document}

### 2015-03-23 out/web/pivotBranch.js

@@ -1,0 +1,1541 @@
+(function() {
+  var callWithJQuery,
+    __indexOf = [].indexOf || function(item) { for (var i = 0, l = this.length; i < l; i++) { if (i in this && this[i] === item) return i; } return -1; },
+    __slice = [].slice,
+    __bind = function(fn, me){ return function(){ return fn.apply(me, arguments); }; },
+    __hasProp = {}.hasOwnProperty;
+
+  callWithJQuery = function(pivotModule) {
+    if (typeof exports === "object" && typeof module === "object") {
+      return pivotModule(require("jquery"));
+    } else if (typeof define === "function" && define.amd) {
+      return define(["jquery"], pivotModule);
+    } else {
+      return pivotModule(jQuery);
+    }
+  };
+
+  callWithJQuery(function($) {
+
+    /*
+    Utilities
+     */
+    var PivotData, addSeparators, aggregatorTemplates, aggregators, dayNamesEn, derivers, getSort, locales, mthNamesEn, naturalSort, numberFormat, pivotTableRenderer, renderers, sortAs, usFmt, usFmtInt, usFmtPct, zeroPad;
+    addSeparators = function(nStr, thousandsSep, decimalSep) {
+      var rgx, x, x1, x2;
+      nStr += '';
+      x = nStr.split('.');
+      x1 = x[0];
+      x2 = x.length > 1 ? decimalSep + x[1] : '';
+      rgx = /(\d+)(\d{3})/;
+      while (rgx.test(x1)) {
+        x1 = x1.replace(rgx, '$1' + thousandsSep + '$2');
+      }
+      return x1 + x2;
+    };
+    numberFormat = function(opts) {
+      var defaults;
+      defaults = {
+        digitsAfterDecimal: 2,
+        scaler: 1,
+        thousandsSep: ",",
+        decimalSep: ".",
+        prefix: "",
+        suffix: "",
+        showZero: false
+      };
+      opts = $.extend(defaults, opts);
+      return function(x) {
+        var result;
+        if (isNaN(x) || !isFinite(x)) {
+          return "";
+        }
+        if (x === 0 && !opts.showZero) {
+          return "";
+        }
+        result = addSeparators((opts.scaler * x).toFixed(opts.digitsAfterDecimal), opts.thousandsSep, opts.decimalSep);
+        return "" + opts.prefix + result + opts.suffix;
+      };
+    };
+    usFmt = numberFormat();
+    usFmtInt = numberFormat({
[…]

### 2015-03-23 src/pdbf/common/Visualization.java

@@ -2,6 +2,7 @@ package pdbf.common;
 
 public abstract class Visualization {
     public String query;
+    public String queryB;
     public double x1;
     public double x2;
     public double y1;

### 2015-03-23 src/pdbf/html/CompleteRun_HTML.java

@@ -9,17 +9,17 @@ public class CompleteRun_HTML {
     public static void main(String[] args) {
 	LaTeX_Compiler.main(args);
 	HTML_Compiler.main(args);
-	
+
 	new File("dim.json").delete();
 	new File("config.json").delete();
 	new File("db.sql").delete();
 	new File("db.json").delete();
 	new File("null.png").delete();
-	
+
 	new File("test.aux").delete();
 	new File("test.log").delete();
 	new File("test.synctex.gz").delete();
-	
+
 	new File("beamer_test.aux").delete();
 	new File("beamer_test.log").delete();
 	new File("beamer_test.nav").delete();
@@ -27,8 +27,8 @@ public class CompleteRun_HTML {
 	new File("beamer_test.snm").delete();
 	new File("beamer_test.toc").delete();
 	new File("beamer_test.synctex.gz").delete();
-	
+
 	System.out.println("Finished!");
     }
-    
+
 }

### 2015-03-23 src/pdbf/tests/CreateConfigJSON.java

@@ -1,61 +1,0 @@
-package pdbf.tests;
-
-import java.io.File;
-import java.util.ArrayList;
-import org.apache.commons.io.FileUtils;
-
-import pdbf.common.LineChart;
-import pdbf.common.Overlay;
-import pdbf.common.Text;
-import pdbf.common.Tools;
-import pdbf.common.Visualization;
-import pdbf.common.VisualizationTypeAdapter;
-
-import com.google.gson.Gson;
-import com.google.gson.GsonBuilder;
-
-public class CreateConfigJSON {
-
-    public static void main(String[] args) {
-	GsonBuilder builder = new GsonBuilder();
-	builder.setPrettyPrinting();
-	builder.registerTypeAdapter(Visualization.class,
-		new VisualizationTypeAdapter());
-	builder.disableHtmlEscaping();
-	Gson gson = builder.create();
-
-	ArrayList<Overlay> overlays = new ArrayList<Overlay>();
-
-	LineChart chart = new LineChart("SELECT * FROM data2 WHERE Zahl2 > 0;",
-		0.5150353982300885, 0.8407079646017699, 0.2496875, 0.3601875,
-		1, false, "Zahl1", "Zahl2", 1.0, 1.0, "topright", "{}");
-	Overlay cur = new Overlay("Tabelle1", chart);
-	overlays.add(cur);
-
-	Text text = new Text("SELECT * FROM test;",
-		0.12831858407079646017699115044248,
-		0.33185840707964601769911504424779, 0.2515625, 0.2671875, 1);
-	cur = new Overlay("SQL1", text);
-	overlays.add(cur);
-
-	text = new Text("SELECT col2 FROM test WHERE col1 = 1;",
-		0.12831858407079646017699115044248,
-		0.34955752212389380530973451327434, 0.2671875, 0.2984375, 1);
-	cur = new Overlay("SQL2", text);
-	overlays.add(cur);
-
-	text = new Text("INSERT INTO test VALUES (1, 123);",
-		0.12831858407079646017699115044248,
-		0.30088495575221238938053097345133, 0.2984375, 0.334375, 1);
-	cur = new Overlay("SQL3", text);
-	overlays.add(cur);
-
-	try {
-	    String out = gson.toJson(overlays);
-	    FileUtils.writeStringToFile(new File("config.json"), out, Tools.utf8);
-	} catch (Exception e) {
-	    e.printStackTrace();
-	}
-    }
-    
-}

### 2015-03-27 test.tex
arbeit arbeit

@@
 { \centering %
-\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}]
+\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
 {SELECT
@@
 c.optlevel      AS [Opt. level],
+l.name          AS [Layout name],
 CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) AS [Chunk size (tuples)],

### 2015-03-27 src/pdbf/common/BarChart.java

@@ -1,5 +1,5 @@
 package pdbf.common;
 
 public class BarChart extends Chart {
-    public int overlap;
+    
 }

### 2015-03-27 Documentation.tex

\documentclass[11pt]{scrartcl}
\usepackage{PDBF}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage[a4paper,vmargin={28mm,28mm},hmargin={28mm,28mm}]{geometry}

\title{PDBF Documentation}
\author{
 Patrick Bender\\
  \texttt{s9pabend@stud.uni-saarland.de}
  \and
 Christian Goldstein\\
  \texttt{c.goldstein@goldstein-systemberatung.de}
}
\date{\today{}, Saarbrücken}

\makeatletter
\def\@maketitle{%
  \vspace*{-\topskip}%
  \begingroup\centering%
  \let \footnote \thanks
  \hrule height \z@%
    {\LARGE \@title \par}%
    \vskip 1.5em 
    {\large
      \lineskip .5em 
      \begin{tabular}[t]{c}%
        \@author
      \end{tabular}\par}%
    \vskip 1em 
    {\large \@date}%
  \par\endgroup%
  \vskip 1.5em%
}
\makeatother

\urlstyle{rm}

\begin{document}
\maketitle

TODO: Write small abstract \newpage

\section{Chart options}
Macros: \\[3pt]
\verb|\lineChart[options]{query}| \\
\verb|\barChart[options]{query}| \\[8pt]
\noindent Options: \\[3pt]
\begin{tabular}{ll}
\textbf{width} & \\
Desc.:	&	Sets the width of the chart. \\
Argument:	&LaTeX length \\
Default:&	no default value. value must always be set! \\[4pt]

\textbf{height} \\
Desc.:	&	Sets the height of the chart. \\
Argument: &	LaTeX length \\
Default:&	no default value. value must always be set! \\[4pt]

\textbf{xunit} \\
Desc.:	&	Sets the a label for the x axis. \\
Argument: &	String \\
Default:& 	\verb|""| (which means hide) \\[4pt]

\textbf{yunit} \\
Desc.:	&	Sets the a label for the y axis. \\
Argument: 	&String \\
Default:& 	\verb|""| (which means hide) \\[4pt]

\textbf{options} \\
Desc.:	&	Sets options that are directly passed to the dygraphs library (\href{http://dygraphs.com/options.html}{$\rightarrow$ Documentation}). \\
& You need to wrap the JSON-String with \verb|{}| if you want to use \verb|[| or \verb|]| or \verb|,|. \\
Argument: 	&JSON-String \\
Default:&	\verb|{}| (which means empty object) \\[4pt]

\textbf{logscale} \\
Desc.:	&	If set, the y axis uses log scale. \\
& Can also be redefined globally (\verb|\pdbfLogscale|) \\
Argument:&	\verb|{true, false}| \\
Default:&	\verb|false| \\[4pt]

\textbf{quality} \\
Desc.:	&	Sets the quality for the image version of the chart. \\
& Can also be redefined globally (\verb|\pdbfQuality|). \\
Argument:	&Number $>$ 0 \\
Default:&	\verb|1.0| \\[4pt]

\textbf{scale} \\
Desc.:	&	Sets the scale for the chart. 1.0 corresponds roughly to 12pt in \LaTeX. \\
& Can also be redefined globally (\verb|\pdbfScale|). \\
Argument:	&Number $>$ 0 \\
Default:&	\verb|1.0| \\[4pt]

\textbf{legendpos} \\
Desc.: 	&	Sets the position of the legend of the chart. \\
& Can also be redefined globally (\verb|\pdbfLegendpos|).\\
Argument:&	\verb|{topright, topcenter, topleft, belowleft, belowcenter, belowright}| \\
Default:&	\verb|topright| \\[4pt]
\end{tabular} \\

\newpage
\section{F.A.Q.}
\begin{itemize}
\item F.: The chart is horribly slow when i try to zoom! \\
A.: Add the following to options: animatedZooms: false 
\end{itemize}

\end{document}

### 2015-03-30 external-tools/capture.js
!

@@ -1,0 +1,37 @@
+var system = require('system');
+if (system.args.length !== 2) {
+console.log('Usage: ' + system.args[0] + ' webpagefile');
+phantom.exit();
+} 
+
+//try {
+	var page = require('webpage').create();
+
+	page.open(system.args[1], function() {
+		var w = page.evaluate(function() {
+			return outw;
+		});
+		var h = page.evaluate(function() {
+			return outh;
+		});
+		page.evaluate(function() {
+			document.body.bgColor = 'white';
+		});
+		page.viewportSize = {
+			width: w,
+			height: h
+		};
+		window.setTimeout(function () {
+			var name = page.evaluate(function() {
+				return json.name;
+			});
+			page.render(name +'.png', {format: 'png', quality: '100'});
+			phantom.exit();
+		}, 500);
+	});
+//} catch(e) {
+//	console.log(e);
+//	phantom.exit();
+//}
+
+

### 2015-03-30 out/build/pdf.js

@@ -1654,8 +1654,7 @@ PDFJS.imageResourcesPath = (PDFJS.imageResourcesPath === undefined ?
  * to workers.
  * @var {boolean}
  */
-PDFJS.disableWorker = (PDFJS.disableWorker === undefined ?
-                       false : PDFJS.disableWorker);
+PDFJS.disableWorker = true;
 
 /**
  * Path and filename of the worker file. Required when the worker is enabled in
@@ -2399,9 +2398,7 @@ var WorkerTransport = (function WorkerTransportClosure() {
         // In the developer build load worker_loader which in turn loads all the
         // other files and resolves the promise. In production only the
         // pdf.worker.js file is needed.
-        Util.loadScript(PDFJS.workerSrc, function() {
-          PDFJS.fakeWorkerFilesLoadedCapability.resolve();
-        });
+        PDFJS.fakeWorkerFilesLoadedCapability.resolve();
       }
       PDFJS.fakeWorkerFilesLoadedCapability.promise.then(function () {
         warn('Setting up fake worker.');

### 2015-03-30 out/web/capture.js

@@ -1,37 +1,0 @@
-var system = require('system');
-if (system.args.length !== 2) {
-console.log('Usage: ' + system.args[0] + ' webpagefile');
-phantom.exit();
-} 
-
-//try {
-	var page = require('webpage').create();
-
-	page.open(system.args[1], function() {
-		var w = page.evaluate(function() {
-			return outw;
-		});
-		var h = page.evaluate(function() {
-			return outh;
-		});
-		page.evaluate(function() {
-			document.body.bgColor = 'white';
-		});
-		page.viewportSize = {
-			width: w,
-			height: h
-		};
-		window.setTimeout(function () {
-			var name = page.evaluate(function() {
-				return json.name;
-			});
-			page.render(name +'.png', {format: 'png', quality: '100'});
-			phantom.exit();
-		}, 500);
-	});
-//} catch(e) {
-//	console.log(e);
-//	phantom.exit();
-//}
-
-

### 2015-03-30 src/pdbf/common/Table.java

@@ -2,14 +2,4 @@ package pdbf.common;
 
 public class Table extends Visualization {
 
-    public Table(String query, double x1, double x2, double y1, double y2, long page) {
-	super(query, x1, x2, y1, y2, page);
-    }
-
-    /*
-     * Needed for GSON
-     */
-    public Table() {
-    }
-
 }

### 2015-03-30 src/pdbf/html/HTML_Compiler.java

@@ -17,17 +17,19 @@ public class HTML_Compiler {
 	
 	String basename = args[0].substring(0, args[0].length()-4);
 	String pdfname = basename + ".pdf";
-	String outfile = "out/web/" + basename + ".html";
-		
+	String outfile = basename + ".html";
+	
 	try {
 	    String viewer;
 	    String viewerHEAD = FileUtils.readFileToString(new File("out/web/templateHEADalasql.html"), Tools.utf8);
 	    String viewerTAIL = FileUtils.readFileToString(new File("out/web/templateTAILalasql.html"), Tools.utf8);
+	    String all = FileUtils.readFileToString(new File("out/web/all"), Tools.utf8);
 	    viewer = viewerHEAD + 
 		    "pdf_base64 = \"" + Tools.encodeFileToBase64Binary(new File(pdfname)) + "\";\r\n" + 
 		    "db_base64 = \"" + Tools.encodeFileToBase64Binary(new File("db.sql")) + "\";\r\n" + 
 		    "json_base64 = \"" + Tools.encodeFileToBase64Binary(new File("config.json")) + "\";\r\n" + 
 		    "dbjson_base64 = \"" + Tools.escapeQuotes(new File("db.json")) + "\";\r\n" +
+		    "\r\n" + all +		    
 		    viewerTAIL;
 	    String insert1 = "%<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>" +
   		      	     "<head><meta charset=\"utf-8\"><!--\n";

### 2015-03-30 src/pdbf/html/MinifyResources.java

@@ -1,0 +1,44 @@
+package pdbf.html;
+
+import java.io.File;
+import java.io.IOException;
+import java.util.Arrays;
+
+import org.apache.commons.io.FileUtils;
+
+import pdbf.common.Tools;
+
+public class MinifyResources {
+
+    public static void main(String[] args) {
+	String command[] = {"java", "-jar", "lib/compiler.jar", "out/web/base64.js", "out/web/compatibility.js", "out/web/dygraphs.js", "out/web/alasql.js", 
+		"out/web/excanvas.compiled.js", "out/web/diff_match_patch.js", "out/web/codemirror-compressed.js",
+		"out/web/jquery-1.11.2.min.js", "out/web/pivot.js", "out/web/jquery-ui-1.9.2.custom.min.js",
+		"out/web/l10n.js", "out/web/viewer.js", "out/web/main.js",
+		"--js_output_file", "out/web/all", "--language_in", "ECMASCRIPT5", "--compilation_level", 
+		"WHITESPACE_ONLY", "--charset", "UTF-8"}; 
+	
+	try {
+	    ProcessBuilder pb = new ProcessBuilder(Arrays.asList(command));
+	    pb.inheritIO();
+	    Process p = pb.start();
+	    p.waitFor();
+	} catch (Exception e) {
+	    e.printStackTrace();
+	}
+	
+	try {
+	    String pdfJS = FileUtils.readFileToString(new File("out/build/pdf.js"), Tools.utf8);
+	    String pdfworkerJS = FileUtils.readFileToString(new File("out/build/pdf.worker.js"), Tools.utf8);
+	    String viewerCSS = FileUtils.readFileToString(new File("out/web/viewer.css"), Tools.utf8);
+	    String pivotCSS = FileUtils.readFileToString(new File("out/web/pivot.css"), Tools.utf8);
+	    String codemirrorCSS = FileUtils.readFileToString(new File("out/web/codemirror.css"), Tools.utf8);
+	    String out = pdfJS + pdfworkerJS + "</script><style>" + viewerCSS + pivotCSS + codemirrorCSS + "</style>";
+	    FileUtils.writeStringToFile(new File("out/web/all"), out, Tools.utf8, true);
+	} catch (IOException e) {
+	    e.printStackTrace();
+	}
+	//TODO: Minify HTML and css
+    }
+
+}

### 2015-03-30 src/pdbf/latex/LaTeX_Compiler.java

@@ -32,6 +32,7 @@ public class LaTeX_Compiler {
 
     private static ArrayList<Process> processes = new ArrayList<Process>();
     private static ArrayList<String> cleanupfiles = new ArrayList<String>();
+    private static ArrayList<String> copyfiles = new ArrayList<String>();
     private static Gson gson;
     private static String suffix;
     private static Dimension dimOrg;
@@ -49,7 +50,7 @@ public class LaTeX_Compiler {
 	}
 
 	if (args.length != 1) {
-	    System.out.println("Usage: Java_Compiler.jar pathToLaTeX");
+	    System.out.println("Usage: Java_Compiler.jar LaTeX_file");
 	    System.exit(-1);
 	}
 	String latexPath = args[0];
@@ -59,6 +60,8 @@ public class LaTeX_Compiler {
 	    System.exit(-1);
 	}
 
+	String latexFolder = latex.getParent();
+	
 	ArrayList<String> commands = new ArrayList<String>(Arrays.asList(pathToLaTeXScript));
 	commands.add(latex.getAbsolutePath());
 
@@ -150,9 +153,18 @@ public class LaTeX_Compiler {
 	    }
 	}
 	
+	//copy images to latex file
+	for (String s : copyfiles) {
+	    File source = new File(s);
+	    String target = latexFolder + File.separator + source.getName();
+	    try {
+		FileUtils.copyFile(source, new File(target));
+	    } catch (IOException e) {
+		e.printStackTrace();
+	    }
+	}
+	
 	//TODO: javascript. maybe implement chart legend via dygraphs plugin apis
-
-	//TODO: logscale bei zweiter chart broken
 	
 	System.out.println("Compiling LaTeX (2/2)...");
 	try {
@@ -243,7 +255,14 @@ public class LaTeX_Compiler {
 	Chart c = (Chart)o.type;
 	cleanupfiles.add("out/web/" + i + ".html");
 	cleanupfiles.add("" + o.name + ".png");
+	copyfiles.add("" + o.name + ".png");
 	try {
+	    if (c.aggregationattributeBig.equals("")) {
+		c.aggregationattributeBig = c.aggregationattribute;
+	    }
+	    if (c.aggregationBig.equals("")) {
+		c.aggregationBig = c.aggregation;
+	    }
 	    Dimension dim = new Dimension(dimOrg.width * c.quality, dimOrg.height * c.quality);
 	    String viewer;
[…]

### 2015-03-31 test.tex
!
fixed bug with xref

@@
 \usepackage{wrapfig}
+\usepackage{ngerman}
 \usepackage[a4paper,vmargin={30mm,40mm},hmargin={28mm,28mm}]{geometry}
@@
 }
-\date{\today{}, Saarbrücken}
+\date{\today{}, Saarbr"ucken}
 
@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
+%\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
 
@@
     \begin{minipage}{\textwidth}
-    \lineChart[width=1.0\textwidth, height=0.45\textwidth, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}\\%
+    ~%\lineChart[width=1.0\textwidth, height=0.45\textwidth, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}\\%
     \caption{Lot of points}
@@
 
-\newpage
+\clearpage
 \begin{wrapfigure}{r}{0.52\textwidth}
 { \centering %
-\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
-{SELECT
-m.cpuarchitecture AS Machine,
-IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
-IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
-c.compiler      AS Compiler,
-c.optlevel      AS [Opt. level],
-l.name          AS [Layout name],
-CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provi- ded at],
-ROUND(v.runtime, 3) 		AS [Query time]
-FROM fact_experiment AS v
-    JOIN dim_machine m ON m.id = v.machineid
-    JOIN dim_dataset d ON d.id = v.datasetid
-    JOIN dim_compiler c ON c.id = v.compilerid
-    JOIN dim_layout l ON l.id = v.layoutid
-    JOIN dim_implementation i ON i.id = v.implementationid
-    JOIN dim_query q ON q.id = v.queryid
-ORDER BY m.id DESC, d.tableschema, q.querytext,
-    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
-{SELECT
-m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
-d.tableschema   AS [Table schema],
-q.querytext     AS [Query],
-c.compiler      AS Compiler,
-c.optlevel      AS [Opt. level],
-l.name          AS [Layout name],
-CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) AS [Chunk size (tuples)],
-IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
-IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
-t.timestamp 	AS [Time of experiment],
-ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
-FROM fact_experiment AS v
-    JOIN dim_machine m ON m.id = v.machineid
-    JOIN dim_dataset d ON d.id = v.datasetid
-    JOIN dim_compiler c ON c.id = v.compilerid
-    JOIN dim_layout l ON l.id = v.layoutid
-    JOIN dim_implementation i ON i.id = v.implementationid
-    JOIN dim_query q ON q.id = v.queryid
-    JOIN dim_time t ON t.id = v.timeid
-ORDER BY m.id DESC, d.tableschema, q.querytext,
-    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+~%\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
+%{SELECT
+%m.cpuarchitecture AS Machine,
+%IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
+%IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
+%c.compiler      AS Compiler,
+%c.optlevel      AS [Opt. level],
+%l.name          AS [Layout name],
+%CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provi- ded at],
+%ROUND(v.runtime, 3) 		AS [Query time]
+%FROM fact_experiment AS v
+%    JOIN dim_machine m ON m.id = v.machineid
+%    JOIN dim_dataset d ON d.id = v.datasetid
+%    JOIN dim_compiler c ON c.id = v.compilerid
+%    JOIN dim_layout l ON l.id = v.layoutid
+%    JOIN dim_implementation i ON i.id = v.implementationid
+%    JOIN dim_query q ON q.id = v.queryid
+%ORDER BY m.id DESC, d.tableschema, q.querytext,
+%    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+%{SELECT
+%m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
+%d.tableschema   AS [Table schema],
+%q.querytext     AS [Query],
+%c.compiler      AS Compiler,
+%c.optlevel      AS [Opt. level],
+%l.name          AS [Layout name],
+%CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) AS [Chunk size (tuples)],
+%IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
+%IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
+%t.timestamp 	AS [Time of experiment],
+%ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
+%FROM fact_experiment AS v
+%    JOIN dim_machine m ON m.id = v.machineid
+%    JOIN dim_dataset d ON d.id = v.datasetid
+%    JOIN dim_compiler c ON c.id = v.compilerid
+%    JOIN dim_layout l ON l.id = v.layoutid
+%    JOIN dim_implementation i ON i.id = v.implementationid
+%    JOIN dim_query q ON q.id = v.queryid
+%    JOIN dim_time t ON t.id = v.timeid
+%ORDER BY m.id DESC, d.tableschema, q.querytext,
+%    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
 \\ }

### 2015-04-08 testdata/chunked_column_layouts/draft.tex
!

\documentclass{sig-alternate}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{enumerate}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
%\usepackage{booktabs} % for better looking tables
\usepackage{wrapfig}
\usepackage{array}
\usepackage{multirow}

% For SQL and AWK code formatting:
\usepackage{listings}
\usepackage{color}
\definecolor{darkgray}{rgb}{0.6,0.6,0.6}
\definecolor{darkgreen}{rgb}{0.3,0.5,0.3}
\lstdefinestyle{SQL}{
  language=SQL,
  showstringspaces=false,
  basicstyle=\rm\ttfamily,
  backgroundcolor=\color{white},
  commentstyle=\ttfamily\itshape,
  keywordstyle=\ttfamily\color{darkgreen},
  morekeywords={FUNCTION, BOOLEAN, RETURNS, DETERMINISTIC, RETURN, REFERENCES, WITH, SEQUENCE, TRUNCATE, START, RULE, TO, DO, IS, INSTEAD, NEW, PARTITION, RANK},
  stringstyle=\ttfamily\color{blue},
  breaklines=true,
  framexleftmargin=0.5em,
  frame=single,
  rulecolor=\color{black},
  rulesepcolor=\color{darkgray},
  xleftmargin=1em,
  xrightmargin=1em,
  columns=fixed
}

\newcommand{\hide}[1]{}
\sloppy

\begin{document}

\title{The Surprising Runtime Fragility Found when Querying a Single Table}
\title{On the Surprising Runtime Fragility Found when Querying a Single Table}
\subtitle{\vspace*{-0.4cm}[Experiments and Analysis Mini-Paper]}

\numberofauthors{1}

\author{
\alignauthor
Endre Palatinus \hspace{0.5cm} Jens Dittrich 
\\\vspace*{0.2cm}
\affaddr{Information Systems Group, Saarland University}\\
\affaddr{http://infosys.cs.uni-saarland.de}\\
}

\date{}

\maketitle

\section{Problem statement}

Given a table with two attributes and a simple query reading those attributes, which data layout (row, column or a suitable PAX-layout) should we choose in order to get the best possible performance? This sounds really simple, yet this paper shows it is not. We will explore the parameter space that has an impact on the performance of such queries including: (1)~the data type used in the schema, (2)~branches in the code, (3)~the CPU architecture, (4)~the compiler, (5)~the optimization level, as well as (6)~compile time vs. runtime layouts. In addition, we will show that a task like this is not easy to measure as we observe considerable variance throughout our measurements which makes it difficult to argue along means over different runs of an experiment. Therefore, we compute confidence intervals for all measurements and exploit this to detect outliers and define classes of methods that we are not allowed to distinguish statistically.
Our results indicate that a carefully or ill-chosen compilation setup can trigger a performance gain or loss of factor~1.6 to factor~28, depending on the setup. We also introduce robustness graphs displaying the impact of non-optimal layouts on runtime performance.
Finally, we give a clear guideline on when to use which method.

%TODO \vspace{-0.3cm}
\section{Data Layouts in Main Memory}
\label{sec:layouts}

The two most common data layouts used in todays database management systems are row and column layout. These are only the two extremes when vertically partitioning a table. In-between these extremes there exists a full spectrum of column-grouped layouts, which under certain settings can beat both of the aforementioned traditional layouts for legacy disk-based row-stores~\cite{VP}. However, for main-memory systems column grouped layouts have not proved to be of much use for OLTP workloads~\cite{grund2010hyrise}, unless the schema is very wide~\cite{JITHybrid}.

Another axis of partitioning a table is horizontal partitioning. This is usually based on the values of a column with low cardinality, e.g.~geographical regions, but this is not necessarily a requirement. Therefore an alternative to laying out all records of a table in column layout is to do this \textit{within} horizontal partitions, so-called chunks of the table. This can be done e.g.~by taking repeatedly $k$~records from the table and laying them out in column layout. We denote the special case when $k=2^N$ by \mbox{ColumnXN}. Row layout is the same as ColumnX0, and column layout is equivalent to ColumnXN, where $2^N$ is larger or equal to the cardinality of the table. The chunks of these layouts are analogous to PAX~pages~\cite{PAX}, however, we can choose any chunk size that is a multiple of the tuple size, while for PAX we are restricted to multiples of the disk's block size. The possible horizontal partitionings of a table having 2 columns and 8 records, using column layout inside the partitions are illustrated in Figure~\ref{fig:ColumnXN}. Here we can see the two extremes: row- and column layout, and chunked column layouts with a chunk-size of 2 (ColumnX1) and 4 (ColumnX2).

\begin{figure}[!h]
\vspace{-5pt}
\center
\includegraphics[width=0.45\textwidth]{pdf/ColumnXNIllustration.pdf}
\vspace{-10pt}
\caption{Horizontal partitionings of a table having 2 columns and 8 records, using column layout inside the partitions, and chunk sizes of powers of 2.}  
\label{fig:ColumnXN}
\vspace{-10pt}
\end{figure}

\section{The six-dimensional Parameter Space of our Experiments}

\noindent {\bf (1) The datatype used in the schema.} Our dataset is a single table with two integer columns, with a total size of 2~GB. Depending on the data type chosen (1-byte, 4-byte, or 8-byte integers) we get the following scenarios:

\begin{table}[!ht]
  \vspace{-5pt}
  \scriptsize
  \centering
  \begin{tabular}{| l | l | r |}
    \hline
    {\bf Label} & {\bf Schema} & {\bf Tuple count} \\ \hline \hline
    char & \texttt{Table\_char (a int1, b int1)} & 1024 * 1024 * 1024 \\ \hline
    int & \texttt{Table\_int \  (a int4, b int4)} & 256 * 1024 * 1024 \\ \hline
    long & \texttt{Table\_long (a int8, b int8)} & 128 * 1024 * 1024 \\ \hline
    \end{tabular}
    \vspace{-5pt}
    \caption{The schemas used in our experiments}
    \label{tab:scenario}
    \vspace{-5pt}
\end{table}

\noindent {\bf (2) The presence of branches in the query.} We use two queries requiring the tuples to be reconstructed for processing as shown in Figure~\ref{fig:queries}. We have chosen Q2 to factor out the possible runtime overhead caused by the branches in Q1 due to using the MIN function. Since Q2 has no branches, the measured query times are not affected by branch-mispredictions. We also tried a branch-free implementation of the min calculation in Q1 which, however, was consistently slower.

\begin{figure}[!ht]
\vspace{5pt}
\begin{lstlisting}[style=SQL]
Q1: SELECT MIN(a+b) FROM T;
Q2: SELECT SUM(a*b) FROM T;
\end{lstlisting}
   \vspace{-10pt}
    \caption{The queries used in the experiments}
    \label{fig:queries}
    \vspace{-10pt}
\end{figure}

\noindent {\bf (3) The CPU architecture.} The performance characteristics of a main-memory database system are influenced the most by the machine's CPU. As there are usually significant changes between the subsequent CPU architectures, we have chosen machines equipped with Intel CPUs of four subsequent architectures, all running Debian 7.8.0 with Linux kernel version 3.2.0-4-amd64 as shown in Table~\ref{tab:machines}.

\begin{table}[!ht]
\vspace{-5pt}
 \scriptsize
  \centering
%  \begin{tabular}{| l | l | l l | }
  \begin{tabular}{| l | l | r <{\hspace{-9pt}} l | }
     \hline
     {\bf CPU} & {\bf Architecture} & {\bf RAM} &  \\ \hline \hline
     Xeon 5150& Nehalem& 16 GB DDR2 & @ 266 MHz \\ \hline
     Xeon X5690& Westmere& 192 GB DDR3 &@ 1066 MHz \\ \hline
     Xeon E5-2407& Sandy Bridge& 48 GB DDR3 & @ 1333 MHz \\ \hline
     Xeon E7-4870 v2& Ivy Bridge& 512 GB DDR3 & @ 1600 MHz \\ \hline
  \end{tabular}
  \vspace{-10pt}
  \caption{The machines used in our experiments}
  \label{tab:machines}
  \vspace{-5pt}
\end{table}

\noindent {\bf (4) The compiler.} In our experiments we have chosen the three most commonly used compilers: clang++
(3.0-6.2), g++ (Debian 4.7.2-5), and icpc (15.0.0).

%\begin{table}[htbp]
%  \centering
%  \begin{tabular}{| l | l | l | }
%    \hline
%   {\bf Compiler} & {\bf Version} \\ \hline \hline
%        clang++ & 3.0-6.2 \\ \hline
%        g++ & Debian 4.7.2-5 \\ \hline
%        icpc & 15.0.0  \\ \hline
%  \end{tabular}
%  \caption{The compilers used in the experiments}
%  \label{tab:compilers}
%\end{table}

\noindent {\bf (5) The optimization level.} We intuitively expect to get higher performance from higher optimization levels, yet there is no guarantee from the compiler's side that this will also hold in practice. Thus, we have decided to evaluate three standard levels: \texttt{-O1}, \texttt{-O2}, and \texttt{-O3}.

\noindent {\bf (6)~Compile time vs.~runtime layouts.} The tables in our dataset are physically stored in a one-dimensional array of integers, using the chunked column layout linearisation order described in Section~\ref{sec:layouts}. Any query fired against this dataset needs to take care of determining the (virtual) address of any attribute value, and possibly reconstructing tuples as well. To do this it is required to know the chunk size, which can either be specified prior to compiling a given query, i.e.~at compile time, or only provided at runtime. In the first case a separate executable is compiled using templates for executing a query adjusted to a specified layout. This means the layout itself is hard-coded to allow for any compiler optimization to take place that might exploit the chunk size. If the chunk size is not known in advance, a single generic executable is generated that takes the chunk size as a runtime parameter when executing the query. This latter approach inherently has a higher query time for relatively small chunk sizes, caused by the CPU-overhead for executing the loops processing chunks that contain too few elements, yielding many short-lived loops.

\section{Methodology}

The most common way of measuring the performance of algorithms, systems, or components in the database community is to report the average runtime out of 3 or 5 runs. Let's look at an example: assume we measured runtimes of a query when executed against two different layouts. Layout~A has an average runtime of 1.75 seconds and Layout~B of 1.82 seconds. In this case we would clearly declare Layout~A as superior to Layout~B.
\begin{wrapfigure}{R}{0.19\textwidth}
\vspace*{-15pt}
\hspace*{-8pt}
\includegraphics[width=.18\textwidth]{pdf/methodology_runtimes.pdf}
\vspace{-15pt}
\caption{Query times for two different layouts, each measured five times}  
\label{fig:methodology_runtimes}
\vspace{-10pt}
\end{wrapfigure}
However, if we take a look at the runtimes of all 5 runs in Figure~\ref{fig:methodology_runtimes}, we can see that Layout~A has a high variance (0.06), whereas the query time for Layout~B is rather stable (it's variance is 0.00075). Most system designers would probably prefer Layout~B, due to its performance being more predictable. This example demonstrates that reporting the average runtime alone is not sufficient for comparing two solutions~\cite{jain1991art}. Therefore at a minimum the variance or standard deviation of the sample should be provided along with the average to get a proper description of the sample.

We should keep in mind that when experimentally comparing multiple systems, we only get a \textit{sample} of their performance metrics which can only be used to \textit{estimate} the populations' performance metrics. Thus, there is always a level of uncertainty in our estimates, which renders the necessity of expressing this uncertainty in some way. One possible way to do this is to use confidence intervals, which express in natural language: ``There is a 95\% chance that the actual average runtime of System~A is between 1.7 and 1.8 seconds.''

\begin{table}[!b]
\includegraphics[width=.478\textwidth]{pdf/problem_bests_table.pdf}
\vspace*{-20pt}
\caption{The best layouts and most efficient ways of implementing a given query on a given table and executed on a given machine.}  
\label{fig:problem_bests_table}
%\vspace{-10pt}
\end{table}

\noindent\textbf{Confidence intervals.}
To create a confidence interval we first have to choose our confidence level, typically 90\%, 95\% or 99\%, denoted by $1 - \alpha$, where $\alpha$ is called the significance level. 
%A higher level of confidence yields a wider confidence interval for the same sample. The use-case determines the suitable confidence level, however, the latter trade-off has to be kept in mind when choosing one.
We require the sample size $n$, the sample mean $\overline{x} $, sample standard deviation $\sigma$, and the significance level $\alpha$. Then the confidence interval is defined as follows: $( \overline{x} - C \times \frac{\sigma}{\sqrt{n}} , \overline{x} + C \times \frac{\sigma}{\sqrt{n}}) $, where $C$ is the so-called confidence coefficient.
The choice of the confidence coefficient is determined by the sample size~\cite{jain1991art}. If we have a large sample ($n \geq 30$), we can use the $1 - \alpha / 2$-quantile of the standard normal distribution for the confidence coefficient: $C = Z_{1-\alpha / 2}$. However, we ran only 5 measurements, thus we have a sample size of $n=5$. Therefore, we should only use the $1 - \alpha / 2$-quantile of the Student's t-distribution with $n-1$ degrees of freedom: $C = t_{[1-\alpha / 2 , n-1]}$. The prerequisite is that the population needs to have a normal distribution, which is a fair assumption for our runtime measurements.
%To illustrate the difference between these two coefficients: for $\alpha = 0.05$, and $n = 5$ we get $Z_{0.975} = 1.96$ and $t_{[0.975, 4]} = 2.776$, thus for a small sample we get a wider confidence interval due to the confidence coefficient alone.
For instance, the 95\% confidence intervals for our example in Figure~\ref{fig:methodology_runtimes} are:  (0.23,~3.27) for Layout~A, and (1.65,~1.99) for Layout~B. This makes Layout~B a safer choice, if predictability is of great importance for the system designer. (See~\cite{jain1991art} for details.)
When looking at the measured query times on Layout~A in Figure~\ref{fig:methodology_runtimes}, we can see that the relatively wide confidence interval for Layout~A is due to the large variance of the sample: the points are scattered  out across the (1.4,~2.0) interval. However, a sample can have a large variance even if most measured values are ``near" to each other, and only a few of them having a higher or lower value than the rest. These latter are called outliers.

\noindent\textbf{Outlier detection.}
An outlier is an element of a sample that does not ``fit" into the sample in some way. It is hard to quantify the criteria for labelling an element as an outlier, and it also depends heavily on the use-case. Therefore, the most common technique used for detecting outliers is plotting the sample on a scatter plot, and visually inspecting the plot by a human.
If we assume, that there is only one outlier in the sample, and it is either the minimum, or the maximum value, than we can use Grubb's test~\cite{grubbs1950sample} to automatically detect outliers. The only problem is that this method tends to identify outliers too often for samples with less than eight elements. This explains why it detected an outlier in 27\% of the cases in our experiments. To counter the error rate of the method we have included an additional condition for labelling an outlier: $\text{margin\_or\_error} / \overline{x} \geq$ 2.5\%, where the margin of error is defined as the radius of the confidence interval. This reduces the detection rate to 3\%, and those elements proved to be outliers after manual inspection.

\noindent\textbf{Choosing the best solution when there is no single best solution.}
Choosing the best solution using the average runtime is easy, we simply take the one with the smallest one. We have also seen that this can be arbitrarily wrong, and that is why confidence intervals provide a better basis of comparison than the sample mean. 
However, comparing confidence intervals is not that straight-forward as comparing scalars. If two intervals are disjoint, they are easily comparable. It they are not disjoint, and the mean of one sample is inside the other sample's confidence interval, they are indistinguishable from each other with the same level of confidence, as that of the intervals. Finally, if they are not disjoint, but their means do not fall into the other sample's interval, an independent two-sample t-test (Welch's t-test~\cite{welch1947generalization}) can decide whether they are distinguishable, and if so, which one is better.

\vspace*{-0.2cm}

\section{Results}

Table~\ref{fig:problem_bests_table} displays our recommendations for choosing a data layout and implementation strategy for each machine, schema and query\footnote{We have noticed that varying the chunk size of chunked column-layouts between $2^{12}$ and the biggest possible one ($2^{30}$ for Table\_char, $2^{28}$ for Table\_int, and $2^{27}$ for Table\_long) does not make a significant difference in the query times, regardless of the query, machine, and compiler. Thus, we have excluded those results from our discussion.
}. Notice that in some cases there are multiple best solutions. Looking at these results the question arises: what influences the choice of best solution? Thus, let us investigate the connection between the elements of the parameter space and the best layout. %For this we have calculated the relative occurrence of a layout as the best solution, aggregated over the values of each parameter (machine, schema, query, compiler, and O-level). Here we have taken special care to weight multiple best solutions of a problem with the reciprocal of their count.

\begin{figure}[!h]
\vspace{-5pt}
\centering
\includegraphics[width=0.48\textwidth]{pdf/problem_bests_plot.pdf}
\vspace{-20pt}
\caption{Best layouts and their query times. Drilled-down along machine, schema, and query.}  
\label{fig:problem_bests_plot}
\vspace{-10pt}
\end{figure}

\begin{figure*}[!ht]
\centering
\includegraphics[width=.97\textwidth]{pdf/fragility.pdf}
\vspace{-10pt}
\caption{Robustness graphs for 24 different experiments displaying the impact of non--optimal layouts on runtime performance. The horizontal axis displays the k-th-best data layout picked (where $k$ is in 1$,\ldots, N=$252; normalized to 100\%). The vertical axis displays the performance overhead of that method over the best one (displayed till at most 250\% overhead; notice that some curves leave their plot).
The bigger the area under the curve, the more likely it is that a decision for a non-optimal data layout will trigger a performance loss. }  
\label{fig:fragility}
\vspace{-10pt}
\end{figure*}

In Figure~\ref{fig:problem_bests_plot} we can see the query times of the best layouts, drilled-down along machine, schema, and query. We can immediately notice the radical difference between Nehalem and the other three CPU architectures. The oldest one, Nehalem, prefers layouts with smaller chunk sizes, i.e.~close to row layout. The three newer ones on the other hand prefer larger chunk sizes, i.e.~close to column layout. For the latter CPUs we can further notice that the best layout for a dataset is often the one, where the following holds: \texttt{2~* field\_size * chunk\_size = 4KB} --- which is when the chunk perfectly fits the memory page: columnX12 for \texttt{char}, columnX10 for \texttt{int} and columnX9 for \texttt{long}.

\begin{table}[!ht]
  \vspace{-10pt}
  \scriptsize
  \centering
  \begin{tabular}{| l || r | r | r | r | r | r |}
    \hline
\multirow{2}{*}{\bf Machine}  &  \multicolumn{2}{c|}{\bf char} &  \multicolumn{2}{c|}{\bf int} & \multicolumn{2}{c|}{\bf long} \\ \cline{2-7}
 & {\bf Q1} & {\bf Q2} & {\bf Q1} & {\bf Q2} & {\bf Q1} & {\bf Q2} \\ \hline \hline
Nehalem & 11.2 & 6.6 & 2.7 & 2.6 & 2.0 & 2.0 \\ \hline
Westmere & 29.6 & 6.7 & 3.7 & 4.7 & 3.7 & 2.8 \\ \hline
Sandy Bridge & 20.6 & 5.2 & 4.3 & 4.5 & 2.9 & 2.8 \\ \hline
Ivy Bridge & 14.4 & 4.3 & 3.2 & 3.7 & 2.6 & 2.3 \\ \hline
    \end{tabular}
    \vspace{-5pt}
    \caption{The quotient of worst and best query times for each experiment. We observe up to a factor 29.6 difference in runtime.}
    \label{tab:fragility}
    \vspace{-5pt}
    \end{table}
So far we have seen the best solutions, but have not talked about the performance of the other ones. In Table~\ref{tab:fragility} we show the ratio of the worst and the best query times, drilled-down along machine, schema, and query. For \texttt{char} we can get factor 4 to factor 29 worse by choosing the wrong layout and/or compiler. At this point it would be interesting to know, how does the best solution's runtime compare to the others solutions' runtimes?

%When looking at the plots in Figure~\ref{fig:fragility} we can note that best solutions for \texttt{char} and Q1 excel in performance, as only 5\% of the solutions have a query time that is at most 50\% worse than the best.

% When looking at the results per machine in Figure~\ref{fig:table_layout_machine}, we can see that our oldest machine, with the Nehalem CPU, prefers layouts with smaller chunk sizes. However, the newer CPUs prefer layouts with larger chunk sizes. We get more insight into the latter phenomena when looking at the connection between the schema and the layout in Figure~\ref{fig:table_layout_schema_all}. Here we can notice that the best layout for a dataset is most often the one, where the following holds: size(field) * chunk\_size = 2KB. Finally, the query doesn't influence too much which layout is the best, as seen in Figure~\ref{fig:table_layout_query_all}.

% It is an interesting question, whether the compiler and O-level have an effect on the performance of the layouts, i.e.~which one is chosen as the best one. TODO: The relative occurrence of a layout as the best solution, aggregated over each compiler and O-level can be seen in Figure~\ref{fig:tables_solution_layout}. We can conclude that the compiler and O-level does not make too much of a difference of a given layout, yet this doesn't mean that all compilers are equally efficient. When looking at the Compiler column in Figure~\ref{fig:problem_bests} we can see that \texttt{icpc} (50\%) and \texttt{g++} (46\%) are the most efficient ones, and \texttt{clang++} (4\%) was rarely providing the best runtime.
\vspace*{-0.2cm}

\section{Robustness Graphs}

In Figure~\ref{fig:problem_bests_plot} we could already observe that the set of best layouts for a particular experiment is not always considerably faster than the second or even the $k$-th-best method. A~database architect may be willing to live with a data layout that is suboptimal for a very specific case, but does not incur too much performance overhead in the general case. To facilitate this decision, we introduce robustness graphs. Given runtime measurements for $N$ different methods ($N=252$ in our case), we depict the overhead of the $k$-th best method over the best method set. The result for 24 different experiments is shown in Figure~\ref{fig:fragility}. The robustness graphs show that for some situations, e.g.~Q1 on \texttt{char} run on Sandy Bridge, one should be more careful in choosing the layout than in others, e.g.~Q2 on \texttt{long} run on Westmere. We are planning to explore these kind of graphs as a general technique in a longer  paper.

\section{Conclusions and Guidelines}

Our guideline for choosing the best layout is as follows:
For servers equipped with a Nehalem CPU it is a safe bet to use row layout, while for machines with the subsequent Westmere, Sandy Bridge, and Ivy Bridge architectures it is just fine to use column layout. For the latter machines we can exploit the schema for some fine-tuning, by creating PAX--blocks with the same size as the virtual memory pages. Having branches in the query is an additional argument for this optimization. The compiler, O-level, and compile time vs. runtime layouts will not change the choice of best layout, but they are to be chosen carefully for the best performance.

We have shown how misleading it can be to choose the best solution along means. Take the case of Q1 on \texttt{long} run on Nehalem, where 20\% of all possible solutions are statistically indistinguishable from best solutions with 95\% confidence.

\noindent {\bf Acknowledgments.} Research supported by BMBF.

% ensure same length columns on last page (might need two sub-sequent latex runs)
\balance

\vspace{-5pt}

\scriptsize
\bibliographystyle{abbrv}
\bibliography{chunkedColumns}

\end{document}

### 2015-04-08 out/web/main.js

@@ -539,7 +539,7 @@ function getChartOptions(json, zoomFactor, values, chart) {
 		point: {
 			show: false
 		},
-		completeScale: zoomFactor
+		completeScale: zoomFactor*1.45
 		/*
 		 * labels : columns, logscale : json.type.I.logScale, animatedZooms :
 		 * true, labelsSeparateLines : true, legend : "always",
@@ -565,11 +565,6 @@ function getChartOptions(json, zoomFactor, values, chart) {
 				+ json.type.I.options);
 	}
 	jQuery.extend(true, options, addOpt);
-	
-	if (options.axis.x.type == "timeseries") {
-		//options.data.xFormat = "%Y/%m/%d";
-        //options.axis.x.tick = { format: '%Y-%m-%d' };
-	}
 	
 	/*
 	 * try { function mergeAintoB(a, b) { for ( var key in a) { if (typeof

### 2015-04-08 src/pdbf/common/LineChart.java

@@ -1,5 +1,5 @@
 package pdbf.common;
 
-public class LineChart extends Chart{
+public class LineChart extends Chart {
 
 }

### 2015-04-08 src/pdbf/common/MultiplotChart.java

@@ -1,0 +1,5 @@
+package pdbf.common;
+
+public class MultiplotChart extends Chart {
+
+}

### 2015-04-13 testdata/chunked_column_layouts/draft.tex
!

@@
 \sloppy
+
+\usepackage{PDBF}
+
+\DBSQLFile{somedb.sql}
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}
+\renewcommand\pdbfScale{1.00}
+\renewcommand\pdbfQuality{1.00}
 
@@
 \centering
-\includegraphics[width=.97\textwidth]{pdf/fragility.pdf}
-\vspace{-10pt}
+\MultiplotChart[width=.97\textwidth, height=.641\textwidth, xunit={Relative size of the set of best solutions}, yunit={Slowdown compared to best solution}, xCount=6, yCount=4, yValues={["runtimeA AS [engine A]", "runtimeB AS [engine B]", "runtimeA AS [engine A]", "runtimeB AS [engine B]"]}, xValues={["date <= '2008/05/10'", "date > '2008/05/10'", "date <= '2008/05/10'", "date > '2008/05/10'", "date <= '2008/05/10'", "date > '2008/05/10'"]}, rightArr={["Nehalem", "Westmere", "Sandy Bridge", "Ivy Bridge"]}, topArr={["Q1, char", "Q2, char", "Q1, int", "Q2, int", "Q1, long", "Q2, long"]}, forceXequal, forceYequal, yFirst, options={"legend": { "show": true }, "axis": { "x": { "tick": { "fit":true } } }}]{SELECT date, ?
+FROM data2 WHERE ?;}
 \caption{Robustness graphs for 24 different experiments displaying the impact of non--optimal layouts on runtime performance.

### 2015-04-13 testdata/chunked_column_layouts/draft.tex
!

@@
 \DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}
+\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}
 \renewcommand\pdbfScale{1.00}
-\renewcommand\pdbfQuality{1.00}
+\renewcommand\pdbfQuality{4.00}
 

### 2015-04-13 src/pdbf/latex/LaTeX_Compiler.java

@@ -77,7 +77,7 @@ public class LaTeX_Compiler {
 
 	// Read JSON
 	GsonBuilder builder = new GsonBuilder();
-	builder.disableHtmlEscaping();
+	builder.disableHtmlEscaping().serializeNulls();
 	builder.registerTypeAdapter(Visualization.class, new VisualizationTypeAdapter());
 	builder.registerTypeAdapter(Alasql.class, new Alasql());
 	builder.registerTypeAdapter(Table.class, new Table());
@@ -231,9 +231,10 @@ public class LaTeX_Compiler {
 		    }
 		    if (cols > 0) {
 			while (rs.next()) {
-			    String[] data = new String[cols];
+			    Object[] data = new Object[cols];
+			    Object o;
 			    for (int i = 1; i <= cols; ++i) {
-				data[i - 1] = rs.getString(i);
+				data[i - 1] = rs.getObject(i);
 			    }
 			    table.data.add(new Data(data));
 			}

### 2015-04-20 beamer_test.tex
!

@@
   \texttt{s9pabend@stud.uni-saarland.de}
-  \\[5pt]
- Christian Goldstein\\
-  \texttt{c.goldstein@goldstein-systemberatung.de}
 }
@@
 
-\DBSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
-\DBSQLFile{somedb.sql}
-\DBSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
+\dbSQLText{CREATE TABLE table1 (column1 INTEGER, column2 DATE, column3 FLOAT);}
+\dbSQLFile{somedb.sql}
+\dbSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtest, dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}%
 
@@
 	{ \centering 
-	\lineChart[width=0.8\textwidth, height=0.6\textwidth, xunit=Money (in €), yunit=Date, options={"title": "Title"}]{SELECT * FROM data1;}%
+	\chart[width=0.8\textwidth, height=0.6\textwidth, chartType=Line, xunit=Money (in €), yunit=Date, options={"title": "Title"}]{SELECT * FROM data1;}%
 	\\}
@@
 	{\centering
-	    \lineChart[width=1.0\textwidth, height=0.5\textwidth, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}%
+	    \chart[width=1.0\textwidth, height=0.5\textwidth, chartType=Line, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}%
 	\\}
@@
     \begin{minipage}{.48\textwidth}
-    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}%
+    \chart[width=\textwidth, height=0.8\textwidth, chartType=Line, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}%
     \caption{Multi-column Line Chart}
@@
     \begin{minipage}{.48\textwidth}
-    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}%
+    \chart[width=\textwidth, height=0.8\textwidth, chartType=Line, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}%
     \caption{Multi-column Bar Chart}
@@
     \begin{minipage}{.48\textwidth}
-    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}, overlap=2]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \chart[width=\textwidth, height=0.8\textwidth, chartType=Bar, xunit=Date, yunit={Runtime [in sec]}, overlap=2]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
     \caption{Multi-column overlapping Bar Chart}
@@
     \begin{minipage}{.48\textwidth}
-    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}, options={"strokeWidth": 0, "drawPoints": true, "pointSize": 3, "highlightCircleSize": 4, "xRangePad": 10}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \chart[width=\textwidth, height=0.8\textwidth, chartType=Line, xunit=Date, yunit={Runtime [in sec]}, options={"strokeWidth": 0, "drawPoints": true, "pointSize": 3, "highlightCircleSize": 4, "xRangePad": 10}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
     \caption{Multi-column Dot Chart}

### 2015-04-20 test.tex
!

--- 2015-03-27 test.tex
@@
 { \centering %
-\lineChart[width=0.4\textwidth, height=0.3\textwidth,  options={"title": "Title"}]{SELECT * FROM data1;}%
+\chart[width=0.4\textwidth, height=0.3\textwidth, chartType=Line, options={"title": "Title"}]{SELECT * FROM data1;}%
 \\ }
@@
      { \centering %
-     \lineChart[width=0.3\textwidth, height=0.4\textwidth, xunit=Date, yunit=Money (in €), logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
+     \chart[width=0.3\textwidth, height=0.4\textwidth, chartType=Line, xunit=Date, yunit=Money (in €), logscale=true]{SELECT * FROM data1 WHERE date >= '2008/05/11';}%
      \\ }
@@
     \begin{minipage}{.48\textwidth}
-    \lineChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}
+    \chart[width=\textwidth, height=0.8\textwidth, chartType=Line, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS engine_A, runtimeB AS engine_B FROM data2;}
     \caption{Multi-column Line Chart}
@@
     \begin{minipage}{.48\textwidth}
-    \barChart[width=\textwidth, height=0.8\textwidth, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+    \chart[width=\textwidth, height=0.8\textwidth, chartType=Bar, xunit=Date, yunit={Runtime [in sec]}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
     \caption{Multi-column Bar Chart}
@@
     \begin{minipage}{\textwidth}
-    \lineChart[width=1.0\textwidth, height=0.45\textwidth, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}\\%
+    \chart[width=1.0\textwidth, height=0.45\textwidth, chartType=Line, xunit=Row Number, yunit={Runtime [in sec]}]{SELECT ROWNUM(), runtime FROM fact_experiment;}\\%
     \caption{Lot of points}
@@
 
-\newpage
+\clearpage
 \begin{wrapfigure}{r}{0.52\textwidth}

### 2015-04-20 testdata/chunked_column_layouts/draft.tex
!

--- 2015-04-08 testdata/chunked_column_layouts/draft.tex
@@
 \sloppy
+
+\usepackage{PDBF}
+
+\dbSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{dim_time, dim_compiler, dim_dataset, dim_implementation, dim_layout, dim_query, dim_machine, fact_experiment}
+\renewcommand\pdbfScale{1.00}
+\renewcommand\pdbfQuality{4.00}
 
@@
 \centering
-\includegraphics[width=.97\textwidth]{pdf/fragility.pdf}
-\vspace{-10pt}
+\multiplotChart[
+width=.97\textwidth, 
+height=.641\textwidth, 
+xunit={Relative size of the set of best solutions}, 
+yunit={Slowdown compared to best solution}, 
+xCount=6, yCount=4, 
+yValues={[ 
+"cpuarchitecture='Nehalem'", 
+"cpuarchitecture='Westmere'", 
+"cpuarchitecture='Sandy Bridge'", 
+"cpuarchitecture='Ivy Bridge'" 
+]}, 
+xValues={[
+"querytext='SELECT MIN(a+b) FROM t' AND tableschema='(char a, char b)'", 
+"querytext='SELECT SUM(a*b) FROM t' AND tableschema='(char a, char b)'", 
+"querytext='SELECT MIN(a+b) FROM t' AND tableschema='(int a, int b)'", 
+"querytext='SELECT SUM(a*b) FROM t' AND tableschema='(int a, int b)'", 
+"querytext='SELECT MIN(a+b) FROM t' AND tableschema='(long a, long b)'", 
+"querytext='SELECT SUM(a*b) FROM t' AND tableschema='(long a, long b)'"
+]}, 
+rightArr={["Nehalem", "Westmere", "Sandy Bridge", "Ivy Bridge"]}, 
+topArr={["Q1, char", "Q2, char", "Q1, int", "Q2, int", "Q1, long", "Q2, long"]}, 
+forceXequal, 
+forceYequal, 
+yFirst,
+chartType=SignaturePlot,
+options={
+	"padding" : { "top" : 6, "right" : 8 }, 
+	"axis": { 
+		"x" : { 
+			"tick" : { "values" : [0, 0.25, 0.5, 0.75, 1], "rotate" : 90 }, 
+			"min" : 0, 
+			"max" : 1, 
+			"padding" : { "left" : 0, "right" : 0 } 
+		}, 
+		"y": { 
+			"tick" : { "values" : [0, 0.5, 1, 1.5, 2, 2.5] }, 
+			"min" : 0, 
+			"max" : 2.5, 
+			"padding" : { "top" : 0, "bottom" : 0 } 
+		} 
+	}, 
+	"legend" : { "show" : false } 
+} 
+]
+{SELECT runtimes 
+  FROM (SELECT machineid, datasetid, queryid, compilerid,   layoutid, implementationid, timeid,   GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid, timeid) AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_query q ON q.id = v.queryid 
+    WHERE ? AND ?;}
 \caption{Robustness graphs for 24 different experiments displaying the impact of non--optimal layouts on runtime performance.

### 2015-04-20 src/pdbf/common/Chart.java

@@ -1,5 +1,5 @@
 package pdbf.common;
 
-public abstract class Chart extends Visualization {
+public class Chart extends Visualization {
 
 }

### 2015-04-20 src/pdbf/common/LineChart.java

@@ -1,5 +1,0 @@
-package pdbf.common;
-
-public class LineChart extends Chart {
-
-}

### 2015-04-23 testdata/chunked_column_layouts/draft.tex
!

@@
 
-\begin{table}[!b]
-\includegraphics[width=.478\textwidth]{pdf/problem_bests_table.pdf}
-\vspace*{-20pt}
-\caption{The best layouts and most efficient ways of implementing a given query on a given table and executed on a given machine.}  
-\label{fig:problem_bests_table}
-%\vspace{-10pt}
-\end{table}
-
 \noindent\textbf{Confidence intervals.}
@@
 
+\begin{table}[!b] 
+\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
+{SELECT
+m.cpuarchitecture AS Machine,
+IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
+IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
+c.compiler      AS Compiler,
+c.optlevel      AS [Opt. level],
+l.name          AS [Layout name],
+CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provided at],
+ROUND(v.runtime, 3) 		AS [Query time]
+FROM fact_experiment AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_compiler c ON c.id = v.compilerid
+    JOIN dim_layout l ON l.id = v.layoutid
+    JOIN dim_implementation i ON i.id = v.implementationid
+    JOIN dim_query q ON q.id = v.queryid
+ORDER BY m.id DESC, d.tableschema, q.querytext,
+    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+{SELECT
+m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
+d.tableschema   AS [Table schema],
+q.querytext     AS [Query],
+c.compiler      AS Compiler,
+c.optlevel      AS [Opt. level],
+l.name          AS [Layout name],
+CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) AS [Chunk size (tuples)],
+IIF(i.templateparam, 'Compile time', 'Runtime') AS [Chunk size provided at],
+IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') AS [Chunk sizes considered],
+t.timestamp 	AS [Time of experiment],
+ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
+FROM fact_experiment AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_compiler c ON c.id = v.compilerid
+    JOIN dim_layout l ON l.id = v.layoutid
+    JOIN dim_implementation i ON i.id = v.implementationid
+    JOIN dim_query q ON q.id = v.queryid
+    JOIN dim_time t ON t.id = v.timeid
+ORDER BY m.id DESC, d.tableschema, q.querytext,
+    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+\vspace*{-5pt}
+\caption{The best layouts and most efficient ways of implementing a given query on a given table and executed on a given machine.}  
+\label{fig:problem_bests_table}
+\vspace{-10pt}
+\end{table}
+
 \noindent\textbf{Choosing the best solution when there is no single best solution.}
@@
 
-\vspace*{-0.2cm}
+\vspace*{-0.15cm}
 

### 2015-04-23 src/pdbf/latex/LaTeX_Compiler.java

@@ -60,7 +60,7 @@ public class LaTeX_Compiler {
 	    System.exit(-1);
 	}
 
-	String latexFolder = latex.getParent();
+	String latexFolder = latex.getAbsoluteFile().getParent();
 
 	ArrayList<String> commands = new ArrayList<String>(Arrays.asList(pathToLaTeXScript));
 	commands.add(latex.getAbsolutePath());
@@ -163,7 +163,7 @@ public class LaTeX_Compiler {
 	    }
 	}
 
-	// copy images to latex file
+	// Copy images to latex file
 	for (String s : copyfiles) {
 	    File source = new File(s);
 	    String target = latexFolder + File.separator + source.getName();

### 2015-04-27 testdata/chunked_column_layouts/draft.tex
!

--- 2015-04-20 testdata/chunked_column_layouts/draft.tex
@@
 
+\begin{table}[!b] 
+\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Bests, aggregationattribute=Query time]
+{SELECT
+m.cpuarchitecture 																													AS Machine,
+IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   			AS [Table schema],
+IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     																				AS [Query],
+c.compiler      																														AS Compiler,
+c.optlevel      																														AS [Opt. level],
+l.name          																														AS [Layout name],
+CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) 																		AS [Chunk size provided at],
+runtimes																																AS [Query time]
+  FROM (SELECT machineid, datasetid, queryid, compilerid, layoutid, implementationid, GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid) AS v
+  JOIN dim_machine m ON m.id = v.machineid 
+  JOIN dim_dataset d ON d.id = v.datasetid 
+  JOIN dim_compiler c ON c.id = v.compilerid 
+  JOIN dim_layout l ON l.id = v.layoutid 
+  JOIN dim_implementation i ON i.id = v.implementationid 
+  JOIN dim_query q ON q.id = v.queryid 
+  ORDER BY m.id DESC, d.tableschema, q.querytext, c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+{SELECT 
+  m.cpu + ' (' + m.cpuarchitecture + ')' 						AS Machine, 
+  d.tableschema 											AS [Table schema], 
+  q.querytext 												AS [Query], 
+  c.compiler 												AS Compiler, 
+  c.optlevel 												AS [Opt. level], 
+  l.name 													AS [Layout name], 
+  CAST(COALESCE(l.chunksize, 'n/a') AS VARCHAR) 			AS [Chunk size (tuples)], 
+  IIF(i.templateparam, 'Compile time', 'Runtime') 				AS [Chunk size provided at], 
+  IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') 			AS [Chunk sizes considered], 
+  runtimes 												AS [Query time] 
+  FROM (SELECT machineid, datasetid, queryid, compilerid, layoutid, implementationid, GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid) AS v
+  JOIN dim_machine m ON m.id = v.machineid 
+  JOIN dim_dataset d ON d.id = v.datasetid 
+  JOIN dim_compiler c ON c.id = v.compilerid 
+  JOIN dim_layout l ON l.id = v.layoutid 
+  JOIN dim_implementation i ON i.id = v.implementationid 
+  JOIN dim_query q ON q.id = v.queryid 
+  ORDER BY m.id DESC, d.tableschema, q.querytext, c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+\vspace*{-5pt}
+\caption{The best layouts and most efficient ways of implementing a given query on a given table and executed on a given machine.}  
+\label{fig:problem_bests_table}
+\vspace{-10pt}
+\end{table}
+
 \noindent\textbf{Choosing the best solution when there is no single best solution.}

### 2015-05-05 testdata/chunked_column_layouts/draft.tex
!

@@
 \begin{table}[!b] 
-\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Bests, aggregationattribute=Query time]
+\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time]
 {SELECT

### 2015-05-12 Documentation.tex
!

@@
 \usepackage{hyperref}
-\usepackage[a4paper,vmargin={28mm,28mm},hmargin={28mm,28mm}]{geometry}
+\usepackage{ngerman}
+\usepackage[a4paper,vmargin={20mm,20mm},hmargin={28mm,28mm}]{geometry}
 
@@
   \texttt{s9pabend@stud.uni-saarland.de}
-  \and
- Christian Goldstein\\
-  \texttt{c.goldstein@goldstein-systemberatung.de}
 }
-\date{\today{}, Saarbrücken}
+\date{\today{}, Saarbr"ucken}
 
@@
 
-TODO: Write small abstract \newpage
+%TODO: Write small abstract
+
+% TODO: picasso diag
+
+%Required packages:
+%\RequirePackage{zref-savepos}
+%\RequirePackage{zref-abspage}
+%\RequirePackage{zref-user}
+%\RequirePackage{xcolor}
+%\RequirePackage{graphicx}
+%\RequirePackage{xstring}
+%\RequirePackage{xparse}
+%\RequirePackage{geometry}
+
+%need to use geometry package
+
+\newpage
 
@@
 Macros: \\[3pt]
-\verb|\lineChart[options]{query}| \\
-\verb|\barChart[options]{query}| \\[8pt]
+\verb|\chart[options][queryForOverlay]{queryForPage}| \\[8pt]
 \noindent Options: \\[3pt]
 \begin{tabular}{ll}
+\textbf{chartType} & \\
+Desc.:	&	Sets the type of the chart. \\
+Argument:	&\verb|line or bar or signatureplot| (more info here)  \\
+Default:&	\verb|line| \\[4pt]
+
 \textbf{width} & \\
@@
 \textbf{options} \\
-Desc.:	&	Sets options that are directly passed to the dygraphs library (\href{http://dygraphs.com/options.html}{$\rightarrow$ Documentation}). \\
+Desc.:	&	Sets options that are directly passed to the c3 chart library (\href{http://c3js.org/reference.html}{$\rightarrow$ Documentation}). \\
 & You need to wrap the JSON-String with \verb|{}| if you want to use \verb|[| or \verb|]| or \verb|,|. \\
@@
 
-\textbf{logscale} \\
-Desc.:	&	If set, the y axis uses log scale. \\
-& Can also be redefined globally (\verb|\pdbfLogscale|) \\
-Argument:&	\verb|{true, false}| \\
-Default:&	\verb|false| \\[4pt]
+%TODO: not yet implemented
+%\textbf{logscale} \\
+%Desc.:	&	If set, the y axis uses log scale. \\
+%& Can also be redefined globally (\verb|\pdbfLogscale|) \\
+%Argument:&	\verb|true or false| \\
+%Default:&	\verb|false| \\[4pt]
 
@@
 
-\textbf{legendpos} \\
-Desc.: 	&	Sets the position of the legend of the chart. \\
-& Can also be redefined globally (\verb|\pdbfLegendpos|).\\
-Argument:&	\verb|{topright, topcenter, topleft, belowleft, belowcenter, belowright}| \\
-Default:&	\verb|topright| \\[4pt]
+%TODO: not yet implemented
+%\textbf{legendpos} \\
+%Desc.: 	&	Sets the position of the legend of the chart. \\
+%& Can also be redefined globally (\verb|\pdbfLegendpos|).\\
+%Argument:&	\verb|{topright, topcenter, topleft, belowleft, belowcenter, belowright}| \\
+%Default:&	\verb|topright| \\[4pt]
 \end{tabular} \\
@@
 \begin{itemize}
-\item F.: The chart is horribly slow when i try to zoom! \\
-A.: Add the following to options: animatedZooms: false 
+\item Q.: The overlay is not on the right position \\
+A.: Most likely you use pages with different sizes in your document. This is currently not supported by PBDF 
+\item Q.: I get \verb|Error: paperwidth value missing! Did you forgot to specify the| \\ \verb|papersize via the geometry package?| when compiling my document \\
+A.: You need to explicitly specify the papersize of your document via the geometry package (e.g. \verb|\usepackage[letterpaper]{geometry}|)
 \end{itemize}

### 2015-05-12 test.tex
!

@@
 \sql{SELECT col2 FROM test WHERE col1=1;}\\[3pt]
-\sqlmult{INSERT INTO test}{INSERT INTO test VALUES (1, 123);}\\
-\sqlmult{VALUES (1, 123);}{INSERT INTO test VALUES (1, 123);}\\[8pt]
+\sql[INSERT INTO test]{INSERT INTO test VALUES (1, 123);}\\
+\sql[VALUES (1, 123);]{INSERT INTO test VALUES (1, 123);}\\[8pt]
 Lorem ipsum \sql{SELECT * FROM test;} dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
@@
 \pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
-{SELECT
-m.cpuarchitecture AS Machine,
-IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
-IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
-c.compiler      AS Compiler,
-c.optlevel      AS [Opt. level],
-l.name          AS [Layout name],
-CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provi- ded at],
-ROUND(v.runtime, 3) 		AS [Query time]
-FROM fact_experiment AS v
-    JOIN dim_machine m ON m.id = v.machineid
-    JOIN dim_dataset d ON d.id = v.datasetid
-    JOIN dim_compiler c ON c.id = v.compilerid
-    JOIN dim_layout l ON l.id = v.layoutid
-    JOIN dim_implementation i ON i.id = v.implementationid
-    JOIN dim_query q ON q.id = v.queryid
-ORDER BY m.id DESC, d.tableschema, q.querytext,
-    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
-{SELECT
+[SELECT
 m.cpu + ' (' +  m.cpuarchitecture + ')' AS Machine,
@@
     JOIN dim_time t ON t.id = v.timeid
+ORDER BY m.id DESC, d.tableschema, q.querytext,
+    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;]
+{SELECT
+m.cpuarchitecture AS Machine,
+IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
+IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
+c.compiler      AS Compiler,
+c.optlevel      AS [Opt. level],
+l.name          AS [Layout name],
+CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provi- ded at],
+ROUND(v.runtime, 3) 		AS [Query time]
+FROM fact_experiment AS v
+    JOIN dim_machine m ON m.id = v.machineid
+    JOIN dim_dataset d ON d.id = v.datasetid
+    JOIN dim_compiler c ON c.id = v.compilerid
+    JOIN dim_layout l ON l.id = v.layoutid
+    JOIN dim_implementation i ON i.id = v.implementationid
+    JOIN dim_query q ON q.id = v.queryid
 ORDER BY m.id DESC, d.tableschema, q.querytext,

### 2015-05-12 testdata/chunked_column_layouts/draft.tex
!

--- 2015-04-23 testdata/chunked_column_layouts/draft.tex
@@
 \begin{table}[!b] 
-\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
-{SELECT
-m.cpuarchitecture AS Machine,
-IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   AS [Table schema],
-IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     AS [Query],
-c.compiler      AS Compiler,
-c.optlevel      AS [Opt. level],
-l.name          AS [Layout name],
-CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) AS [Chunk size provided at],
-ROUND(v.runtime, 3) 		AS [Query time]
-FROM fact_experiment AS v
-    JOIN dim_machine m ON m.id = v.machineid
-    JOIN dim_dataset d ON d.id = v.datasetid
-    JOIN dim_compiler c ON c.id = v.compilerid
-    JOIN dim_layout l ON l.id = v.layoutid
-    JOIN dim_implementation i ON i.id = v.implementationid
-    JOIN dim_query q ON q.id = v.queryid
-ORDER BY m.id DESC, d.tableschema, q.querytext,
-    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
-{SELECT
+\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time]
+[SELECT 
   m.cpu + ' (' + m.cpuarchitecture + ')' 						AS Machine, 
@@
   IIF(i.logscaleparam, 'Powers of 2', 'Multiples of 2') 			AS [Chunk sizes considered], 
-t.timestamp 	AS [Time of experiment],
-ROUND(v.runtime, 3) 		AS [Query time (in Sec)]
-FROM fact_experiment AS v
+  runtimes 												AS [Query time] 
+  FROM (SELECT machineid, datasetid, queryid, compilerid, layoutid, implementationid, GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid) AS v
   JOIN dim_machine m ON m.id = v.machineid 
@@
   JOIN dim_query q ON q.id = v.queryid 
-    JOIN dim_time t ON t.id = v.timeid
-ORDER BY m.id DESC, d.tableschema, q.querytext,
-    c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
+  ORDER BY m.id DESC, d.tableschema, q.querytext, c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;]
+{SELECT
+m.cpuarchitecture 																													AS Machine,
+IIF(d.tableschema = '(long a, long b)', 'long', IIF(d.tableschema = '(char a, char b)', 'char', IIF(d.tableschema = '(int a, int b)', 'int', 'n/a')))   			AS [Table schema],
+IIF(q.querytext = 'SELECT MIN(a+b) FROM t', 'Q1', 'Q2')     																				AS [Query],
+c.compiler      																														AS Compiler,
+c.optlevel      																														AS [Opt. level],
+l.name          																														AS [Layout name],
+CAST(COALESCE(IIF(i.templateparam, 'C', 'R'), 'n/a') AS VARCHAR) 																		AS [Chunk size provided at],
+runtimes																																AS [Query time]
+  FROM (SELECT machineid, datasetid, queryid, compilerid, layoutid, implementationid, GRUBBS_FILTER(ARRAY(runtime)) as runtimes FROM fact_experiment GROUP BY machineid, datasetid, queryid, compilerid, layoutid, implementationid) AS v
+  JOIN dim_machine m ON m.id = v.machineid 
+  JOIN dim_dataset d ON d.id = v.datasetid 
+  JOIN dim_compiler c ON c.id = v.compilerid 
+  JOIN dim_layout l ON l.id = v.layoutid 
+  JOIN dim_implementation i ON i.id = v.implementationid 
+  JOIN dim_query q ON q.id = v.queryid 
+  ORDER BY m.id DESC, d.tableschema, q.querytext, c.compiler, c.optlevel, i.templateparam, i.logscaleparam, l.chunksize;}
 \vspace*{-5pt}

### 2015-05-12 Documentation.tex
!

@@
 
+%problems with utf8 package. for german umlauts use ngerman package instead. 
+%if you really have to use the utf8 package then you cant use special characters inside queries.
+
 \newpage
@@
 Macros: \\[3pt]
-\verb|\chart[options][queryForOverlay]{queryForPage}| \\[8pt]
+\verb|\chart[options][queryForOverlay]{queryForPage}| \\
+Note: if \verb|queryForOverlay| is omitted the \verb|QueryForPage| is used for the overlay\\[8pt]
 \noindent Options: \\[3pt]
@@
 
+%TODO: not yet implemented
+%\textbf{overlap} \\
+%Desc.:	&	unused \\
+%Argument:&	\verb|-| \\
+%Default:&	\verb|-| \\[4pt]
+
+legendpos
+
+includeZero
+
+drawPoints
+
+fillGraph
+
+showRangeSelector
+
 \textbf{quality} \\
-Desc.:	&	Sets the quality for the image version of the chart. \\
+Desc.:	&	Sets the quality for the image version of the chart in the pdf. 1.0 corresponds roughly to 120 pixels per inch \\
 & Can also be redefined globally (\verb|\pdbfQuality|). \\
@@
 \textbf{scale} \\
-Desc.:	&	Sets the scale for the chart. 1.0 corresponds roughly to 12pt in \LaTeX. \\
+Desc.:	&	Scales the size of the chart. 1.0 corresponds roughly to font-size that is currently choosen in \LaTeX. \\
 & Can also be redefined globally (\verb|\pdbfScale|). \\
@@
 
+%TODO: Pivot
+%TODO: MultiplotChart
+
+%\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"aggregation": "\pdbf@aggregation",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"aggregationattribute": "\pdbf@aggregationattribute",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"aggregationBig": "\pdbf@aggregationBig",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"aggregationattributeBig": "\pdbf@aggregationattributeBig",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"xCount": \pdbf@xCount,}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"yCount": \pdbf@yCount,}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"leftArr": "\pdbf@leftArr",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"rightArr": "\pdbf@rightArr",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"topArr": "\pdbf@topArr",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"bottomArr": "\pdbf@bottomArr",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"xValues": "\pdbf@xValues",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"yValues": "\pdbf@yValues",}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"yFirst": \pdbf@yFirst,}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"forceXequal": \pdbf@forceXequal,}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"forceYequal": \pdbf@forceYequal,}% 
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"cols": "\pdbf@cols",}%
+%		\immediate\write\tempfile{\space\space\space\space\space\space\space\space\space\space\space\space"rows": "\pdbf@rows",}%
+
+%TODO: \sql[textForPage]{queryForOverlay}
+%TODO: \dbSQLText{sqlQueryString}
+%TODO: \dbSQLFile{fileWithSqlQueries}
+%TODO: \dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}
+%jdbcConnectionURL consist of ... (e.g. jdbc:postgresql://localhost:5432/postgres)
+%Note: jdbcConnectionURL, user, and password are not stored in the output document
+
 \newpage
@@
 A.: You need to explicitly specify the papersize of your document via the geometry package (e.g. \verb|\usepackage[letterpaper]{geometry}|)
+%TODO: check utf8 error message and put a faq entry here
 \end{itemize}

### 2015-05-14 Documentation.tex
!

--- 2015-03-27 Documentation.tex
@@
 \documentclass[11pt]{scrartcl}
+\usepackage[utf8]{inputenc}
 \usepackage{PDBF}
@@
 \usepackage{hyperref}
-\usepackage[a4paper,vmargin={28mm,28mm},hmargin={28mm,28mm}]{geometry}
+\usepackage[a4paper,vmargin={20mm,40mm},hmargin={20mm,20mm}]{geometry}
 
@@
 
+\def\a{5cm}
+\def\b{10.5cm}
+
+\def\option#1#2#3#4{%
+\noindent \begin{tabular}{|p{\a}|p{\b}|}
+\hline
+\textbf{#1} & \\
+\hline
+Desc.: & #2 \\
+\hline
+Argument: & #3\\
+\hline
+Default:& #4\\
+\hline
+\end{tabular} \\[4pt]%
+}
+
+\dbSQLText{CREATE TABLE test(a int, b int); INSERT INTO test VALUES(1, 10); INSERT INTO test VALUES(2, 100); INSERT INTO test VALUES(3, 120); INSERT INTO test VALUES(4, 100);}
+
 \begin{document}
@@
 
-TODO: Write small abstract \newpage
+\section{Abstract}
+TODO ...
 
-\section{Chart options}
-Macros: \\[3pt]
-\verb|\lineChart[options]{query}| \\
-\verb|\barChart[options]{query}| \\[8pt]
-\noindent Options: \\[3pt]
-\begin{tabular}{ll}
-\textbf{width} & \\
-Desc.:	&	Sets the width of the chart. \\
-Argument:	&LaTeX length \\
-Default:&	no default value. value must always be set! \\[4pt]
-
-\textbf{height} \\
-Desc.:	&	Sets the height of the chart. \\
-Argument: &	LaTeX length \\
-Default:&	no default value. value must always be set! \\[4pt]
-
-\textbf{xunit} \\
-Desc.:	&	Sets the a label for the x axis. \\
-Argument: &	String \\
-Default:& 	\verb|""| (which means hide) \\[4pt]
-
-\textbf{yunit} \\
-Desc.:	&	Sets the a label for the y axis. \\
-Argument: 	&String \\
-Default:& 	\verb|""| (which means hide) \\[4pt]
-
-\textbf{options} \\
-Desc.:	&	Sets options that are directly passed to the dygraphs library (\href{http://dygraphs.com/options.html}{$\rightarrow$ Documentation}). \\
-& You need to wrap the JSON-String with \verb|{}| if you want to use \verb|[| or \verb|]| or \verb|,|. \\
-Argument: 	&JSON-String \\
-Default:&	\verb|{}| (which means empty object) \\[4pt]
-
-\textbf{logscale} \\
-Desc.:	&	If set, the y axis uses log scale. \\
-& Can also be redefined globally (\verb|\pdbfLogscale|) \\
-Argument:&	\verb|{true, false}| \\
-Default:&	\verb|false| \\[4pt]
-
-\textbf{quality} \\
-Desc.:	&	Sets the quality for the image version of the chart. \\
-& Can also be redefined globally (\verb|\pdbfQuality|). \\
-Argument:	&Number $>$ 0 \\
-Default:&	\verb|1.0| \\[4pt]
-
-\textbf{scale} \\
-Desc.:	&	Sets the scale for the chart. 1.0 corresponds roughly to 12pt in \LaTeX. \\
-& Can also be redefined globally (\verb|\pdbfScale|). \\
-Argument:	&Number $>$ 0 \\
-Default:&	\verb|1.0| \\[4pt]
-
-\textbf{legendpos} \\
-Desc.: 	&	Sets the position of the legend of the chart. \\
-& Can also be redefined globally (\verb|\pdbfLegendpos|).\\
-Argument:&	\verb|{topright, topcenter, topleft, belowleft, belowcenter, belowright}| \\
-Default:&	\verb|topright| \\[4pt]
-\end{tabular} \\
+\section{Requirements}
+Required packages:\\
+zref-savepos\\
+zref-abspage\\
+zref-user\\
+xcolor\\
+graphicx\\
+xstring\\
+xparse\\
+geometry\\
+need to use geometry package\\
+problems with utf8 package. for german umlauts use ngerman package instead. \\
+if you really have to use the utf8 package then you cant use special characters inside queries.\\
+All options in all modes because of switcher.\\
 
 \newpage
+\section{Chart}
+\begin{figure}[h!]%
+\hspace{-218pt}Examples:\\
+    \centering
+    \begin{minipage}{.48\textwidth}
+    \chart[width=\textwidth, height=0.7\textwidth, chartType=line]{SELECT * FROM test;}
+    \caption{Line Chart}
+    \end{minipage}
+    \hspace{11pt}
+    \begin{minipage}{.48\textwidth}
+    \chart[width=\textwidth, height=0.7\textwidth, chartType=bar]{SELECT * FROM test;}
+    \caption{Bar Chart}
+    \end{minipage} \\[8pt]
+
+    \begin{minipage}{1.0\textwidth}
+    \chart[width=\textwidth, height=0.36\textwidth, chartType=signatureplot]{SELECT b FROM test;}
+    \caption{Signatureplot Chart}
+    \end{minipage}
+%    \begin{minipage}{.48\textwidth}
+%    \chart[width=\textwidth, height=0.8\textwidth, chartType=signatureplot]{SELECT * FROM test;}
+%    \caption{Signatureplot Chart}
+%    \end{minipage}
+%    \hspace{11pt}
+%    \begin{minipage}{.48\textwidth}
+%    \chart[width=\textwidth, height=0.8\textwidth, chartType=Line, xunit=Date, yunit={Runtime [in sec]}, options={"strokeWidth": 0, "drawPoints": true, "pointSize": 3, "highlightCircleSize": 4, "xRangePad": 10}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
+%    \caption{Multi-column Dot Chart}
+%    \end{minipage}
+\end{figure}
+
+\noindent Macros: \\[3pt]
+\verb|\chart[options][queryForOverlay]{queryForPage}| \\
+Note: if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\[8pt]
+\noindent Options:\\[3pt]
+%
+Common options\\[4pt]
+%
+\option
+{width}
+{Sets the width of the chart.}
+{\LaTeX\relax length}
+{No default value. Value must always be set!}
+
+\option
+{height}
+{Sets the height of the chart.}
+{\LaTeX\relax length}
+{No default value. Value must always be set!}
+
+\option
+{quality}
+{Sets the quality for the image version of the chart in the pdf. 1.0 corresponds roughly to 120 pixels per inch. Can also be redefined globally (\texttt{pdbfQuality}).}
+{Number $>$ 0}
+{\texttt{1.0}}
+
+\option
+{scale}
+{Scales the size of the chart. 1.0 corresponds roughly to font-size that is currently choosen in \LaTeX. Can also be redefined globally (\texttt{pdbfScale}).}
+{Number $>$ 0}
+{\texttt{1.0}}
+
+\option
+{chartType}
+{Sets the type of the chart.}
+{\texttt{line} or \texttt{bar} or \texttt{signatureplot}}
+{\texttt{line}}
+
+\option
+{xunit}
+{Sets the a label for the x-axis.}
+{String}
+{\texttt{""} (which means hide)}
+
+\option
+{yunit}
+{Sets the a label for the y-axis.}
+{String}
+{\texttt{""} (which means hide)}
+
+\option
+{options}
+{Sets options that are directly passed to the c3 chart library (\href{http://c3js.org/reference.html}{$\rightarrow$ Documentation}).
+You need to wrap the JSON-String with \texttt{\{\}} if you want to use \texttt{[} or \texttt{]} or \texttt{,}.}
+{JSON-String}
+{\texttt{\{\}} (which means empty object)}
+
+\option %TODO: not yet implemented
+{logscale}
+{If set, the y axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option %TODO: not yet implemented
+{overlap}
+{Unused.}
+{\texttt{TODO: ...}}
+{\texttt{TODO: ...}}
+
+\option %TODO: not yet implemented
+{legendpos}
+{Sets the position of the legend of the chart.}
+{\texttt{TODO: ...}}
+{\texttt{TODO: ...}}
+
+\option %TODO: not yet implemented
+{includeZero}
+{Set the minimum of the range of the y-axis to zero.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option %TODO: not yet implemented
+{drawPoints}
+{Whether to show each point in line.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option %TODO: not yet implemented
+{fillGraph}
+{Whether to fill the area below the graph.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option %TODO: not yet implemented
+{showRangeSelector}
+{Whether to show a range selector for the x-axis below the chart.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\newpage
+\section{Multiplot Chart}
+\begin{figure}[h!]%
+    Example:\\
+    \multiplotChart[width=\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}
+    \caption{Multiplot Line Chart}
+\end{figure}
+
+\noindent Macros: \\[3pt]
+\verb|\multiplotChart[options][queryForOverlay]{queryForPage}| \\
+Note: If \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\
+Note: Both \verb|queryForOverlay| and \verb|queryForPage| must contain exactly two occurences of the ? character. These are later replaced with values from the xValues/yValues option.\\[8pt]
+\noindent Options: \\[3pt]
+%
+Note: For multiplot charts all options of chart are also valid options.\\[4pt]
+%
+\option
+{xCount}
+{Sets number of columns.}
+{Number $>$ 0}
+{No default value. Value must always be set!}
+
+\option
+{yCount}
+{Sets number of rows.}
+{Number $>$ 0}
+{No default value. Value must always be set!}
+
+\option
+{leftArr}
+{Sets the labels for the left side.}
+{Either a JavaScript array of Strings where each string corresponds to exactly one row (e.g \{"row1", "row2"\}) or a JavaScript array of Objects with a c property which corresponds to the number of rows this text should span and a text property which corresponds to the text-string.}
+{If this option is not set, then the value of the xunit option is used spanning over the whole site.}
+
+\option
+{rightArr}
+{Sets the labels for the right side.}
+{Either a JavaScript array of Strings where each string corresponds to exactly one row (e.g \{"row1", "row2"\}) or a JavaScript array of Objects with a c property which corresponds to the number of rows this text should span and a text property which corresponds to the text-string.}
+{If this option is not set, then the value that is used for the query in this row is used.}
+
+\option
+{bottomArr}
+{Sets the labels for the bottom side.}
+{Either a JavaScript array of Strings where each string corresponds to exactly one column (e.g \{"column1", "column2"\}) or a JavaScript array of Objects with a c property which corresponds to the number of columns this text should span and a text property which corresponds to the text-string.}
+{If this option is not set, then the value of the yunit option is used spanning over the whole site.}
+
+\option
+{topArr}
+{Sets the labels for the tod side.}
+{Either a JavaScript array of Strings where each string corresponds to exactly one column (e.g \{"column1", "column2"\}) or a JavaScript array of Objects with a c property which corresponds to the number of columns this text should span and a text property which corresponds to the text-string.}
+{If this option is not set, then the value that is used for the query in this column is used.}
+
+\option
+{xValues}
+{The values that replace the first ? character in the query. If yFirst is set, they replace the second ? character in the query instead.}
+{A JavaScript array that has as much entries as the xCount options value}
+{No default value. Value must always be set!}
+
+\option
+{yValues}
+{The values that replace the second ? character in the query. If yFirst is set, they replace the first ? character in the query instead.}
+{A JavaScript array that has as much entries as the yCount options value}
+{No default value. Value must always be set!}
+
+\option
+{yFirst}
+{If this option is set, then first ?
+character is replaced with values from yValues option and second ?
+character is replaced with values from xValues.
+If this option is not set, then first ?
+character is replaced with values from xValues option and second ?
+character is replaced with values from yValues.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option
+{forceXequal}
+{If this option is set, then all columns have the same x-axis range as the uppermost chart. If this option is not set, then all charts have individual x-axis ranges.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option
+{forceYequal}
+{If this option is set, then all rows have the same y-axis range as the leftmost chart. If this option is not set, then all charts have individual y-axis ranges.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\newpage
+\section{Pivot Table}
+\begin{figure}[h!]%
+    Example:\\
+    \pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}
+    \caption{Pivot Table}
+\end{figure}
+
+\noindent Macros: \\[3pt]
+\verb|\pivotTable[options][queryForOverlay]{queryForPage}| \\
+Note: if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\[8pt]
+\noindent Options: \\[3pt]
+%
+\option
+{width}
+{Sets the width of the chart.}
+{\LaTeX\relax length}
+{No default value. Value must always be set!}
+
+\option
+{height}
+{Sets the height of the chart.}
+{\LaTeX\relax length}
+{No default value. Value must always be set!}
+
+\option
+{quality}
+{Sets the quality for the image version of the chart in the pdf. 1.0 corresponds roughly to 120 pixels per inch. Can also be redefined globally (\texttt{pdbfQuality}).}
+{Number $>$ 0}
+{\texttt{1.0}}
+
+\option
+{scale}
+{Scales the size of the chart. 1.0 corresponds roughly to font-size that is currently choosen in \LaTeX. Can also be redefined globally (\texttt{pdbfScale}).}
+{Number $>$ 0}
+{\texttt{1.0}}
+
+\option
+{aggregation}
+{Sets the aggregation function for the page.}
+{\texttt{Count} or \texttt{Count Unique Values} or \texttt{List Unique Values} or \texttt{Sum} or \texttt{Integer Sum} or \texttt{Average} or \texttt{Minimum} or \texttt{Maximum} or \texttt{Sum over Sum} or \texttt{80\% Upper Bound} or \texttt{80\% Lower Bound} or \texttt{Sum as Fraction of Total} or \texttt{Sum as Fraction of Rows} or \texttt{Sum as Fraction of Columns} or \texttt{Count as Fraction of Total} or \texttt{Count as Fraction of Rows} or \texttt{Count as Fraction of Columns}}
+{\texttt{Minimum}}
+
+\option
+{aggregationBig}
+{Sets the aggregation function for the overlay.}
+{\texttt{Count} or \texttt{Count Unique Values} or \texttt{List Unique Values} or \texttt{Sum} or \texttt{Integer Sum} or \texttt{Average} or \texttt{Minimum} or \texttt{Maximum} or \texttt{Sum over Sum} or \texttt{80\% Upper Bound} or \texttt{80\% Lower Bound} or \texttt{Sum as Fraction of Total} or \texttt{Sum as Fraction of Rows} or \texttt{Sum as Fraction of Columns} or \texttt{Count as Fraction of Total} or \texttt{Count as Fraction of Rows} or \texttt{Count as Fraction of Columns}}
+{If this option is not set, then the value of the aggregation option is used}
+
+\option
+{aggregationattribute}
+{Sets the attribute for the aggregation function in the page.}
+{The name of an attribute that is present in the result of the sql query for the page}
+{No default value. Value must always be set!}
+
+\option
+{aggregationattributeBig}
+{Sets the attribute for the aggregation function in the overlay.}
+{The name of an attribute that is present in the result of the sql query for the overlay}
+{If this option is not set, then the value of the aggregationattribute option is used}
+
+\option
+{cols}
+{The attributes for the columns in the pivot table.}
+{JavaScript array of strings (e.g. ["a", "b", "c"]). The name of attributes have to be present in the result of the sql query for the overlay and page}
+{\texttt{[]} (which means empty array)}
+
+\option
+{rows}
+{The attributes for the rows in the pivot table.}
+{JavaScript array of strings (e.g. ["a", "b", "c"]). The name of attributes have to be present in the result of the sql query for the overlay and page}
+{\texttt{[]} (which means empty array)}
+
+\newpage
+\section{Sql}
+Example:\\[3pt]
+\sql{SELECT * FROM test;}\\[3pt]
+%
+\noindent Macros: \\[3pt]
+\verb|\sql[textForPage]{queryForOverlay}| \\
+Note: if \verb|textForPage| is omitted the \verb|queryForOverlay| is used as text for the page.\\[8pt]
+
+\section{Sql data sources}
+\noindent Macros: \\[3pt]
+\verb|\dbSQLText{sqlQueryString}| \\
+\verb|\dbSQLFile{fileWithSqlQueries}| \\
+\verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
+Note: jdbcConnectionURL consist of jdbc followed by the name of the dbms (currently only postgresql and mysql are supported) followed by the url of the dbms followed by the database name (e.g.
+jdbc:postgresql://localhost:5432/postgres).\\
+Note: jdbcConnectionURL, user, and password are not stored in the output document.\\[8pt]
+
 \section{F.A.Q.}
 \begin{itemize}
-\item F.: The chart is horribly slow when i try to zoom! \\
-A.: Add the following to options: animatedZooms: false 
+\item Q.: The overlay is not on the right position.\\
+A.: Most likely you use pages with different sizes in your document. This is currently not supported by PBDF 
+\item Q.: I get \verb|Error: paperwidth value missing! Did you forgot to specify the| \\ \verb|papersize via the geometry package?| when compiling my document.\\
+A.: You need to explicitly specify the papersize of your document via the geometry package (e.g. \verb|\usepackage[letterpaper]{geometry}|)
+\item Q.: I get \verb|\unhbox \voidb@x \bgroup \let \unhbox \voidb@x \setbox \@tempboxa \hbox| \verb|{u\global \mathchardef \accent@spacefactor \spacefactor }\accent 127 u\egroup| \newline \verb|\spacefactor \accent@spacefactor| or a similar string in a error message.\\
+A.: Chances are high that you use the \verb|\usepackage[utf8]{inputenc}| package or a similar package.
+These are not fully compatible with PDBF.
+If you really have to use the utf8 package then you cant use special characters inside PDBF commands.
 \end{itemize}

### 2015-05-14 test.tex
!

@@
 { \centering %
-\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
+\pivotTable[width=0.5\textwidth, height=0.8\textwidth, scale=0.5, rows={["Machine", "Table schema", "Query"]}, aggregation=Minimum, aggregationattribute=Query time, aggregationBig=Minimum, aggregationattributeBig=Query time (in Sec)]
 [SELECT

### 2015-05-14 testdata/chunked_column_layouts/draft.tex
!

@@
 \begin{table}[!b] 
-\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={"Machine", "Table schema", "Query"}, aggregation=Minimum, aggregationattribute=Query time]
+\pivotTable[width=0.478\textwidth, height=0.7478\textwidth, scale=0.666, rows={["Machine", "Table schema", "Query"]}, aggregation=Minimum, aggregationattribute=Query time]
 [SELECT 

### 2015-05-14 out/web/d3.js

@@ -7505,7 +7505,7 @@
     return d3_scale_nice(domain, d3_scale_niceStep(d3_scale_linearTickRange(domain, m)[2]));
   }
   function d3_scale_linearTickRange(domain, m) {
-    if (m == null) m = 10;
+    if (m == null) m = 8;
     var extent = d3_scaleExtent(domain), span = extent[1] - extent[0], step = Math.pow(10, Math.floor(Math.log(span / m) / Math.LN10)), err = m / span * step;
     if (err <= .15) step *= 10; else if (err <= .35) step *= 5; else if (err <= .75) step *= 2;
     extent[0] = Math.ceil(extent[0] / step) * step;

### 2015-05-18 testdata/chunked_column_layouts/draft.tex
!

@@
 \centering
-\includegraphics[width=0.48\textwidth]{pdf/problem_bests_plot.pdf}
+\multiplot[
+width=0.48\textwidth, 
+height=0.3329\textwidth, 
+charttype=scatter,
+xunit={Best layouts}, 
+yunit={Average query time [seconds]}, 
+xCount=1, yCount=4, 
+yValues={[ 
+"cpuarchitecture='Nehalem'", 
+"cpuarchitecture='Westmere'", 
+"cpuarchitecture='Sandy Bridge'", 
+"cpuarchitecture='Ivy Bridge'" 
+]}, 
+xValues={[
+"querytext='SELECT MIN(a+b) FROM t' AND tableschema='(char a, char b)'"%, 
+%"querytext='SELECT SUM(a*b) FROM t' AND tableschema='(char a, char b)'", 
+%"querytext='SELECT MIN(a+b) FROM t' AND tableschema='(int a, int b)'", 
+%"querytext='SELECT SUM(a*b) FROM t' AND tableschema='(int a, int b)'", 
+%"querytext='SELECT MIN(a+b) FROM t' AND tableschema='(long a, long b)'", 
+%"querytext='SELECT SUM(a*b) FROM t' AND tableschema='(long a, long b)'"
+]}, 
+rightArr={["Nehalem", "Westmere", "Sandy Bridge", "Ivy Bridge"]}, 
+topArr={[""]}, 
+forceXequal, 
+forceYequal, 
+yFirst,
+chartType=SignaturePlot,
+options={
+	"padding" : { "top" : 6, "right" : 8 }, 
+	"axis": { 
+%		"x" : { 
+%			"tick" : { "values" : [0, 0.25, 0.5, 0.75, 1], "rotate" : 90 }, 
+%			"min" : 0, 
+%			"max" : 1, 
+%			"padding" : { "left" : 0, "right" : 0 } 
+%		}, 
+		"y": { 
+			"tick" : { "values" : [0, 0.2, 0.4, 0.6, 0.8, 1.0]}, 
+			"min" : 0, 
+			"max" : 1, 
+			"padding" : { "top" : 0, "bottom" : 0 } 
+		} 
+	}, 
+	"legend" : { "show" : false } 
+} 
+]]{}
 \vspace{-20pt}

### 2015-05-18 Documentation.tex
!

@@
 
+\section{SQL queries}
+attributes with spaces sourrounded by square bracets. E.g. SELECT a AS [This is a test] FROM test;\\
+TODO: write about special sql functions. link to alasql.\\
+\verb|GRUBBS_FILTER(arr, alpha)|: \\
+\verb|MEAN(arr)|: \\
+\verb|STDDEV_SAMP(arr)|: \\
+\verb|MARGIN_OF_ERROR(arr, alpha)|: \\
+\verb|CONF_INT(arr)|: \\
+\verb|T_TEST(arr1, arr2, alpha)|: \\
+\verb|WELCH_TEST(arr1, arr2, alpha)|: \\
+
 \newpage
@@
 \verb|\chart[options][queryForOverlay]{queryForPage}| \\
-Note: if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\[8pt]
+Note: if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay.\\
+Note: The first column of the query result is used for the x-Axis, all other columns are used for y-Axis.\\[8pt] 
 \noindent Options:\\[3pt]

### 2015-05-18 out/web/science.v1.js

@@ -1,1738 +1,0 @@
-(function(exports){
-(function(exports){
-science = {version: "1.9.1"}; // semver
-science.ascending = function(a, b) {
-  return a - b;
-};
-// Euler's constant.
-science.EULER = .5772156649015329;
-// Compute exp(x) - 1 accurately for small x.
-science.expm1 = function(x) {
-  return (x < 1e-5 && x > -1e-5) ? x + .5 * x * x : Math.exp(x) - 1;
-};
-science.functor = function(v) {
-  return typeof v === "function" ? v : function() { return v; };
-};
-// Based on:
-// http://www.johndcook.com/blog/2010/06/02/whats-so-hard-about-finding-a-hypotenuse/
-science.hypot = function(x, y) {
-  x = Math.abs(x);
-  y = Math.abs(y);
-  var max,
-      min;
-  if (x > y) { max = x; min = y; }
-  else       { max = y; min = x; }
-  var r = min / max;
-  return max * Math.sqrt(1 + r * r);
-};
-science.quadratic = function() {
-  var complex = false;
-
-  function quadratic(a, b, c) {
-    var d = b * b - 4 * a * c;
-    if (d > 0) {
-      d = Math.sqrt(d) / (2 * a);
-      return complex
-        ? [{r: -b - d, i: 0}, {r: -b + d, i: 0}]
-        : [-b - d, -b + d];
-    } else if (d === 0) {
-      d = -b / (2 * a);
-      return complex ? [{r: d, i: 0}] : [d];
-    } else {
-      if (complex) {
-        d = Math.sqrt(-d) / (2 * a);
-        return [
-          {r: -b, i: -d},
-          {r: -b, i: d}
-        ];
-      }
-      return [];
-    }
-  }
-
-  quadratic.complex = function(x) {
-    if (!arguments.length) return complex;
-    complex = x;
-    return quadratic;
-  };
-
-  return quadratic;
-};
-// Constructs a multi-dimensional array filled with zeroes.
-science.zeroes = function(n) {
-  var i = -1,
-      a = [];
-  if (arguments.length === 1)
-    while (++i < n)
-      a[i] = 0;
-  else
-    while (++i < n)
-      a[i] = science.zeroes.apply(
-        this, Array.prototype.slice.call(arguments, 1));
-  return a;
-};
-})(this);
[…]

### 2015-05-18 src/pdbf/html/HTML_Compiler.java

@@ -10,60 +10,54 @@ import pdbf.common.Tools;
 
 public class HTML_Compiler {
 
-    public static DecimalFormat df = new DecimalFormat( "0000000000" );
-    
+    public static DecimalFormat df = new DecimalFormat("0000000000");
+
     public static void main(String[] args) {
 	System.out.println("Compiling HTML...");
-	
+
 	String a = new File(args[0]).getName();
-	String filename = a.substring(0, a.length()-4);
-	String basename = args[0].substring(0, args[0].length()-4);
+	String filename = a.substring(0, a.length() - 4);
+	String basename = args[0].substring(0, args[0].length() - 4);
 	String pdfname = filename + ".pdf";
 	String outfile = basename + ".html";
-	
+
 	try {
 	    String viewer;
 	    String viewerHEAD = FileUtils.readFileToString(new File("out/web/templateHEADalasql.html"), Tools.utf8);
 	    String viewerTAIL = FileUtils.readFileToString(new File("out/web/templateTAILalasql.html"), Tools.utf8);
+	    if (CompleteRun_HTML.includeRes) {
+		viewerTAIL += "</script>";
+	    } else {
+		viewerTAIL += "</script> <link rel=\"stylesheet\" href=\"viewer.css\"/>" + "<link rel=\"stylesheet\" href=\"pivot.css\"/>" + "<link rel=\"stylesheet\" href=\"codemirror.css\"/>" + "<link rel=\"stylesheet\" href=\"jquery.dataTables.css\"/>" + "<link rel=\"stylesheet\" href=\"c3.css\"/>" + "" + "<script src=\"base64.js\"></script>" + "<script src=\"d3.js\"></script>" + "<script src=\"alasql.js\"></script>" + "<script src=\"codemirror-compressed.js\"></script>"
[…]

### 2015-05-20 Documentation.tex
cleanup

--- 2015-05-14 Documentation.tex
@@
 \section{Abstract}
-TODO ...
+PDBF documents are a hybrid format. They are a valid PDF and a valid HTML page at the same time. 
+If you change the file extension to PDF and open it with an PDF viewer you can see the static part of the document.
+If you change the file extension to HTML and open it with a Browser (currently Chrome/Firefox/IE 10 supported) you can see the dynamic part of the document.
+For example if the static version contains an image of a chart which displays some data, then the dynamic version contains the actual raw data used to render the chart and renders the chart when opening the document.
+The advantage is that you can open the chart in an overlay view by clicking on the enlarge symbol at the upper left corner and then start to change parameters of the chart.
+For example you can remove filtering functions that are applied to the raw data or change confidence levels to other values an see the results of these changes directly in the chart.
+\\[4pt]
+%
+\noindent PDBF files are created from LaTeX code and data in SQL format.
+The raw data can either be a SQL statements string, a file with SQL statements, or in a database (currently PostgreSQL/MySQL/MariaDB supported).
+In the LaTeX code one can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query.\\[4pt]
+%
+\noindent PDBF toolkit is written in Java and LaTeX an can therefore be run on Windows/Mac/Linux. PDBF documents are also platform independent and run on any OS with a browser/PDF viewer.\\[4pt]
+%
+\noindent This toolkit is licensed unter the MIT License.\\
+
+\noindent Thanks to the authors of:\\
+phantomJS (\url{https://github.com/ariya/phantomjs})\\
+Apache Commons IO (\url{http://commons.apache.org/proper/commons-io/})\\
+Apache Commons Codec (\url{http://commons.apache.org/proper/commons-codec/})\\
+google-gson (\url{https://github.com/google/gson})\\
+postgreSQL JDBC Driver (\url{https://jdbc.postgresql.org/})\\
+MariaDB JDBC Driver (\url{https://mariadb.com/kb/en/mariadb/about-the-mariadb-java-client/})\\
+AlaSQL (\url{https://github.com/agershun/alasql})\\
+C3 (\url{https://github.com/masayuki0812/c3})\\
+D3 (\url{https://github.com/mbostock/d3})\\
+Codemirror (\url{https://github.com/codemirror/codemirror})\\
+google-diff-match-patch (\url{https://code.google.com/p/google-diff-match-patch/})\\
+explorercanvas (\url{https://code.google.com/p/explorercanvas/})\\
+DataTables (\url{https://github.com/DataTables/DataTables})\\
+jQuery (\url{https://github.com/jquery/jquery})\\
+jQuery-UI (\url{https://github.com/jquery/jquery-ui})\\
+jStat (\url{https://github.com/jstat/jstat})\\
+pivottable (\url{https://github.com/nicolaskruchten/pivottable})\\
+PDF.js (\url{https://github.com/mozilla/pdf.js})\\
+\newpage
+\tableofcontents
+\newpage
 
 \section{Requirements}
-Required packages:\\
+These \LaTeX\relax packages are required for PDBF documents:\\
 zref-savepos\\
@@
 geometry\\
-need to use geometry package\\
-problems with utf8 package. for german umlauts use ngerman package instead. \\
-if you really have to use the utf8 package then you cant use special characters inside queries.\\
-All options in all modes because of switcher.\\
+
+\noindent Furthermore it is needed that your \LaTeX\relax document uses the geometry package to specify the page size.\\
+
+\noindent \textbf{Warning:} There are problems with inputenc UTF8 package. If you really have to use the inputenc UTF8 package then you cant use non-ASCII characters inside queries.
+
+\section{SQL queries for PDBF elements}
+Attributes with spaces sourrounded by square bracets. E.g. SELECT a AS [This is a test] FROM test;\\
+
+\noindent TODO: write about special sql functions. AlaSQL (\url{https://github.com/agershun/alasql}).\\
+\verb|GRUBBS_FILTER(arr, alpha)|: \\
+\verb|MEAN(arr)|: \\
+\verb|STDDEV_SAMP(arr)|: \\
+\verb|MARGIN_OF_ERROR(arr, alpha)|: \\
+\verb|CONF_INT(arr)|: \\
+\verb|T_TEST(arr1, arr2, alpha)|: \\
+\verb|WELCH_TEST(arr1, arr2, alpha)|:
+
+\section{Sql data sources}
+\noindent Macros: \\[3pt]
+\verb|\dbSQLText{sqlQueryString}| \\
+\verb|\dbSQLFile{fileWithSqlQueries}| \\
+\verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
+\textbf{Note:} jdbcConnectionURL consist of jdbc followed by the name of the dbms (currently only postgresql and mysql are supported) followed by the url of the dbms followed by the database name (e.g.
+jdbc:postgresql://localhost:5432/postgres).\\
+\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output document.
 
 \newpage
-\section{Chart}
+\section{PDBF elements}
+Currently there are four different PDBF elements: charts, pivot tables, multiplot charts, and sql statements.
+
+\subsection{Chart}
 \begin{figure}[h!]%
@@
     \begin{minipage}{.48\textwidth}
-    \chart[width=\textwidth, height=0.7\textwidth, chartType=line]{SELECT * FROM test;}
+    \chart[width=\textwidth, height=0.6\textwidth, chartType=line]{SELECT * FROM test;}
     \caption{Line Chart}
@@
     \begin{minipage}{.48\textwidth}
-    \chart[width=\textwidth, height=0.7\textwidth, chartType=bar]{SELECT * FROM test;}
+    \chart[width=\textwidth, height=0.6\textwidth, chartType=bar]{SELECT * FROM test;}
     \caption{Bar Chart}
@@
 \verb|\chart[options][queryForOverlay]{queryForPage}| \\
-Note: if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\[8pt]
+\textbf{Note:} if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay.\\
+\textbf{Note:} The first column of the query result is used for the x-Axis, all other columns are used for y-Axis.\\[8pt] 
 \noindent Options:\\[3pt]
@@
 \newpage
-\section{Multiplot Chart}
+\subsection{Multiplot Chart}
 \begin{figure}[h!]%
@@
 \verb|\multiplotChart[options][queryForOverlay]{queryForPage}| \\
-Note: If \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\
-Note: Both \verb|queryForOverlay| and \verb|queryForPage| must contain exactly two occurences of the ? character. These are later replaced with values from the xValues/yValues option.\\[8pt]
+\textbf{Note:} If \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\
+\textbf{Note:} Both \verb|queryForOverlay| and \verb|queryForPage| must contain exactly two occurences of the ? character. These are later replaced with values from the xValues/yValues option.\\[8pt]
 \noindent Options: \\[3pt]
 %
-Note: For multiplot charts all options of chart are also valid options.\\[4pt]
+\textbf{Note:} For multiplot charts all options of chart are also valid options.\\[4pt]
 %
@@
 \newpage
-\section{Pivot Table}
+\subsection{Pivot Table}
 \begin{figure}[h!]%
@@
 \verb|\pivotTable[options][queryForOverlay]{queryForPage}| \\
-Note: if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\[8pt]
+\textbf{Note:} if \verb|queryForOverlay| is omitted the \verb|queryForPage| is used for the overlay\\[8pt]
 \noindent Options: \\[3pt]
@@
 \newpage
-\section{Sql}
+\subsection{Sql}
 Example:\\[3pt]
@@
 \verb|\sql[textForPage]{queryForOverlay}| \\
-Note: if \verb|textForPage| is omitted the \verb|queryForOverlay| is used as text for the page.\\[8pt]
-
-\section{Sql data sources}
-\noindent Macros: \\[3pt]
-\verb|\dbSQLText{sqlQueryString}| \\
-\verb|\dbSQLFile{fileWithSqlQueries}| \\
-\verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
-Note: jdbcConnectionURL consist of jdbc followed by the name of the dbms (currently only postgresql and mysql are supported) followed by the url of the dbms followed by the database name (e.g.
-jdbc:postgresql://localhost:5432/postgres).\\
-Note: jdbcConnectionURL, user, and password are not stored in the output document.\\[8pt]
+\textbf{Note:} if \verb|textForPage| is omitted the \verb|queryForOverlay| is used as text for the page.\\[8pt]
 

### 2015-05-20 Documentation.tex
!

@@
 \usepackage{wrapfig}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage[a4paper,vmargin={20mm,40mm},hmargin={20mm,20mm}]{geometry}
@@
 \date{\today{}, Saarbrücken}
+
+\hypersetup{
+  colorlinks=true,
+  linkcolor=black,
+  urlcolor=blue!70!black
+}
 
@@
 \section{SQL queries for PDBF elements}
-Attributes with spaces sourrounded by square bracets. E.g. SELECT a AS [This is a test] FROM test;\\
+PDBF uses AlaSQL as database engine.
+For all available features visit \url{https://github.com/agershun/alasql}.
+Some functionality might not yet be available because we currently use not the latest version.\\
+Mostly sql syntax is standart sql, but for example attributes with spaces in their name have to be sourrounded by square brackets. E.g. SELECT a AS [This is a test] FROM test;\\
 
-\noindent TODO: write about special sql functions. AlaSQL (\url{https://github.com/agershun/alasql}).\\
+\noindent We extended AlaSQL with the following statistical sql functions:\\[5pt]
 \verb|GRUBBS_FILTER(arr, alpha)|: \\
+\indent Desc.: Filters an array of values with a \href{http://en.wikipedia.org/wiki/Grubbs'_test_for_outliers}{grubbs test}\\
+\indent Returns the filtered array.\\[3pt]
+\indent \option
+{arr}
+{Array on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}
+\indent \option
+{alpha}
+{significance level}
+{Number $>$ 0 AND Number $<$ 1}
+{\texttt{0.05}}\\[3pt]
+%
 \verb|MEAN(arr)|: \\
+\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Arithmetic_mean}{arithmetic mean} of an array of values.\\
+\indent Returns the arithmetic mean of the array.\\[3pt]
+\indent \option
+{arr}
+{Array on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}\\[3pt]
+%
 \verb|STDDEV_SAMP(arr)|: \\
+\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Standard_deviation}{ sample standard deviation} of an array of values.\\
+\indent Returns the sample standard deviation of the array.\\[3pt]
+\indent \option
+{arr}
+{Array on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}\\[3pt]
+%
 \verb|MARGIN_OF_ERROR(arr, alpha)|: \\
-\verb|CONF_INT(arr)|: \\
+\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Margin_of_error}{margin of error} of an array of values.\\
+\indent Returns the margin of error of the array.\\[3pt]
+\indent \option
+{arr}
+{Array on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}
+\indent \option
+{alpha}
+{significance level}
+{Number $>$ 0 AND Number $<$ 1}
+{\texttt{0.05}}\\[3pt]
+%
+\verb|CONF_INT(arr, alpha)|: \\
+\indent Desc.: Calculates the \href{http://en.wikipedia.org/wiki/Confidence_interval}{confidence interval} of an array.\\
+\indent Returns the confidence interval of the array.\\[3pt]
+\indent \option
+{arr}
+{Array on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}
+\indent \option
+{alpha}
+{significance level}
+{Number $>$ 0 AND Number $<$ 1}
+{\texttt{0.05}}\\[3pt]
+%
 \verb|T_TEST(arr1, arr2, alpha)|: \\
-\verb|WELCH_TEST(arr1, arr2, alpha)|:
-
+\indent Desc.: Performs a \href{http://en.wikipedia.org/wiki/Student's_t-test}{Student's t-test} on two arrays.\\
+\indent Returns true if they are indistinguishable, and returns false otherwise.\\[3pt]
+\indent \option
+{arr1, arr2}
+{Arrays on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}
+\indent \option
+{alpha}
+{significance level}
+{Number $>$ 0 AND Number $<$ 1}
+{\texttt{0.05}}\\[3pt]
+%
+\verb|WELCH_TEST(arr1, arr2, alpha)|:\\
+\indent Desc.: Performs a \href{http://en.wikipedia.org/wiki/Welch's_t_test}{Welch's t-test} on two arrays.\\
+\indent Returns true if they are indistinguishable, and returns false otherwise.\\[3pt]
+\indent \option
+{arr1, arr2}
+{Arrays on which the operation is performed.}
+{JavaScript array}
+{No default value. Value must always be set!}
+\indent \option
+{alpha}
+{significance level}
+{Number $>$ 0 AND Number $<$ 1}
+{\texttt{0.05}}\\[3pt]
+%
 \section{Sql data sources}

### 2015-05-20 minimal.tex
!

\documentclass[10pt]{scrartcl}
\usepackage{PDBF}
\usepackage[a4paper]{geometry}

\dbSQLText{CREATE TABLE test(a int, b int); INSERT INTO test VALUES(1, 10); INSERT INTO test VALUES(2, 100); INSERT INTO test VALUES(3, 120); INSERT INTO test VALUES(4, 100);}
%\dbSQLFile{somedb.sql}
%\dbSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtable}%

\begin{document}
\noindent
\sql{SELECT * FROM test;}\\[3pt]
\chart[width=\textwidth, height=0.45\textwidth, chartType=line]{SELECT * FROM test;}\\[3pt]
\chart[width=\textwidth, height=0.45\textwidth, chartType=bar]{SELECT * FROM test;}\\[3pt]
\chart[width=\textwidth, height=0.45\textwidth, chartType=signatureplot]{SELECT b FROM test;}\\[3pt]
\multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}\\[3pt]
\pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}
\end{document}

### 2015-05-20 Documentation.tex
fixed some small bugs in the documentation

@@
 %
-\noindent This toolkit is licensed unter the MIT License.\\
-
+\noindent This toolkit is licensed unter the MIT License.\\[4pt]
+%
 \noindent Thanks to the authors of:\\
@@
 PDF.js (\url{https://github.com/mozilla/pdf.js})\\
+google-closure-compiler (\url{https://github.com/google/closure-compiler})\\
 \newpage
@@
 {Sets the width of the chart.}
-{\LaTeX\relax length}
+{\LaTeX~length}
 {No default value. Value must always be set!}
@@
 {Sets the height of the chart.}
-{\LaTeX\relax length}
+{\LaTeX~length}
 {No default value. Value must always be set!}

### 2015-05-26 Documentation.tex
Update Documentation.tex

@@
 PDBF documents are a hybrid format. They are a valid PDF and a valid HTML page at the same time. 
-If you change the file extension to PDF and open it with an PDF viewer you can see the static part of the document.
-If you change the file extension to HTML and open it with a Browser (currently Chrome/Firefox/IE 10 supported) you can see the dynamic part of the document.
-For example if the static version contains an image of a chart which displays some data, then the dynamic version contains the actual raw data used to render the chart and renders the chart when opening the document.
+If you change the file extension to PDF and open it with an PDF viewer, you can see the static part of the document.
+If you change the file extension to HTML and open it with a Browser (currently Chrome/Firefox/Safari/IE 10 supported), you can see the dynamic part of the document.
+For example, if the static version contains an image of a chart which displays some data, then the dynamic version contains the actual raw data used to render the chart and renders the chart when opening the document.
 The advantage is that you can open the chart in an overlay view by clicking on the enlarge symbol at the upper left corner and then start to change parameters of the chart.
@@
 %
-\noindent PDBF files are created from LaTeX code and data in SQL format.
-The raw data can either be a SQL statements string, a file with SQL statements, or in a database (currently PostgreSQL/MySQL/MariaDB supported).
-In the LaTeX code one can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query.\\[4pt]
+\noindent PDBF files are created from LaTeX source code and a relational database.
+The raw data can either be a SQL statement string, a file with SQL statements, or contained in a database (currently PostgreSQL/MySQL/MariaDB supported).
+In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query.
+Read more in the [documentation](http://uds-datalab.github.io/PDBF/) (which is itself a PDBF document).\\[4pt]
 %
-\noindent PDBF toolkit is written in Java and LaTeX an can therefore be run on Windows/Mac/Linux. PDBF documents are also platform independent and run on any OS with a browser/PDF viewer.\\[4pt]
+\noindent PDBF toolkit is written in Java and LaTeX and can be used to compile documents on Windows (we are currently working to extend this to Mac and Linux).
+PDBF documents are also platform independent and run on any desktop OS (Windows, Linux, Mac) with a browser/PDF viewer.\\[4pt]
 %

### 2015-05-26 Documentation.tex
updated documentation

@@
 For example you can remove filtering functions that are applied to the raw data or change confidence levels to other values an see the results of these changes directly in the chart.
-\\[4pt]
-%
+\\
+
 \noindent PDBF files are created from LaTeX source code and a relational database.
@@
 In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query.
-Read more in the [documentation](http://uds-datalab.github.io/PDBF/) (which is itself a PDBF document).\\[4pt]
-%
+Read more in the [documentation](http://uds-datalab.github.io/PDBF/) (which is itself a PDBF document).\\
+
 \noindent PDBF toolkit is written in Java and LaTeX and can be used to compile documents on Windows (we are currently working to extend this to Mac and Linux).
-PDBF documents are also platform independent and run on any desktop OS (Windows, Linux, Mac) with a browser/PDF viewer.\\[4pt]
-%
-\noindent This toolkit is licensed unter the MIT License.\\[4pt]
-%
+PDBF documents are also platform independent and run on any desktop OS (Windows, Linux, Mac) with a browser/PDF viewer.\\
+
+\noindent This toolkit is licensed unter the MIT License.\\
+
+\newpage
 \noindent Thanks to the authors of:\\

### 2015-05-27 Documentation.tex
added some notes to config. changed documentclass of minimal example.

@@
 These \LaTeX\relax packages are required for PDBF documents:\\
-zref-savepos\\
-zref-abspage\\
-zref-user\\
+zref\\
 xcolor\\

### 2015-05-28 Documentation.tex
Fixed mac/linux compatibility. Updated documentation and readme. Added better error message for latex compile errors.

@@
 
-\noindent PDBF toolkit is written in Java and LaTeX and can be used to compile documents on Windows (we are currently working to extend this to Mac and Linux).
+\noindent PDBF toolkit is written in Java and LaTeX and can be used to compile documents on Windows, Mac and Linux.
 PDBF documents are also platform independent and run on any desktop OS (Windows, Linux, Mac) with a browser/PDF viewer.\\
@@
 Some functionality might not yet be available because we currently use not the latest version.\\
-Mostly sql syntax is standart sql, but for example attributes with spaces in their name have to be sourrounded by square brackets. E.g. SELECT a AS [This is a test] FROM test;\\
+Mostly sql syntax is standart sql, but for example attributes with spaces in their name have to be sourrounded by square brackets. E.g. SELECT attribute AS [attribute with spaces] FROM test;\\
 

### 2015-05-28 src/pdbf/latex/LaTeX_Compiler.java

@@ -82,7 +82,14 @@ public class LaTeX_Compiler {
 	    System.out.println("Usage: PDBF.jar LaTeX_file");
 	    System.exit(-1);
 	}
+	
 	String latexPath = args[0];
+	
+	if (!latexPath.endsWith(".tex")) {
+	    System.err.println("Error: Specified file has the wrong extension. Only .tex is supported!");
+	    System.exit(-1);
+	}
+	
 	File latex = new File(latexPath);
 	if (!latex.exists()) {
 	    System.err.println("Error: LaTeX file does not exist!");
@@ -115,7 +122,8 @@ public class LaTeX_Compiler {
 	    Process p = pb.start();
 	    p.waitFor();
 	} catch (Exception e) {
-	    e.printStackTrace();
+	    System.err.println("Error: LaTeX compilation failed! Reason: \n" + e.getMessage());
+	    System.exit(-1);
 	}
 
 	// Read JSON
@@ -240,7 +248,8 @@ public class LaTeX_Compiler {
 	    Process p = pb.start();
 	    p.waitFor();
 	} catch (Exception e) {
-	    e.printStackTrace();
+	    System.err.println("Error: LaTeX compilation failed! Reason: \n" + e.getMessage());
+	    System.exit(-1);
 	}
 
 	// Remove database entries from config

### 2015-05-28 Documentation.tex
Fixed typo in documentation

@@
 PDBF uses AlaSQL as database engine.
-For all available features visit \url{https://github.com/agershun/alasql}.
-Some functionality might not yet be available because we currently use not the latest version.\\
-Mostly sql syntax is standart sql, but for example attributes with spaces in their name have to be sourrounded by square brackets. E.g. SELECT attribute AS [attribute with spaces] FROM test;\\
+For all available features of the database visit \url{https://github.com/agershun/alasql}.
+Some functionality might not yet be available because we currently use not the latest version.\\[4pt]
+\textbf{Note:} Mostly sql syntax is standart sql, but for example attributes with spaces in their name have to be sourrounded by square brackets.
+E.g.
+SELECT attribute AS [attribute with spaces] FROM test;\\
 

### 2015-06-03 Documentation.tex
added minimal ui for multiplot

@@
 {topArr}
-{Sets the labels for the tod side.}
+{Sets the labels for the top side.}
 {Either a JavaScript array of Strings where each string corresponds to exactly one column (e.g \{"column1", "column2"\}) or a JavaScript array of Objects with a c property which corresponds to the number of columns this text should span and a text property which corresponds to the text-string.}

### 2015-06-03 out/web/main.js

@@ -74,7 +74,7 @@ function display(json, page, phantomJS) {
 					containerOver.setAttribute('style', 'position:fixed; z-index:-1; border:1px solid black; padding:10px; background:#DDDDDD; width:95%; height:87%; opacity:0; visibility:hidden; -webkit-transition:opacity 500ms ease-out; -moz-transition:opacity 500ms ease-out; -o-transition:opacity 500ms ease-out; transition:opacity 500ms ease-out; overflow:auto; white-space: nowrap;');
 					containerOver.id = json.name + "Big";
 					containerOver.className = "centerhv";
-					buildContainerMultiplotChartBig(json, containerOver, true);
+					buildContainerChartBig(json, containerOver, true);
 				}
 				containerOver.style.visibility = 'visible';
 				containerOver.style['z-index'] = 9;
@@ -198,12 +198,12 @@ function buildContainerChartBig(json, containerOver, initial) {
 		json.chart = c3.generate(optionsBig);
 	};
 
-	json.jsonBig = jQuery.extend(true, {}, json);
-	var chartdataCpy = getChartData(json);
-	json.resultBig = chartdataCpy.res;
 	var viewerContainer = document.getElementById("viewerContainer");
 	var tip = "Tip: Scroll to zoom the graph. Click and drag to pan the graph.<br/>";
 	var ref = prepopulateContainerOver(containerOver, viewerContainer, tip, [ json ], updateData, 'graph', true);
+	json.jsonBig = jQuery.extend(true, {}, json);
+	var chartdataCpy = getChartData(json);
+	json.resultBig = chartdataCpy.res;
 	var containerContent = ref.containerContent;
 	var containerControl = ref.containerControl;
 	var containerOptions = ref.options;
@@ -257,6 +257,8 @@ function buildContainerChartBig(json, containerOver, initial) {
 		json.chart = c3.generate(optionsBig);
 		containerOver.style['font-size'] = '' + rawZoomFactor * basetextsize + 'pt';
 	}
+	
+	containerOver.updateData = updateData;
 }
 
 function buildContainerChart(container, json, zoomFactor, style, containerOver) {
@@ -288,6 +290,10 @@ function buildContainerChart(container, json, zoomFactor, style, containerOver) 
 function buildContainerPivotBig(json, containerOver, initial) {
 	var basetextsize = 8;
 
+	var viewerContainer = document.getElementById("viewerContainer");
[…]

### 2015-07-04 src/pdbf/latex/LaTeX_Compiler.java
implemented DataText element

@@ -20,6 +20,7 @@ import com.google.gson.Gson;
 import com.google.gson.GsonBuilder;
 
 import pdbf.common.Chart;
+import pdbf.common.DataText;
 import pdbf.common.Database;
 import pdbf.common.Dimension;
 import pdbf.common.Overlay;
@@ -37,6 +38,7 @@ public class LaTeX_Compiler {
     private static ArrayList<String> cleanupfiles = new ArrayList<String>();
     private static ArrayList<String> copyfiles = new ArrayList<String>();
     private static ArrayList<String> preloadfiles = new ArrayList<String>();
+    private static ArrayList<String> datatextfiles = new ArrayList<String>();
     private static Gson gson;
     private static String suffix;
     private static Dimension dimOrg;
@@ -141,8 +143,8 @@ public class LaTeX_Compiler {
 	    dimOrg = gson.fromJson(json2, Dimension.class);
 	    for (int i = 0; i < overlays.length; ++i) {
 		Visualization v = overlays[i].type;
-		if (v instanceof Text) {
-		    Text t = (Text) overlays[i].type;
+		if (v instanceof Text || v instanceof DataText) {
+		    Visualization t = overlays[i].type;
 		    t.x1 = t.x1 / dimOrg.width;
 		    t.x2 = t.x2 / dimOrg.width;
 		    t.y1 = (t.y1 + 65536 * t.fontsize) / dimOrg.height;
@@ -204,6 +206,9 @@ public class LaTeX_Compiler {
 	    if (overlays[i].type instanceof Chart) {
 		processChart(overlays[i]);
 	    }
+	    if (overlays[i].type instanceof DataText) {
+		processDataChart(overlays[i]);
+	    }
 	}
 
 	for (Process p : processes) {
@@ -237,6 +242,20 @@ public class LaTeX_Compiler {
 	}
 	try {
 	    FileUtils.writeStringToFile(new File("preload"), preload);
+	} catch (IOException e2) {
+	    e2.printStackTrace();
+	}
+	
+	String aux = "";
+	for (String data : datatextfiles) {
+	    try {
+		aux += "\\expandafter\\gdef\\csname pdbf@" + data.substring(0, data.length()-5) + "\\endcsname{" + FileUtils.readFileToString(new File(data)) + "}\n";
+	    } catch (IOException e) {
+		e.printStackTrace();
+	    }
+	}
+	try {
[…]

### 2015-07-06 Documentation.tex
finished dataText

@@
 \verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
-\textbf{Note:} jdbcConnectionURL consist of jdbc followed by the name of the dbms (currently only postgresql and mysql are supported) followed by the url of the dbms followed by the database name (e.g.
+\textbf{Note:} jdbcConnectionURL consist of jdbc followed by the name of the dbms (only postgresql and mysql are currently supported) followed by the url of the dbms followed by the database name (e.g.
 jdbc:postgresql://localhost:5432/postgres).\\
-\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output document.
+\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.
 
@@
 
-\option %TODO: not yet implemented
-{logscale}
-{If set, the y axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
-{\texttt{true} or \texttt{false}}
-{\texttt{false}}
-
-\option %TODO: not yet implemented
-{overlap}
-{Unused.}
-{\texttt{TODO: ...}}
-{\texttt{TODO: ...}}
-
-\option %TODO: not yet implemented
-{legendpos}
-{Sets the position of the legend of the chart.}
-{\texttt{TODO: ...}}
-{\texttt{TODO: ...}}
-
-\option %TODO: not yet implemented
-{includeZero}
-{Set the minimum of the range of the y-axis to zero.}
-{\texttt{true} or \texttt{false}}
-{\texttt{false}}
-
-\option %TODO: not yet implemented
-{drawPoints}
-{Whether to show each point in line.}
-{\texttt{true} or \texttt{false}}
-{\texttt{false}}
-
-\option %TODO: not yet implemented
-{fillGraph}
-{Whether to fill the area below the graph.}
-{\texttt{true} or \texttt{false}}
-{\texttt{false}}
-
-\option %TODO: not yet implemented
-{showRangeSelector}
-{Whether to show a range selector for the x-axis below the chart.}
-{\texttt{true} or \texttt{false}}
-{\texttt{false}}
+%\option %TODO: not yet implemented
+%{logscale}
+%{If set, the y axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
+%{\texttt{true} or \texttt{false}}
+%{\texttt{false}}
+%
+%\option %TODO: not yet implemented
+%{overlap}
+%{Unused.}
+%{\texttt{TODO: ...}}
+%{\texttt{TODO: ...}}
+%
+%\option %TODO: not yet implemented
+%{legendpos}
+%{Sets the position of the legend of the chart.}
+%{\texttt{TODO: ...}}
+%{\texttt{TODO: ...}}
+%
+%\option %TODO: not yet implemented
+%{includeZero}
+%{Set the minimum of the range of the y-axis to zero.}
+%{\texttt{true} or \texttt{false}}
+%{\texttt{false}}
+%
+%\option %TODO: not yet implemented
+%{drawPoints}
+%{Whether to show each point in line.}
+%{\texttt{true} or \texttt{false}}
+%{\texttt{false}}
+%
+%\option %TODO: not yet implemented
+%{fillGraph}
+%{Whether to fill the area below the graph.}
+%{\texttt{true} or \texttt{false}}
+%{\texttt{false}}
+%
+%\option %TODO: not yet implemented
+%{showRangeSelector}
+%{Whether to show a range selector for the x-axis below the chart.}
+%{\texttt{true} or \texttt{false}}
+%{\texttt{false}}
 
@@
 
-\newpage
 \subsection{Sql}
@@
 \textbf{Note:} if \verb|textForPage| is omitted the \verb|queryForOverlay| is used as text for the page.\\[8pt]
+
+\noindent Options: \\[3pt]
+
+\option
+{color}
+{The textcolor of the displayed text.}
+{\texttt{white}, \texttt{black}, \texttt{red}, \texttt{green}, \texttt{blue}, \texttt{cyan}, \texttt{magenta}, \texttt{yellow} or xcolor syntax (see \href{http://mirror.unicorncloud.org/CTAN/macros/latex/contrib/xcolor/xcolor.pdf}{here}).
+Can also be redefined globally (\texttt{pdbfDynamicTextColor})}
+{\texttt{blue}}
+
+\subsection{DataText}
+Produce text from sql-statements.\\[3pt]
+Example:\\[3pt]
+The table ``test'' contains \dataText{SELECT COUNT(*) FROM test;} tuples.\\
+The table ``test'' contains the following tuples: \dataText{SELECT * FROM test;}.\\[3pt]
+%
+\noindent Macros: \\[3pt]
+\verb|\dataText{query}| \\[8pt]
+
+\noindent Options: \\[3pt]
+
+\option
+{color}
+{The textcolor of the displayed text.}
+{\texttt{white}, \texttt{black}, \texttt{red}, \texttt{green}, \texttt{blue}, \texttt{cyan}, \texttt{magenta}, \texttt{yellow} or xcolor syntax (see \href{http://mirror.unicorncloud.org/CTAN/macros/latex/contrib/xcolor/xcolor.pdf}{here}).
+Can also be redefined globally (\texttt{pdbfDynamicTextColor})}
+{\texttt{blue}}
 
@@
 These are not fully compatible with PDBF.
-If you really have to use the utf8 package then you cant use special characters inside PDBF commands.
+If you really have to use the utf8 package then you cant use special characters inside options and queries of the PDBF package.
 \end{itemize}

### 2015-07-06 src/pdbf/latex/LaTeX_Compiler.java

@@ -128,52 +128,8 @@ public class LaTeX_Compiler {
 	    System.exit(-1);
 	}
 
-	// Read JSON
-	GsonBuilder builder = new GsonBuilder();
-	builder.disableHtmlEscaping().serializeNulls();
-	builder.registerTypeAdapter(Visualization.class, new VisualizationTypeAdapter());
-	builder.registerTypeAdapter(Alasql.class, new Alasql());
-	builder.registerTypeAdapter(Table.class, new Table());
-	gson = builder.create();
-	Overlay[] overlays = null;
-	try {
-	    String json = FileUtils.readFileToString(new File("config.json"), Tools.utf8);
-	    overlays = gson.fromJson(json, Overlay[].class);
-	    String json2 = FileUtils.readFileToString(new File("dim.json"), Tools.utf8);
-	    dimOrg = gson.fromJson(json2, Dimension.class);
-	    for (int i = 0; i < overlays.length; ++i) {
-		Visualization v = overlays[i].type;
-		if (v instanceof Text || v instanceof DataText) {
-		    Visualization t = overlays[i].type;
-		    t.x1 = t.x1 / dimOrg.width;
-		    t.x2 = t.x2 / dimOrg.width;
-		    t.y1 = (t.y1 + 65536 * t.fontsize) / dimOrg.height;
-		    t.y2 = t.y2 / dimOrg.height;
-		} else if (v instanceof Chart) {
-		    Chart c = (Chart) overlays[i].type;
-		    c.x1 = c.x1 / dimOrg.width;
-		    c.x2 = c.x2 / dimOrg.width;
-		    c.y1 = c.y1 / dimOrg.height;
-		    c.y2 = c.y2 / dimOrg.height;
-		}
-
-		if (!(v instanceof Database)) {
-		    v.zoom = v.fontsize / 12.0 * v.zoom;
-		    if (v.aggregationattributeBig.equals("")) {
-			v.aggregationattributeBig = v.aggregationattribute;
-		    }
-		    if (v.aggregationBig.equals("")) {
-			v.aggregationBig = v.aggregation;
-		    }
-		}
-	    }
-	} catch (Exception e) {
-	    e.printStackTrace();
-	}
-	if (overlays == null) {
-	    System.err.println("Error: Deserialization failed!");
-	    System.exit(-1);
-	}
+	Overlay[] overlays = readJSONconfig();
+	
 	// Split
 	File f = new File("db.sql");
 	File f2 = new File("db.json");
[…]

### 2015-07-08 Documentation.tex
implemented linking dataText with charts, fixed some bugs

@@
 %
-Common options\\[4pt]
-%
+\option
+{name}
+{Sets the name of this element. This is only useful for linking Elements. See DataText.}
+{String}
+{\texttt{undefined}}
+
 \option
@@
 \noindent Macros: \\[3pt]
-\verb|\sql[textForPage]{queryForOverlay}| \\
+\verb|\sql[options][textForPage]{queryForOverlay}| \\
 \textbf{Note:} if \verb|textForPage| is omitted the \verb|queryForOverlay| is used as text for the page.\\[8pt]
@@
 \noindent Macros: \\[3pt]
-\verb|\dataText{query}| \\[8pt]
+\verb|\dataText[options]{query}| \\[8pt]
 
@@
 {\texttt{blue}}
+
+\option
+{linkTo}
+{When this option is set then this DataText and the element with name equal to the value of \texttt{linkTo} are linked.
+When the cursor is moved above the DataText then in the linked element the specified \texttt{linkSelector} specifies which part of the element is highlighted and annotated with the value of \texttt{linkLabel}.}
+{String which value is the name of some other element. \textbf{Currently only chart elements can be used for \texttt{linkTo}.}}
+{\texttt{undefined}}
+
+\option
+{linkSelector}
+{The selector that specifies which part of the \texttt{linkTo} element is highlighted.}
+{\texttt{Number}, \texttt{Date} or \texttt{String} depending on the type of the \texttt{linkTo} element.
+For charts the type depends on the datatype of the xAxis and the value specifies which part of the xAxis is highlighted.}
+{\texttt{undefined}}
+
+\option
+{linkLabel}
+{The text that is displayed near the highlighted area.}
+{String}
+{\texttt{undefined}}
 

### 2015-07-08 minimal.tex
added some notes to config. changed documentclass of minimal example.
implemented DataText element
finished dataText
implemented linking dataText with charts, fixed some bugs

@@
-\documentclass[10pt]{scrartcl}
+\documentclass[10pt]{minimal}
 \usepackage{PDBF}
@@
 \sql{SELECT * FROM test;}\\[3pt]
-\chart[width=\textwidth, height=0.45\textwidth, chartType=line]{SELECT * FROM test;}\\[3pt]
+The table ``test'' contains \dataText{SELECT COUNT(*) FROM test;} tuples.\\[3pt]
+The following points are linked to the chart below: \dataText[linkTo=testname, linkSelector=2, linkLabel=Value at 2]{SELECT * FROM test WHERE a=2;}, \dataText[linkTo=testname, linkSelector=3, linkLabel=Value at 3]{SELECT * FROM test WHERE a=3;}.\\[3pt]
+\chart[width=\textwidth, height=0.45\textwidth, chartType=line, name=testname]{SELECT * FROM test;}\\[3pt]
 \chart[width=\textwidth, height=0.45\textwidth, chartType=bar]{SELECT * FROM test;}\\[3pt]

### 2015-07-22 Documentation.tex
implemented dataTables and fixed some minor bugs

@@
-\documentclass[11pt]{scrartcl}
+\documentclass[11pt]{article}
 \usepackage[utf8]{inputenc}
@@
 geometry\\
+array\\
 
@@
 
-\subsection{Sql}
+\subsection{SQL}
+Clickable SQL-statements\\[3pt]
 Example:\\[3pt]
@@
 \subsection{DataText}
-Produce text from sql-statements.\\[3pt]
+Write text that is the result of a SQL-statement.\\[3pt]
 Example:\\[3pt]
@@
 
+\subsection{DataTable}
+Write a \LaTeX\space table that is the result of a SQL-statement.\\[3pt]
+Example:\\[3pt]
+\dataTable{SELECT * FROM test;}\\[5pt]
+%
+\noindent Macros: \\[3pt]
+\verb|\dataTable[options]{query}| \\[8pt]
+
+\noindent Options: \\[3pt]
+
+\option
+{color}
+{The textcolor of the displayed text.}
+{\texttt{white}, \texttt{black}, \texttt{red}, \texttt{green}, \texttt{blue}, \texttt{cyan}, \texttt{magenta}, \texttt{yellow} or xcolor syntax (see \href{http://mirror.unicorncloud.org/CTAN/macros/latex/contrib/xcolor/xcolor.pdf}{here}).
+Can also be redefined globally (\texttt{pdbfDynamicTextColor})}
+{\texttt{blue}}
+
+\option
+{verticalLines}
+{Specifies the vertical lines for the table.}
+{\texttt{a} for all, \texttt{i} for inner, \texttt{o} for outer, \texttt{n} for none, or you can use the default tabular syntax (e.g. \texttt{|l|c|r|})}
+{\texttt{a}}
+
+\option
+{horizontalLinesHeader}
+{Specifies the horizontal lines for the table header.}
+{\texttt{a} for all, \texttt{i} for inner, \texttt{o} for outer, \texttt{n} for none, or you can specify a custom pattern that is repeatedly applied with the following format: h means line, b means no line, the space character delimits format specifiers.
+E.g.
+``hh b'' means: double line followed by no line followed by double line followed by no line ...
+and so on}
+{\texttt{a}}
+
+\option
+{horizontalLinesBody}
+{Specifies the horizontal lines for the table body, excluding the border above the first tuple, which is controlled by horizontalLinesHeader.}
+{\texttt{a} for all, \texttt{i} for inner, \texttt{o} for outer, \texttt{n} for none, or you can specify a custom pattern that is repeatedly applied with the following format: h means line, b means no line, the space character delimits format specifiers.
+E.g.
+``hh b'' means: double line followed by no line followed by double line followed by no line ...
+and so on}
+{\texttt{a}}
+
 \section{F.A.Q.}

### 2015-07-22 minimal.tex
implemented dataTables and fixed some minor bugs

@@
-\documentclass[10pt]{minimal}
+\documentclass[10pt]{article}
 \usepackage{PDBF}
@@
 \multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}\\[3pt]
-\pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}
+\pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}\\[3pt]
+\dataTable[verticalLines=i, horizontalLinesHeader=b hh, horizontalLinesBody=i]{SELECT * FROM test;}\\[3pt]
 \end{document}

### 2015-08-05 data/main.js
removed unused folders

@@ -1540,7 +1540,7 @@ function getChartOptions(json, zoomFactor, values, chart) {
 			alert('Unknown chart type. Type was: "' + json.type.I.chartType + '"');
 	}
 	
-	if (values.json != undefined && isNaN(values.json[0][values.x]) && (options == undefined || options.axis == undefined || options.axis.x == undefined || options.axis.x.type == undefined)) {
+	if (values.json != undefined && (isNaN(values.json[0][values.x]) || values.json[0][values.x] == "") && (options == undefined || options.axis == undefined || options.axis.x == undefined || options.axis.x.type == undefined)) {
 		if (isValidDate(values.json[0][values.x])) {
 			options.axis.x.type = "timeseries";
 		} else {

### 2015-08-05 src/pdbf/common/Tools.java

@@ -27,8 +27,9 @@ public class Tools {
 	return Base64.encodeBase64String(string.getBytes(utf8));
     }
     
-    public static String escapeQuotes(File file) throws IOException {
+    public static String escapeSpecialChars(File file) throws IOException {
 	String tmp = FileUtils.readFileToString(file);
+	tmp = tmp.replace("\\", "\\\\");
 	tmp = tmp.replace("\"", "\\\"");
 	return tmp;
     }

### 2015-08-05 src/pdbf/latex/LaTeX_Compiler.java

@@ -384,7 +384,7 @@ public class LaTeX_Compiler {
 	    String viewer;
 	    String viewerHEAD = FileUtils.readFileToString(new File("data/templateHEADimages.html"), Tools.utf8);
 	    String viewerTAIL = FileUtils.readFileToString(new File("data/templateTAILimages.html"), Tools.utf8);
-	    viewer = viewerHEAD + "dim_base64 = \"" + Tools.encodeStringToBase64Binary(gson.toJson(dim)) + "\";\r\n" + "json_base64 = \"" + Tools.encodeStringToBase64Binary(gson.toJson(o)) + "\";\r\n" + "db_base64 = \"" + Tools.encodeFileToBase64Binary(new File("db.sql")) + "\";\r\n" + "dbjson_base64 = \"" + Tools.escapeQuotes(new File("db.json")) + "\";\r\n" + viewerTAIL;
+	    viewer = viewerHEAD + "dim_base64 = \"" + Tools.encodeStringToBase64Binary(gson.toJson(dim)) + "\";\r\n" + "json_base64 = \"" + Tools.encodeStringToBase64Binary(gson.toJson(o)) + "\";\r\n" + "db_base64 = \"" + Tools.encodeFileToBase64Binary(new File("db.sql")) + "\";\r\n" + "dbjson_base64 = \"" + Tools.escapeSpecialChars(new File("db.json")) + "\";\r\n" + viewerTAIL;
 	    FileUtils.writeStringToFile(new File("data/" + o.name + ".html"), viewer, Tools.utf8);
 	} catch (Exception e) {
 	    e.printStackTrace();
@@ -410,7 +410,7 @@ public class LaTeX_Compiler {
 	    String viewer;
 	    String viewerHEAD = FileUtils.readFileToString(new File("data/templateHEADimages.html"), Tools.utf8);
 	    String viewerTAIL = FileUtils.readFileToString(new File("data/templateTAILimages.html"), Tools.utf8);
[…]

### 2015-08-05 Documentation.tex
small changes
removed unused folders
updated documentation and readme, added version file

@@
 \maketitle
+\noindent \textbf{For PDBF \input{VERSION.md}}
 
 \section{Abstract}
-PDBF documents are a hybrid format. They are a valid PDF and a valid HTML page at the same time. 
+PDBF documents are a hybrid format.
+They are a valid PDF and a valid HTML page at the same time.
+You can now optionally add an VirtualBox OVA file with a complete operating system to the PDBF document.
+Yes, this means that the resulting file is a valid PDF, HTML, and OVA file at the same time.
 If you change the file extension to PDF and open it with an PDF viewer, you can see the static part of the document.
 If you change the file extension to HTML and open it with a Browser (currently Chrome/Firefox/Safari/IE 10 supported), you can see the dynamic part of the document.
-For example, if the static version contains an image of a chart which displays some data, then the dynamic version contains the actual raw data used to render the chart and renders the chart when opening the document.
-The advantage is that you can open the chart in an overlay view by clicking on the enlarge symbol at the upper left corner and then start to change parameters of the chart.
-For example you can remove filtering functions that are applied to the raw data or change confidence levels to other values an see the results of these changes directly in the chart.
-\\
+And if an ova file is attached you can also change the file extension to OVA and install and run the attached operating system.\\
+The difference between the PDF and the HTML version is that the PDF version contains static version of all PDBF elements, whereas the HTML version is dynamic.
+For example you can zoom into graphs, temporarly remove dataseries from the graph, inspect and change the underling query of the PDBF element and see the result of the change directly in the browser.\\
+This approach works completely offline. No internet connection is required, neither at compile time, nor at viewing time.
 
@@
 In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query.
-Read more in the [documentation](http://uds-datalab.github.io/PDBF/) (which is itself a PDBF document).\\
+Read more in the \href{http://uds-datalab.github.io/PDBF/}{documentation}, which is itself is a PDBF document.\\
 
@@
 
-\noindent This toolkit is licensed unter the MIT License.\\
+\noindent A \href{https://infosys.uni-saarland.de/publications/p1972-dittrich.pdf}{demo paper} of our tool is appearing at \href{http://www.vldb.org/2015/}{VLDB 2015}.
+
+\section{License}
+\noindent This toolkit is licensed unter the MIT License. See LICENSE.md file.\\
 
 \newpage
-\noindent Thanks to the authors of:\\
+\tableofcontents
+\newpage
+
+\section{Thanks to}
 phantomJS (\url{https://github.com/ariya/phantomjs})\\
@@
 google-closure-compiler (\url{https://github.com/google/closure-compiler})\\
-\newpage
-\tableofcontents
-\newpage
+
+\section{How to use the PDBF compiler}
+Right now there are only two modes: \\[4pt]
+-- The standart mode which compiles a tex file into the hybrid PDBF format:\\
+\verb|java -jar PDBF.jar LaTeX_file| \\[2pt]
+-- The VM mode which attaches a VirtualBox .ova file to an existing PDBF:\\
+\verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|
 
 \section{Requirements}
-These \LaTeX\relax packages are required for PDBF documents:\\
-zref\\
-xcolor\\
-graphicx\\
-xstring\\
-xparse\\
-geometry\\
-array\\
-
+These \LaTeX\relax packages are required for PDBF documents:
+\begin{itemize}
+\item \verb|zref|
+\item \verb|xcolor|
+\item \verb|graphicx|
+\item \verb|xstring|
+\item \verb|xparse|
+\item \verb|geometry|
+\item \verb|array|
+\end{itemize}
 \noindent Furthermore it is needed that your \LaTeX\relax document uses the geometry package to specify the page size.\\
 
-\noindent \textbf{Warning:} There are problems with inputenc UTF8 package. If you really have to use the inputenc UTF8 package then you cant use non-ASCII characters inside queries.
+\noindent \textbf{Warning:} There are problems with \textbf{inputenc UTF8} and \textbf{pgfpages}.
+If you really have to use the inputenc UTF8 package then you cant use non-ASCII characters inside queries.
+The pgfpages package is incompatible to this package (Its fixable but requires additional effort).
 
@@
 %
-\section{Sql data sources}
+\section{SQL data sources}
 \noindent Macros: \\[3pt]
 \verb|\dbSQLText{sqlQueryString}| \\
-\verb|\dbSQLFile{fileWithSqlQueries}| \\
+\verb|\dbSQLFile{fileWithSQLQueries}| \\
 \verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
@@
 jdbc:postgresql://localhost:5432/postgres).\\
-\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.
+\textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.\\[4pt]
+\textbf{WARNING:} \verb|\dbSQLText| and \verb|\dbSQLFile| are currently horribly, even for medium sized (more than 100 tuples) amounts of data. Use \verb|\dbSQLJDBC| instead.
 
-\newpage
 \section{PDBF elements}
-Currently there are four different PDBF elements: charts, pivot tables, multiplot charts, and sql statements.
-
 \subsection{Chart}
@@
 
+\newpage
 \noindent Macros: \\[3pt]
@@
 
+\option
+{includeZero}
+{Set the minimum of the range of the y-axis to zero.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option
+{drawPoints}
+{Whether to show each point in line.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option
+{fillGraph}
+{Whether to fill the area below the graph.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
+\option
+{showRangeSelector}
+{Whether to show a range selector for the x-axis below the chart.}
+{\texttt{true} or \texttt{false}}
+{\texttt{false} for in-page chart, \texttt{true} for overlay chart}
+
 %\option %TODO: not yet implemented
@@
 %
-%\option %TODO: not yet implemented
-%{includeZero}
-%{Set the minimum of the range of the y-axis to zero.}
-%{\texttt{true} or \texttt{false}}
-%{\texttt{false}}
-%
-%\option %TODO: not yet implemented
-%{drawPoints}
-%{Whether to show each point in line.}
-%{\texttt{true} or \texttt{false}}
-%{\texttt{false}}
-%
-%\option %TODO: not yet implemented
-%{fillGraph}
-%{Whether to fill the area below the graph.}
-%{\texttt{true} or \texttt{false}}
-%{\texttt{false}}
-%
-%\option %TODO: not yet implemented
-%{showRangeSelector}
-%{Whether to show a range selector for the x-axis below the chart.}
-%{\texttt{true} or \texttt{false}}
-%{\texttt{false}}
 
-\newpage
 \subsection{Multiplot Chart}
@@
 \begin{itemize}
-\item Q.: The overlay is not on the right position.\\
-A.: Most likely you use pages with different sizes in your document. This is currently not supported by PBDF 
+\item Q.: The overlay is not on the right position!\\
+A.: Most likely you use pages with different sizes in your document or you use the pgfpages package. This is currently not supported by PBDF 
 \item Q.: I get \verb|Error: paperwidth value missing! Did you forgot to specify the| \\ \verb|papersize via the geometry package?| when compiling my document.\\

### 2015-08-05 src/pdbf/vm/VM_Compiler.java

@@ -12,15 +12,15 @@ public class VM_Compiler {
     public static void main(String[] args) {
 	System.out.println("Compiling VM...");
 
-	String a = new File(args[0]).getName();
+	String a = new File(args[1]).getName();
 	if (!a.toUpperCase().endsWith(".HTML")) {
 	    System.err.println("Error: Only .HTML files are supported!");
 	    System.exit(-1);
 	}
-	String basename = args[0].substring(0, args[0].length() - 5);
+	String basename = args[1].substring(0, args[1].length() - 5);
 	
 	try {
-	    StringBuilder sb = new StringBuilder(FileUtils.readFileToString(new File("Space Invaders.ova"), StandardCharsets.ISO_8859_1));
+	    StringBuilder sb = new StringBuilder(FileUtils.readFileToString(new File(args[2]), StandardCharsets.ISO_8859_1));
 	    String replace = "%PDF-1.5\n%.ovf\0\n1 0 obj\nstream\n<head><meta charset=UTF-8><script>";
 	    sb.replace(0, replace.length(), replace);
 	    
@@ -50,7 +50,7 @@ public class VM_Compiler {
 	    
 	    String removeFromHTML = "%PDF-1.5\n%<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>" + "<head><meta charset=\"utf-8\"><!--\n1337 0 obj\nstream";
 	    String addToHTML = "</script>";
-	    String html = FileUtils.readFileToString(new File(args[0]), StandardCharsets.ISO_8859_1).substring(removeFromHTML.length());
+	    String html = FileUtils.readFileToString(new File(args[1]), StandardCharsets.ISO_8859_1).substring(removeFromHTML.length());
 	    html = addToHTML + html;
 	    
 	    //Fix tar
@@ -109,6 +109,7 @@ public class VM_Compiler {
 	    e.printStackTrace();
 	}
 
+	System.out.println("Finished...");
     }
 
 }

### 2015-08-06 pdbf-doc.tex
changed some filenames
replaced url to demo paper

--- 2015-08-05 Documentation.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage{PDBF}
+\usepackage{pdbf}
 \usepackage{wrapfig}
@@
 
-\noindent A \href{https://infosys.uni-saarland.de/publications/p1972-dittrich.pdf}{demo paper} of our tool is appearing at \href{http://www.vldb.org/2015/}{VLDB 2015}.
+\noindent A \href{https://infosys.uni-saarland.de/publications/p1972-dittrich.html}{demo paper} of our tool is appearing at \href{http://www.vldb.org/2015/}{VLDB 2015}.
 

### 2015-08-07 pdbf-doc.tex
Update pdbf-doc.tex

@@
 The raw data can either be a SQL statement string, a file with SQL statements, or contained in a database (currently PostgreSQL/MySQL/MariaDB supported).
-In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements are supported) is created from the raw data with options and an SQL query.
+In the LaTeX code you can then specify how the PDBF element (currently charts/pivot tables/multiplot charts/sql statements/dataTexts/dataTables are supported) is created from the raw data with options and an SQL query.
 Read more in the \href{http://uds-datalab.github.io/PDBF/}{documentation}, which is itself is a PDBF document.\\

### 2015-08-28 pdbf-doc.tex
added website to documentation, fixed bold font for mutliplot charts
bugfix for pretty printing sql in alasql, fixed y-label offset in c3, codemirror auto-resize, fixed logscale, fixed typo in documentation

@@
 
-\title{PDBF Documentation}
+\title{\href{https://github.com/uds-datalab/PDBF}{PDBF} Documentation}
 \author{
@@
 \maketitle
-\noindent \textbf{For PDBF \input{VERSION.md}}
+\noindent \textbf{Website:} \url{https://github.com/uds-datalab/PDBF}\\
+\noindent \textbf{\input{VERSION.md}}
 
@@
 \textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.\\[4pt]
-\textbf{WARNING:} \verb|\dbSQLText| and \verb|\dbSQLFile| are currently horribly, even for medium sized (more than 100 tuples) amounts of data. Use \verb|\dbSQLJDBC| instead.
+\textbf{WARNING:} \verb|\dbSQLText| and \verb|\dbSQLFile| are currently horribly slow, even for medium sized (more than 100 tuples) amounts of data. Use \verb|\dbSQLJDBC| instead.
 

### 2015-09-03 pdbf-doc.tex
updated documentation Version 1.0.3

@@
 google-closure-compiler (\url{https://github.com/google/closure-compiler})\\
+lz-string (\url{https://github.com/pieroxy/lz-string})\\
 

### 2015-09-08 pdbf-doc.tex
now pdbf.jar should run independently of the current workingdir

@@
 
+\tableofcontents
+\newpage
+
 \section{Abstract}
@@
 
-\newpage
-\tableofcontents
-\newpage
+\section{Getting started}
+\begin{itemize}
+\item Make sure you have a Java Runtime (version $>=$ 1.7) installed
+\item \href{https://github.com/uds-datalab/PDBF/archive/gh-pages.zip}{Download the latest version}
+\item Extract zip and change workingdir to extracted folder
+\item Adjust config.cfg
+\item Try to compile pdbf-doc.tex file with this command: java -jar pdbf.jar pdbf-doc.tex
+\item Open pdbf-doc.html, this is the final output of the compilation process, if you rename it to ".pdf" it is also a valid pdf-document
+\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: java -jar pdbf.jar --vm pdbf-doc.html vldb-Invaders.ova
+\item You can play around with the minimal.tex file which contains a minimal example of PDBF usage
+\item For further information take a look at the \href{http://uds-datalab.github.io/PDBF/}{documentation}
+\end{itemize}
 

### 2015-09-08 external-tools/captureData.js

@@ -1,7 +1,7 @@
 var system = require('system');
 var fs = require('fs');
-if (system.args.length !== 2) {
-	console.log('Usage: ' + system.args[0] + ' webpagefile');
+if (system.args.length !== 3) {
+	console.log('Usage: ' + system.args[0] + ' webpagefile basedir');
 	phantom.exit();
 } 
 
@@ -14,6 +14,8 @@ page.onError = function (msg, trace) {
     });
 	phantom.exit();
 };
+
+fs.changeWorkingDirectory(system.args[2]);
 
 page.open(system.args[1], function() {
 	var size = page.evaluate(function() {

### 2015-09-08 pdbf-doc.tex
updated documentation

@@
 
+\section{Bugs, Suggestions, Feature requests}
+If you encounter bugs, have suggestions or have a feature request, then please go to the issue page open a new issue if necessary and explain your concern.
+You can also write us an email (ichbinkeinreh at t-online.de or jens.dittrich at cs.uni-saarland.de).
+
 \section{Abstract}
@@
 
-\noindent PDBF toolkit is written in Java and LaTeX and can be used to compile documents on Windows, Mac and Linux.
+\noindent PDBF toolkit is written in Java and LaTeX and can be used to compile documents on Windows, Mac, and Linux.
 PDBF documents are also platform independent and run on any desktop OS (Windows, Linux, Mac) with a browser/PDF viewer.\\
 
-\noindent A \href{https://infosys.uni-saarland.de/publications/p1972-dittrich.html}{demo paper} of our tool is appearing at \href{http://www.vldb.org/2015/}{VLDB 2015}.
+\noindent A \href{https://infosys.uni-saarland.de/publications/p1972-dittrich.html}{demo paper} of our tool appeared at \href{http://www.vldb.org/2015/}{VLDB 2015}.
 

### 2015-09-09 hello_world.tex
no message

\documentclass[10pt]{article}
\usepackage{pdbf}
\usepackage[a4paper]{geometry}

% Option 1:
% use q SQL dump inserting data into a mockup database
\dbSQLText{
CREATE TABLE HelloWorld(id int, text varchar); INSERT INTO HelloWorld VALUES(10, 'Hello World!'); INSERT INTO HelloWorld VALUES(120, 'Hallo Welt!');  INSERT INTO HelloWorld VALUES(100, 'bonjour le monde!');  INSERT INTO HelloWorld VALUES(42, 'Hello World!');}

% Option 2:
% use a jdbc connecction to an existing database
% note: this connection is opnly required when creating the document; all data will be embedded into the document when compiling the document; afterwards the jdbc connection is not required anymore

%\dbSQLFile{somedb.sql}
%\dbSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtable}%

\begin{document}

\dataTable[verticalLines=i, horizontalLinesHeader=b hh, horizontalLinesBody=i]{SELECT text FROM HelloWorld WHERE id=10;}\\[3pt]

\noindent
The table ``HelloWorld'' contains the following \dataText{SELECT COUNT(*) FROM HelloWorld;} tuples:

\dataTable[verticalLines=i, horizontalLinesHeader=b hh, horizontalLinesBody=i]{SELECT * FROM HelloWorld;}\\[3pt]

\noindent Column ``id'' contains \dataText{SELECT COUNT(DISTINCT id) FROM HelloWorld;} unique values whereas column ``text'' contains \dataText{SELECT COUNT(DISTINCT text) FROM HelloWorld;} unique values. The smallest id is  \dataText{SELECT min(id) FROM HelloWorld;}, the biggest id is  \dataText{SELECT max(id) FROM HelloWorld;}.

\vspace*{0.5cm}
\noindent\textbf{Notice}:

\noindent If you opened ``hello world.html'', try to rename this file to ``hello world.pdf'' and see what happens.

\noindent If you opened ``hello world.pdf'', try to rename this file to ``hello world.html'' and see what happens.

\end{document}

### 2015-09-14 pdbf-doc.tex
csv as data source Version 1.0.4

@@
 }
+
+\def\pdbfQuality{4.0}
 
@@
 %
-\section{SQL data sources}
+\section{Data sources}
+
+\subsection{SQL}
 \noindent Macros: \\[3pt]
@@
 \verb|\dbSQLFile{fileWithSQLQueries}| \\
+\textbf{Note:} Relative filepaths are relative to the tex file.\\
 \verb|\dbSQLJDBC{jdbcConnectionURL}{user}{password}{commaSeperatedListOfTableNames}| \\
@@
 \textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.\\[4pt]
-\textbf{WARNING:} \verb|\dbSQLText| and \verb|\dbSQLFile| are currently horribly slow, even for medium sized (more than 100 tuples) amounts of data. Use \verb|\dbSQLJDBC| instead.
+
+\subsection{CSV}
+\verb|\dbCSVFile[options]{fileWithCSVData}{tableName}| \\[8pt]
+\textbf{Note:} Relative filepaths are relative to the tex file.\\[4pt]
+\noindent Options:\\[3pt]
+%
+\option
+{headers}
+{Defines the names of columns for the CSV file. If not set the first line of the CSV file is used as header.}
+{Javascript-Array of Strings}
+{\texttt{[]}}
+
+\option
+{quote}
+{Defines the quote character for this CSV file.}
+{String}
+{\texttt{"}}
+
+\option
+{seperator}
+{Defines the seperator character for this CSV file.}
+{String}
+{\texttt{,}}
 

### 2015-09-14 src/pdbf/latex/LaTeX_Compiler.java

@@ -14,6 +14,9 @@ import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.StringTokenizer;
 
+import org.apache.commons.csv.CSVFormat;
+import org.apache.commons.csv.CSVParser;
+import org.apache.commons.csv.CSVRecord;
 import org.apache.commons.io.FileUtils;
 
 import com.google.gson.Gson;
@@ -340,6 +343,7 @@ public class LaTeX_Compiler {
 	try {
 	    Database db = (Database) overlay.type;
 	    File f = new File(baseDir + "pdbf-db.sql");
+	    Alasql alasql = dbc.alasql;
 	    switch (db.type) {
 	    case 1:
 		FileUtils.writeStringToFile(f, db.value1 + System.lineSeparator(), Tools.utf8, true);
@@ -359,10 +363,15 @@ public class LaTeX_Compiler {
 		Connection conn = DriverManager.getConnection(db.value1, db.value2, db.value3);
 		// Create JSON Database
 		StringTokenizer stok = new StringTokenizer(db.value4, ",");
-		Alasql alasql = dbc.alasql;
 		while (stok.hasMoreTokens()) {
 		    Table table = new Table();
 		    String curTable = stok.nextToken().trim();
+		    
+			if (alasql.containsTable(curTable)) {
+			    System.err.println("Error: Database already contains a table with name \"" + curTable + "\"");
+			    System.exit(-1);
+			}
+		    
 		    Statement stmt = conn.createStatement();
 		    ResultSet rs = stmt.executeQuery("SELECT * FROM " + curTable);
 		    ResultSetMetaData rsmd = rs.getMetaData();
@@ -383,6 +392,42 @@ public class LaTeX_Compiler {
 		    alasql.addTable(table, curTable);
 		}
 		conn.close();
+		break;
+	    case 4:
+		File csvData = new File(db.value1);
+		if (!csvData.isAbsolute()) {
+		    csvData = new File(latexFolder + File.separator + db.value1);
+		}
+		CSVFormat csvformat = CSVFormat.RFC4180.withDelimiter(db.seperator).withQuote(db.quote);
+		if (db.headers.length > 0) {
+		    csvformat = csvformat.withHeader(db.headers);
+		} else {
+		    csvformat = csvformat.withHeader();
+		}
+		CSVParser parser = CSVParser.parse(csvData, Tools.utf8, csvformat);
+		
+		Table table = new Table();
+		String curTable = db.value2;
[…]

### 2015-09-26 minimal.tex
changed some filenames
csv as data source Version 1.0.4
corrected line endings
- All PDBF elements use the font size, font family, and font style that was active at the point of their definition. - Lots of Bugfixes

@@
 \documentclass[10pt]{article}
-\usepackage{PDBF}
+\usepackage{pdbf}
 \usepackage[a4paper]{geometry}
@@
 %\dbSQLJDBC{jdbc:postgresql://localhost:5432/postgres}{postgres}{test}{testtable}%
+%\dbCSVFile{result.csv}{test}
 
@@
 \chart[width=\textwidth, height=0.45\textwidth, chartType=bar]{SELECT * FROM test;}\\[3pt]
+\newpage
+{\fontfamily{phv} \large \selectfont
+\noindent
+Change the font before this line to Helvetica and fontsize from 10 to 12.\\[3pt]
 \chart[width=\textwidth, height=0.45\textwidth, chartType=signatureplot]{SELECT b FROM test;}\\[3pt]
-\multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}\\[3pt]
+\multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}\\[3pt]
 \pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}\\[3pt]
 \dataTable[verticalLines=i, horizontalLinesHeader=b hh, horizontalLinesBody=i]{SELECT * FROM test;}\\[3pt]
+}
 \end{document}

### 2015-09-26 pdbf-doc.tex
corrected line endings
- All PDBF elements use the font size, font family, and font style that was active at the point of their definition. - Lots of Bugfixes

@@
 
-\def\pdbfQuality{4.0}
+\def\pdbfQuality{2.0}
 
@@
 \subsection{CSV}
+\textbf{Note:} By default the first line of the CSV file is used as attribute names for the table. You can use the header option to manually specify attribute names.\\[3pt]
 \verb|\dbCSVFile[options]{fileWithCSVData}{tableName}| \\[8pt]
@@
 \section{PDBF elements}
+\textbf{Note:} All PDBF elements use the font size, font family, and font style that was active at the point of their definition.
+You can for example surround all commands with \verb|\textbf{...}|, \verb|\emph{...}| or any other font command.\\[3pt]
 \subsection{Chart}
@@
     \end{minipage} \\[8pt]
-
     \begin{minipage}{1.0\textwidth}
@@
     \end{minipage}
-%    \begin{minipage}{.48\textwidth}
-%    \chart[width=\textwidth, height=0.8\textwidth, chartType=signatureplot]{SELECT * FROM test;}
-%    \caption{Signatureplot Chart}
-%    \end{minipage}
-%    \hspace{11pt}
-%    \begin{minipage}{.48\textwidth}
-%    \chart[width=\textwidth, height=0.8\textwidth, chartType=Line, xunit=Date, yunit={Runtime [in sec]}, options={"strokeWidth": 0, "drawPoints": true, "pointSize": 3, "highlightCircleSize": 4, "xRangePad": 10}]{SELECT date, runtimeA AS [engine A], runtimeB AS [engine B] FROM data2;}
-%    \caption{Multi-column Dot Chart}
-%    \end{minipage}
 \end{figure}
@@
 {quality}
-{Sets the quality for the image version of the chart in the pdf. 1.0 corresponds roughly to 120 pixels per inch. Can also be redefined globally (\texttt{pdbfQuality}).}
-{Number $>$ 0}
-{\texttt{1.0}}
-
-\option
-{scale}
-{Scales the size of the chart. 1.0 corresponds roughly to font-size that is currently choosen in \LaTeX. Can also be redefined globally (\texttt{pdbfScale}).}
+{Sets the quality for the image version of the chart in the pdf. 1.0 corresponds roughly to 240 pixels per inch. Can also be redefined globally (\texttt{pdbfQuality}).}
 {Number $>$ 0}
@@
 
-%\option %TODO: not yet implemented
-%{logscale}
-%{If set, the y axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
-%{\texttt{true} or \texttt{false}}
-%{\texttt{false}}
-%
+\option
+{logscale}
+{If set, the y-axis uses log scale. Can also be redefined globally (\texttt{pdbfLogscale}).}
+{\texttt{true} or \texttt{false}}
+{\texttt{false}}
+
 %\option %TODO: not yet implemented
@@
 {width}
-{Sets the width of the chart.}
+{Sets the width of the table.}
 {\LaTeX~length}
@@
 {height}
-{Sets the height of the chart.}
+{Sets the height of the table.}
 {\LaTeX~length}
@@
 {quality}
-{Sets the quality for the image version of the chart in the pdf. 1.0 corresponds roughly to 120 pixels per inch. Can also be redefined globally (\texttt{pdbfQuality}).}
-{Number $>$ 0}
-{\texttt{1.0}}
-
-\option
-{scale}
-{Scales the size of the chart. 1.0 corresponds roughly to font-size that is currently choosen in \LaTeX. Can also be redefined globally (\texttt{pdbfScale}).}
+{Sets the quality for the image version of the table in the pdf. 1.0 corresponds roughly to 240 pixels per inch. Can also be redefined globally (\texttt{pdbfQuality}).}
 {Number $>$ 0}
@@
 {cols}
-{The attributes for the columns in the pivot table.}
+{The attributes for the columns in the pivot table. \textbf{Note:} This feature is currently broken!}
 {JavaScript array of strings (e.g. ["a", "b", "c"]). The name of attributes have to be present in the result of the sql query for the overlay and page}
@@
 \subsection{DataText}
-Write text that is the result of a SQL-statement.\\[3pt]
+Write text that is the result of a SQL-statement.\\
+\textbf{Note:} DataText can be linked to other PDBF elements (Currently only charts supported).
+This means that if you move the mouse over this DataText then the specified data in the linked PDBF element is highlighted.\\[3pt]
 Example:\\[3pt]

### 2015-09-26 data/base64.js

@@ -1,19 +1,33 @@
 if (!Function.prototype.bind) {
-	Function.prototype.bind = function(oThis) {
-		if (typeof this !== 'function') {
-			throw new TypeError('Function.prototype.bind - what is trying to be bound is not callable');
-		}
+  Function.prototype.bind = function(oThis) {
+    if (typeof this !== 'function') {
+      // closest thing possible to the ECMAScript 5
+      // internal IsCallable function
+      throw new TypeError('Function.prototype.bind - what is trying to be bound is not callable');
+    }
 
-		var aArgs = Array.prototype.slice.call(arguments, 1), fToBind = this, fNOP = function() {
-		}, fBound = function() {
-			return fToBind.apply(this instanceof fNOP ? this : oThis, aArgs.concat(Array.prototype.slice.call(arguments)));
-		};
+    var aArgs   = Array.prototype.slice.call(arguments, 1),
+        fToBind = this,
+        fNOP    = function() {},
+        fBound  = function() {
+          return fToBind.apply(this instanceof fNOP
+                 ? this
+                 : oThis,
+                 aArgs.concat(Array.prototype.slice.call(arguments)));
+        };
 
-		fNOP.prototype = this.prototype;
-		fBound.prototype = new fNOP();
+    fNOP.prototype = this.prototype;
+    fBound.prototype = new fNOP();
 
-		return fBound;
-	};
+    return fBound;
+  };
+}
+
+if (!String.prototype.startsWith) {
+  String.prototype.startsWith = function(searchString, position) {
+    position = position || 0;
+    return this.indexOf(searchString, position) === position;
+  };
 }
 
 /* Array of bytes to base64 string decoding */

### 2015-09-26 data/c3-dev.js

@@ -258,7 +258,6 @@
         
         //////////////////////////////////////////////////////////////////////
         config.point_r = config.point_r * config.completeScale;
-        $$.svg[0][0].style.fontSize = config.completeScale*10;
         $$.svg[0][0].style.strokeWidth = config.completeScale;
         config.padding_right = config.padding_right*config.completeScale;
         config.padding_top = config.padding_top*config.completeScale;
@@ -381,7 +380,7 @@
 
     c3_chart_internal_fn.updateSizes = function () {
         var $$ = this, config = $$.config;
-        var legendHeight = $$.legend ? $$.getLegendHeight() : 0,
+        var legendHeight = $$.legend ? ($$.getLegendHeight() + 5*config.completeScale) : 0,
             legendWidth = $$.legend ? $$.getLegendWidth() : 0,
             legendHeightForBottom = $$.isLegendRight || $$.isLegendInset ? 0 : legendHeight,
             hasArc = $$.hasArcType(),
@@ -1046,7 +1045,7 @@
             size_width: undefined,
             size_height: undefined,
             padding_left: undefined,
-            padding_right: 10,
+            padding_right: 5,
             padding_top: 5,
             padding_bottom: undefined,
             innerTickSize: 3,
@@ -1471,7 +1470,8 @@
         }
 
         domainLength = Math.abs(yDomainMax - yDomainMin);
-        padding = padding_top = padding_bottom = domainLength * 0.1;
+        padding = padding_top = padding_bottom = domainLength * 0.03;
+        padding_bottom = 0;
 
         if (typeof center !== 'undefined') {
             yDomainAbs = Math.max(Math.abs(yDomainMin), Math.abs(yDomainMax));
@@ -1528,7 +1528,7 @@
             maxDataCount = $$.getMaxDataCount();
             padding = maxDataCount > 1 ? (diff / (maxDataCount - 1)) / 2 : 0.5;
         } else {
-            padding = diff * 0.01;
+            padding = diff * 0.03;
         }
         if (typeof config.axis_x_padding === 'object' && notEmpty(config.axis_x_padding)) {
             paddingLeft = isValue(config.axis_x_padding.left) ? config.axis_x_padding.left : padding;
@@ -1537,6 +1537,11 @@
             paddingLeft = paddingRight = config.axis_x_padding;
         } else {
             paddingLeft = paddingRight = padding;
+            if (!$$.hasType('bar')) {
+            	paddingLeft = 0.00000001;
+            } else {
[…]

### 2015-09-26 data/c3.js

@@ -16,21 +16,21 @@ this.y2Orient=b.axis_rotated?b.axis_y2_inner?"bottom":"top":b.axis_y2_inner?"lef
 y:0,y2:0};this.rotated_padding_left=30;this.rotated_padding_right=b.axis_rotated&&!b.axis_x_show?0:30;this.rotated_padding_top=5;this.withoutFadeIn={};this.intervalForObserveInserted=void 0;this.axes.subx=a.selectAll([])};d.initChartElements=function(){this.initBar&&this.initBar();this.initLine&&this.initLine();this.initArc&&this.initArc();this.initGauge&&this.initGauge();this.initText&&this.initText()};d.initWithData=function(a){var b=this,c=b.d3,e=b.config,k=!0;b.axis=new q(b);b.bound=e.bindto;
 b.initPie&&b.initPie();b.initBrush&&b.initBrush();b.initZoom&&b.initZoom();b.selectChart=e.bindto?"function"===typeof e.bindto.node?e.bindto:c.select(e.bindto):c.selectAll([]);b.selectChart.empty()&&(b.selectChart=c.select(document.createElement("div")).style("opacity",0),b.observeInserted(b.selectChart),k=!1);b.selectChart.html("").classed("c3",!0);b.data.xs={};b.data.targets=b.convertDataToTargets(a);e.data_filter&&(b.data.targets=b.data.targets.filter(e.data_filter));e.data_hide&&b.addHiddenTargetIds(!0===
 e.data_hide?b.mapToIds(b.data.targets):e.data_hide);e.legend_hide&&b.addHiddenLegendIds(!0===e.legend_hide?b.mapToIds(b.data.targets):e.legend_hide);b.hasType("gauge")&&(e.legend_show=!1);b.updateSizes();b.updateScales();var g=b.getXDomain(b.data.targets),d=b.getYDomain(b.data.targets,"y");n.xd=function(){return g};n.yd=function(){return d};n.pl=function(){return b.margin.left};n.pb=function(){return b.margin.bottom};b.x.domain(c.extent(g));b.y.domain(d);b.y2.domain(b.getYDomain(b.data.targets,"y2"));
[…]

### 2015-09-26 data/d3.js

@@ -7505,7 +7505,7 @@
     return d3_scale_nice(domain, d3_scale_niceStep(d3_scale_linearTickRange(domain, m)[2]));
   }
   function d3_scale_linearTickRange(domain, m) {
-    if (m == null) m = 8;
+    if (m == null) m = 7.5;
     var extent = d3_scaleExtent(domain), span = extent[1] - extent[0], step = Math.pow(10, Math.floor(Math.log(span / m) / Math.LN10)), err = m / span * step;
     if (err <= .15) step *= 10; else if (err <= .35) step *= 5; else if (err <= .75) step *= 2;
     extent[0] = Math.ceil(extent[0] / step) * step;

### 2015-09-26 data/main.js

@@ -59,6 +59,11 @@ function display(json, page, phantomJS) {
 	} else {
 		style = "position: absolute; width:" + (json.type.I.x2 - json.type.I.x1 + 0.001) * 100 + "%; height:" + (json.type.I.y1 - json.type.I.y2 + 0.001) * 100 + "%; left:" + json.type.I.x1 * 100 + "%; bottom:" + (json.type.I.y2 - 0.001) * 100 + "%;";
 	}
+	if (typeof font_store === "undefined" || typeof font_store[json.name] === "undefined") {
+		style += "font-size: " + (zoomFactor * 12.0) + "pt;"
+	} else {
+		style += "font-size: " + (zoomFactor * 12.0) + "pt; font-family: " + font_store[json.name] + ", sans-serif;"
+	}
 	page.appendChild(container);
 	
 	var containerOver = document.getElementById(json.name + "Big");
@@ -266,6 +271,7 @@ function buildContainerChartBig(json, containerOver, initial) {
 			containerOptions.style.visibility = 'visible';
 		}
 		var optionsBig = getChartOptions(json.jsonBig, rawZoomFactor, json.chartdataBig.values, containerContent);
+		optionsBig.completeScale = 1.25;
 		json.chart = c3.generate(optionsBig);
 	};
 	var update = function() {
@@ -291,6 +297,7 @@ function buildContainerChartBig(json, containerOver, initial) {
 		json.jsonBig.type.I.showRangeSelector = showRangeSelector.checked;
 		
 		var optionsBig = getChartOptions(json.jsonBig, rawZoomFactor, json.chartdataBig.values, containerContent);
+		optionsBig.completeScale = 1.25;
 		
 		// logscale
 		if (json.jsonBig.type.I.logScale == true) {
@@ -338,9 +345,9 @@ function buildContainerChartBig(json, containerOver, initial) {
 	
 	var viewerContainer = document.getElementById("viewerContainer");
 	var tip = "Tip: Scroll to zoom the graph. Click and drag to pan the graph.<br/>";
-	var ref = prepopulateContainerOver(containerOver, viewerContainer, tip, [ json ], updateData, 'graph', true);
 	json.jsonBig = jQuery.extend(true, {}, json);
-	var chartdataCpy = getChartData(json);
+	var ref = prepopulateContainerOver(containerOver, viewerContainer, tip, [ json ], updateData, 'graph', true);
+	var chartdataCpy = getChartData(json.jsonBig);
 	json.resultBig = chartdataCpy.res;
[…]

### 2015-09-26 data/pdf.worker.js

@@ -443,7 +443,7 @@ stateManager||new StateManager(new TextState);var textContent={items:[],styles:O
 {fontFamily:font.fallbackName,ascent:font.ascent,descent:font.descent,vertical:font.vertical};return{str:[],dir:null,width:0,height:0,transform:null,fontName:font.loadedName}}function runBidi(textChunk){var str=textChunk.str.join("");var bidiResult=PDFJS.bidi(str,-1,textState.font.vertical);textChunk.str=bidiResult.str;textChunk.dir=bidiResult.dir;return textChunk}function handleSetFont(fontName,fontRef){return self.loadFont(fontName,fontRef,xref,resources).then(function(translated){textState.font=
 translated.font;textState.fontMatrix=translated.font.fontMatrix||FONT_IDENTITY_MATRIX})}function buildTextGeometry(chars,textChunk){var font=textState.font;textChunk=textChunk||newTextChunk();if(!textChunk.transform){var tsm=[textState.fontSize*textState.textHScale,0,0,textState.fontSize,0,textState.textRise];var trm=textChunk.transform=Util.transform(textState.ctm,Util.transform(textState.textMatrix,tsm));if(!font.vertical)textChunk.height=Math.sqrt(trm[2]*trm[2]+trm[3]*trm[3]);else textChunk.width=
 Math.sqrt(trm[0]*trm[0]+trm[1]*trm[1])}var width=0;var height=0;var glyphs=font.charsToGlyphs(chars);var defaultVMetrics=font.defaultVMetrics;for(var i=0;i<glyphs.length;i++){var glyph=glyphs[i];if(!glyph){width+=textState.wordSpacing*textState.textHScale;continue}var vMetricX=null;var vMetricY=null;var glyphWidth=null;if(font.vertical)if(glyph.vmetric){glyphWidth=glyph.vmetric[0];vMetricX=glyph.vmetric[1];vMetricY=glyph.vmetric[2]}else{glyphWidth=glyph.width;vMetricX=glyph.width*.5;vMetricY=defaultVMetrics[2]}else glyphWidth=
[…]

### 2015-09-26 data/pivot.js

@@ -278,7 +278,7 @@
 							if (data.intable) {
 								return formatter(x[attr]);
 							} else {
-								var ret = "<span class=\"hasTooltip\">" + formatter(x[attr]) + "</span><span style=\"position:absolute;\" ><table style=\"width: 100%;\"><tr>";
+								var ret = "<span class=\"hasTooltip\">" + formatter(x[attr]) + "</span><div style=\"position:absolute;\" ><table style=\"width: 100%;\"><tr>";
 								for (key in x) {
 									ret += "<th class=\"pvtAxisLabel\">" + key + "</th>";
 								}
@@ -389,7 +389,7 @@
 									var ret = "";
 									r += (x[attr]);
 								} else {
-									var ret = "<span class=\"hasTooltip\">" + (x[attr]) + "</span><span style=\"position:absolute;\" ><table style=\"width: 100%; border: 0.2em solid #CDCDCD; background: white;\"><tr>";
+									var ret = "<span class=\"hasTooltip\">" + (x[attr]) + "</span><div style=\"position:absolute;\" ><table style=\"width: 100%; border: 0.2em solid #CDCDCD; background: white;\"><tr>";
 									for (key in x) {
 										ret += "<th class=\"pvtAxisLabel\">" + key + "</th>";
 									}
@@ -1679,11 +1679,11 @@
 						// ////////////////////////////////////////////////////////////////////////////////////////
 						if (!subopts.intable) {
 							$('.hasTooltip').hover(function() {
-								var tooltip = $(this).next('span')[0];
+								var tooltip = $(this).next('div')[0];
 								tooltip.style.visibility = 'visible';
 								tooltip.style.opacity = 1;
 							}, function() {
-								var tooltip = $(this).next('span')[0];
+								var tooltip = $(this).next('div')[0];
 								setTimeout(function(o) {
 									o.style.visibility = 'hidden';
 								}, 500, tooltip);

### 2015-09-26 external-tools/captureData.js

@@ -15,6 +15,10 @@ page.onError = function (msg, trace) {
 	phantom.exit();
 };
 
+/*page.onConsoleMessage = function(msg, lineNum, sourceId) {
+console.log('CONSOLE: ' + msg + ' (from line #' + lineNum + ' in "' + sourceId + '")');
+};*/
+
 fs.changeWorkingDirectory(system.args[2]);
 
 page.open(system.args[1], function() {
@@ -34,7 +38,8 @@ page.open(system.args[1], function() {
 	});
 		
 	window.setTimeout(function () {
-		fs.write(data.name + '.data', data.result, 'w');
+		console.log("Finished " + data.name);
+		fs.write('../' + data.name + '.data', data.result, 'w');
 		phantom.exit();
 	}, 500);
 });

### 2015-09-26 external-tools/captureDatabase.js

@@ -33,7 +33,7 @@ page.open(system.args[1], function() {
 	});
 		
 	window.setTimeout(function () {
-		fs.write('pdbf-db.json', data, 'w');
+		fs.write('../' + 'pdbf-db.json', data, 'w');
 		phantom.exit();
 	}, 500);
 });

### 2015-10-07 pdbf-doc.tex
- implemented basic test case that compiles test documents - small clarifications and bugfixes in pdbf-doc - reduced average number of ticks for charts

@@
 
+\newpage
+
 \section{Getting started}
@@
 \item For further information take a look at the \href{http://uds-datalab.github.io/PDBF/}{documentation}
+\end{itemize}
+
+\section{Features}
+\subsection{Automatic generation of Charts, Multiplot Charts, Pivot tables}
+With PDBF you don't need to manually generate these kinds of elements.
+The PDBF compiler automatically generates a static version for the pdf and the dynamic version for the html part of the PDBF document.
+This also means that your document is always up to date!
+If you change something in the underlying data that generate your PDBF document and then recompile the document, then the data in the document is up to date.
+No need to manually update externally generated charts or pivot tables.
+
+\subsection{Generate your document directly from the results of your experiment}
+The idea of PDBF is to store the results of the experiment directly in the document and to make it more transparent how this chart, pivot table, etc.
+was generated from the result data.
+Therefore we currently support CSV files, SQL files, and SQL servers as data sources and use SQL as description language for the transformation of the raw result data to the final representation in the document.
+
+\subsection{Compile LaTeX to single HTML file}
+You can also use the PDBF compiler to compile your LaTeX files to a single HTML file.
+To do so just run the compiler as on any other document (you dont need to include the pdbf package in your tex file):\\[3pt]
+\texttt{java -jar pdbf.jar sometexfile.tex}\\[3pt]
+The resulting HTML file is saved in the same folder with the same name but html ending.
+
+\section{Build Instructions}
+\begin{itemize}
+\item Run "mvn package" if you only want to compile pdbf.jar
+\item Run "mvn verify" if you want to compile pdbf.jar and run integration tests.
 \end{itemize}
@@
 
+\newpage
+
 \section{How to use the PDBF compiler}
 Right now there are only two modes: \\[4pt]
--- The standart mode which compiles a tex file into the hybrid PDBF format:\\
+-- The standart mode which compiles a tex file into the janiform PDBF format:\\
 \verb|java -jar PDBF.jar LaTeX_file| \\[2pt]
@@
 \section{Requirements}
-These \LaTeX\relax packages are required for PDBF documents:
+The following \LaTeX~packages are required for PDBF documents:
 \begin{itemize}
@@
 \end{itemize}
-\noindent Furthermore it is needed that your \LaTeX\relax document uses the geometry package to specify the page size.\\
+\noindent Furthermore it is needed that your \LaTeX~document uses the geometry package to specify the page size.\\
 
-\noindent \textbf{Warning:} There are problems with \textbf{inputenc UTF8} and \textbf{pgfpages}.
-If you really have to use the inputenc UTF8 package then you cant use non-ASCII characters inside queries.
-The pgfpages package is incompatible to this package (Its fixable but requires additional effort).
+\noindent \textbf{Warning:} There are problems with \textbf{inputenc} and \textbf{pgfpages} package.
+If you really have to use the inputenc package with UTF8 option then you cant use non-ASCII characters inside queries.
+The pgfpages package is currently incompatible to this package.
 
-\section{SQL queries for PDBF elements}
+\section{SQL specifics and special functions}
 PDBF uses AlaSQL as database engine.
 For all available features of the database visit \url{https://github.com/agershun/alasql}.
-Some functionality might not yet be available because we currently use not the latest version.\\[4pt]
+Some functionality might not yet be available because we currently don't use the latest version.\\[4pt]
 \textbf{Note:} Mostly sql syntax is standart sql, but for example attributes with spaces in their name have to be sourrounded by square brackets.
@@
 \noindent We extended AlaSQL with the following statistical sql functions:\\[5pt]
-\verb|GRUBBS_FILTER(arr, alpha)|: \\
-\indent Desc.: Filters an array of values with a \href{http://en.wikipedia.org/wiki/Grubbs'_test_for_outliers}{grubbs test}\\
+\verb|GRUBBS_FILTER(arr, alpha, max)|: \\
+\indent Desc.: Filters an array of values with a two-sided \href{http://en.wikipedia.org/wiki/Grubbs'_test_for_outliers}{grubbs test}.
+The grubbs test is performed on the whole array.
+If a significant outlier is detected (Z > ZCrit) and the relative margin of error is not too high (margin\_of\_error / avg > 0.025) then the outlier is removed.
+This procedure is repeated until one of these two conditions fails.\\
 \indent Returns the filtered array.\\[3pt]
@@
 Clickable SQL-statements\\[3pt]
+\textbf{Note:} Currently SQL is limited to one line of text. No line breaks are possible.\\
 Example:\\[3pt]
@@
 Write text that is the result of a SQL-statement.\\
+\textbf{Note:} Currently DataText is limited to one line of text. No line breaks are possible.\\
 \textbf{Note:} DataText can be linked to other PDBF elements (Currently only charts supported).

### 2015-10-26 pdbf-doc.tex
small improvements in documentation, prepared new test to visually verify the integrity of procuced reference pdfs
part 1 of pdf and tar integration tests
pdf and tar integration tests finished. closing #2

integration tests now delete log and aux files after test

@@
 \item Adjust config.cfg
-\item Try to compile pdbf-doc.tex file with this command: java -jar pdbf.jar pdbf-doc.tex
+\item Try to compile pdbf-doc.tex file with this command: \texttt{java -jar pdbf.jar pdbf-doc.tex}
 \item Open pdbf-doc.html, this is the final output of the compilation process, if you rename it to ".pdf" it is also a valid pdf-document
-\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: java -jar pdbf.jar --vm pdbf-doc.html vldb-Invaders.ova
+\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --vm pdbf-doc.html vldb-Invaders.ova}.
+\item Open pdbf-doc.ova, this is the final output with the attached ova file. Its still a valid pdf and html at the same time.
 \item You can play around with the minimal.tex file which contains a minimal example of PDBF usage
@@
 \texttt{java -jar pdbf.jar sometexfile.tex}\\[3pt]
-The resulting HTML file is saved in the same folder with the same name but html ending.
+The resulting HTML file is saved in the same folder with the same name but with ".html" filename extension.
 
@@
 phantomJS (\url{https://github.com/ariya/phantomjs})\\
-Apache Commons IO (\url{http://commons.apache.org/proper/commons-io/})\\
-Apache Commons Codec (\url{http://commons.apache.org/proper/commons-codec/})\\
+Apache Commons (\url{http://commons.apache.org/})\\
+Apache PDFBox (\url{https://pdfbox.apache.org/})\\
+jTar (\url{https://github.com/kamranzafar/jtar})\\
 google-gson (\url{https://github.com/google/gson})\\
@@
 \verb|java -jar PDBF.jar LaTeX_file| \\[2pt]
+\textbf{Note:} The compiled document can be found in the same folder as the tex file and has the same name as the tex file but ".html" filename extension.\\[8pt]
 -- The VM mode which attaches a VirtualBox .ova file to an existing PDBF:\\
-\verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|
+\verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|\\[2pt]
+\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".ova" filename extension.
 
@@
 %
+\newpage
 \section{Data sources}
@@
 
+\newpage
 \section{PDBF elements}
@@
 
-\newpage
 \noindent Macros: \\[3pt]

### 2015-10-28 minimal.tex
- fixed superfluous space for charts, pivot tables, and multiplot charts - added missing parameters for charts in pdbf.sty - prevent sql injection through table names - added second dataset to minimal example - increased default quality setting - tooltip of charts is now always the same size regardless of sourrounding font size
- fixed wrong font and scale handling for charts and multiplot charts

Version 1.1.1

@@
 
-\dbSQLText{CREATE TABLE test(a int, b int); INSERT INTO test VALUES(1, 10); INSERT INTO test VALUES(2, 100); INSERT INTO test VALUES(3, 120); INSERT INTO test VALUES(4, 100);}
+\dbSQLText{CREATE TABLE test(a int, b int, c int); INSERT INTO test VALUES(1, 10, 15); INSERT INTO test VALUES(2, 100, 90); INSERT INTO test VALUES(3, 120, 110); INSERT INTO test VALUES(4, 100, 150);}
 %\dbSQLFile{somedb.sql}
@@
 \newpage
-{\fontfamily{phv} \large \selectfont
+{\fontfamily{phv} \tiny \selectfont
 \noindent
-Change the font before this line to Helvetica and fontsize from 10 to 12.\\[3pt]
+Change the font before this line to Helvetica and fontsize to tiny.\\[3pt]
 \chart[width=\textwidth, height=0.45\textwidth, chartType=signatureplot]{SELECT b FROM test;}\\[3pt]
-\multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}\\[3pt]
+\multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, topArr={["a smaller or equal 2", "a greater 2"]}, yFirst]{SELECT a,?
+FROM test WHERE ?;}\\[3pt]
 \pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}\\[3pt]

### 2015-10-28 pdbf-doc.tex
- fixed superfluous space for charts, pivot tables, and multiplot charts - added missing parameters for charts in pdbf.sty - prevent sql injection through table names - added second dataset to minimal example - increased default quality setting - tooltip of charts is now always the same size regardless of sourrounding font size
- fixed wrong font and scale handling for charts and multiplot charts

Version 1.1.1

@@
     Example:\\
-    \multiplotChart[width=\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]}, yFirst]{SELECT a,? FROM test WHERE ?;}
+    \multiplotChart[width=\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a > 2", "a <= 2"]},  topArr={["a smaller or equal 2", "a greater 2"]}, yFirst]{SELECT a,?
+FROM test WHERE ?;}
     \caption{Multiplot Line Chart}

### 2015-10-28 src/pdbf/html/CompleteRun_HTML.java

@@ -17,6 +17,10 @@ public class CompleteRun_HTML {
 	String baseDir = Tools.getBaseDir();
 	
 	LaTeX_Compiler.suffix = Tools.getOS();
+	
+	if (!includeRes) {
+	    System.out.println("!!!!Warning!!!! includeRes is off");
+	}
 
 	if (args.length > 3 || args.length < 1 || args[0].equalsIgnoreCase("--help")) {
 	    System.out.println("Usage:\tjava -jar pdbf.jar LaTeX_file\nOR\n\tjava -jar pdbf.jar --vm  PDBF_File.html VM_File.ova\nFor further help visit: https://github.com/uds-datalab/PDBF");

### 2015-11-04 src/pdbf/common/Tools.java
pdbf compiler now correctly supports compressed xref tables (PDF version >= 1.5)
replaced phantom-js for mac with a working version

Version 1.1.2

@@ -1,12 +1,21 @@
 package pdbf.common;
 
+import java.io.ByteArrayInputStream;
+import java.io.ByteArrayOutputStream;
 import java.io.File;
 import java.io.IOException;
 import java.io.UnsupportedEncodingException;
 import java.net.URISyntaxException;
+import java.nio.ByteBuffer;
 import java.nio.charset.Charset;
+import java.nio.charset.StandardCharsets;
 import java.security.CodeSource;
+import java.text.DecimalFormat;
 import java.util.ArrayList;
+import java.util.StringTokenizer;
+import java.util.zip.Deflater;
+import java.util.zip.DeflaterOutputStream;
+import java.util.zip.InflaterOutputStream;
 
 import org.apache.commons.codec.binary.Base64;
 import org.apache.commons.io.FileUtils;
@@ -15,19 +24,34 @@ import pdbf.html.CompleteRun_HTML;
 
 public class Tools {
 
+    public static DecimalFormat df = new DecimalFormat("0000000000");
+
     public static Charset utf8 = Charset.forName("UTF-8");
 
     public static String baseDir = new File(getBaseDir()).getParent() + File.separator;
     public static String refDir = baseDir + "src" + File.separator + "pdbf" + File.separator + "referenceImages" + File.separator;
     public static String testDir = baseDir + "src" + File.separator + "pdbf" + File.separator + "tests" + File.separator;
     public static String suffix = getOS();
-    
+
     public static ArrayList<File> deleteList = new ArrayList<File>();
     public static ArrayList<Process> processes = new ArrayList<Process>();
-    
+
+    public static int skipNewline(StringBuilder sb, int i, boolean forward) {
+	if (i == -1) {
+	    return -1;
+	}
+	if (!forward) {
+	    --i;
+	}
+	while (sb.charAt(i) == '\n' || sb.charAt(i) == '\r') {
+	    i = forward ? i + 1 : i - 1;
+	}
+	return forward ? i : i + 1;
+    }
+
     public static String getOS() {
 	String OS = System.getProperty("os.name").toLowerCase();
-	
+
 	if (OS.contains("win")) {
 	    return "win";
 	} else if (OS.contains("mac")) {
@@ -40,7 +64,7 @@ public class Tools {
 	    return null;
 	}
     }
-    
+
[…]

### 2015-11-04 src/pdbf/vm/VM_Compiler.java

@@ -1,9 +1,12 @@
 package pdbf.vm;
+
 import java.io.File;
 import java.nio.charset.StandardCharsets;
 import java.text.DecimalFormat;
 
 import org.apache.commons.io.FileUtils;
+
+import pdbf.common.Tools;
 
 public class VM_Compiler {
 
@@ -18,7 +21,7 @@ public class VM_Compiler {
 	    System.exit(-1);
 	}
 	String basename = args[1].substring(0, args[1].length() - 5);
-	
+
 	try {
 	    String vmcontent = FileUtils.readFileToString(new File(args[2]), StandardCharsets.ISO_8859_1);
 	    if (vmcontent.toLowerCase().contains("</script>")) {
@@ -28,8 +31,8 @@ public class VM_Compiler {
 	    StringBuilder sb = new StringBuilder(vmcontent);
 	    String replace = "%PDF-1.5\n%ª«¬­.ovf\0\n1 0 obj\nstream\n<head><meta charset=UTF-8><script>";
 	    sb.replace(0, replace.length(), replace);
-	    
-	    //Fix tar
+
+	    // Fix tar
 	    int checksum = 0;
 	    sb.setCharAt(148, ' ');
 	    sb.setCharAt(149, ' ');
@@ -42,73 +45,52 @@ public class VM_Compiler {
 	    for (int i = 0; i < 512; ++i) {
 		checksum += sb.charAt(i);
 	    }
-	    //six digit octal number with leading zeroes
+	    // six digit octal number with leading zeroes
 	    String chk = Integer.toOctalString(checksum);
 	    sb.replace(148, 155, "0000000");
-	    sb.replace(155-chk.length(), 155, chk);
+	    sb.replace(155 - chk.length(), 155, chk);
 	    sb.setCharAt(155, '\0');
-	    
-	    //Add new file to tar
+
+	    // Add new file to tar
 	    String name = "DO_NOT_DELETE\0";
 	    String n = name + sb.substring(name.length(), 513);
 	    int von = sb.length();
 	    sb.append(n);
-	    
+
 	    String removeFromHTML = "%PDF-1.5\n%ª«¬­<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>" + "<head><meta charset=\"utf-8\"><!--\n1337 0 obj\nstream";
 	    String addToHTML = "</script>";
[…]

### 2015-11-30 minimal.tex
turn off automatic line ending conversion of git
- added new test for charts - reverted pivot.js to stable version - updated reference pictures

@@
 
-\dbSQLText{CREATE TABLE test(a int, b int, c int); INSERT INTO test VALUES(1, 10, 15); INSERT INTO test VALUES(2, 100, 90); INSERT INTO test VALUES(3, 120, 110); INSERT INTO test VALUES(4, 100, 150);}
+\dbSQLText{CREATE TABLE test(a int, b int, c int); INSERT INTO test VALUES(1, 1, 1); INSERT INTO test VALUES(2, 2, 4); INSERT INTO test VALUES(3, 3, 8); INSERT INTO test VALUES(4, 4, 16);}
 %\dbSQLFile{somedb.sql}
@@
 The table ``test'' contains \dataText{SELECT COUNT(*) FROM test;} tuples.\\[3pt]
-The following points are linked to the chart below: \dataText[linkTo=testname, linkSelector=2, linkLabel=Value at 2]{SELECT * FROM test WHERE a=2;}, \dataText[linkTo=testname, linkSelector=3, linkLabel=Value at 3]{SELECT * FROM test WHERE a=3;}.\\[3pt]
-\chart[width=\textwidth, height=0.45\textwidth, chartType=line, name=testname]{SELECT * FROM test;}\\[3pt]
-\chart[width=\textwidth, height=0.45\textwidth, chartType=bar]{SELECT * FROM test;}\\[3pt]
-\newpage
+The following points are linked to the chart on the left side below: \dataText[linkTo=testname, linkSelector=2, linkLabel=Value at 2]{SELECT * FROM test WHERE a=2;}, \dataText[linkTo=testname, linkSelector=3, linkLabel=Value at 3]{SELECT * FROM test WHERE a=3;}.\\[3pt]
+\chart[width=0.45\textwidth, height=0.3\textwidth, logscale=true, chartType=line, name=testname]{SELECT * FROM test;}
+\chart[width=0.45\textwidth, height=0.3\textwidth, chartType=bar]{SELECT * FROM test;}\\[3pt]
 {\fontfamily{phv} \tiny \selectfont
@@
 Change the font before this line to Helvetica and fontsize to tiny.\\[3pt]
-\chart[width=\textwidth, height=0.45\textwidth, chartType=signatureplot]{SELECT b FROM test;}\\[3pt]
-\multiplotChart[width=\textwidth, height=0.9\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, topArr={["a smaller or equal 2", "a greater 2"]}, yFirst]{SELECT a,?
+\chart[width=0.6\textwidth, height=0.25\textwidth, chartType=signatureplot]{SELECT b FROM test;}\\[3pt]
+\multiplotChart[width=0.6\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, topArr={["a smaller or equal 2", "a greater 2"]}, yFirst]{SELECT a,?
 FROM test WHERE ?;}\\[3pt]
-\pivotTable[width=\textwidth, height=0.5\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}\\[3pt]
+\pivotTable[width=0.10\textwidth, height=0.15\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}
 \dataTable[verticalLines=i, horizontalLinesHeader=b hh, horizontalLinesBody=i]{SELECT * FROM test;}\\[3pt]

### 2015-11-30 src/pdbf/tests/charts.tex
- added new test for charts - reverted pivot.js to stable version - updated reference pictures

\documentclass[10pt]{article}
\usepackage{pdbf}
\usepackage[a4paper]{geometry}

\dbSQLText{CREATE TABLE test(a int, b int, c int); INSERT INTO test VALUES(1, 10, 15); INSERT INTO test VALUES(2, 100, 90); INSERT INTO test VALUES(3, 120, 110); INSERT INTO test VALUES(4, 100, 150);}

\def\test{
	%
	%%%%%%%%%%%%%%%%%%% Page1
	%
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"tick": { "rotate": 45 }
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"tick": { "rotate": 90 }
			}
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-left",
				"x": 25,
				"y": 0,
				"step": 1
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-left",
				"x": 25,
				"y": 0
			}
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-right",
				"step": 1
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "inset",
			"inset": {
				"anchor": "top-right"
			}
		}
	}]{SELECT * FROM test;}\\
	%
	%%%%%%%%%%%%%%%%%%% Page2
	%
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"position": "right"
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"legend": {
			"show": false
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "inner-center"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "inner-middle"
				}
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "inner-left"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "inner-bottom"
				}
			}
		}
	}]{SELECT * FROM test;}\\
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "outer-right"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "outer-top"
				}
			}
		}
	}]{SELECT * FROM test;}
	\chart[width=0.49\textwidth, height=0.45\textwidth, chartType=line, xunit=xLabel, yunit=yLabel, options={
		"axis": { 
			"x": { 
				"label": {
					"text": "xLabel",
	       			"position": "outer-center"
				}
			},
			"y": { 
				"label": {
					"text": "yLabel",
	       			"position": "outer-middle"
				}
			}
		}
	}]{SELECT * FROM test;}\\
}

\begin{document}
\noindent
\tiny
\test
\normalsize
\test
\LARGE
\test
%
\fontfamily{phv}
\tiny
\test
\normalsize
\test
\LARGE
\test
\end{document}

### 2015-11-30 data/pivot.js

@@ -1,1867 +1,2 @@
-(function() {
-	var callWithJQuery, __indexOf = [].indexOf || function(item) {
-		for (var i = 0, l = this.length; i < l; i++) {
-			if (i in this && this[i] === item)
-				return i;
-		}
-		return -1;
-	}, __slice = [].slice, __bind = function(fn, me) {
-		return function() {
-			return fn.apply(me, arguments);
-		};
-	}, __hasProp = {}.hasOwnProperty;
-
-	callWithJQuery = function(pivotModule) {
-		if (typeof exports === "object" && typeof module === "object") {
-			return pivotModule(require("jquery"));
-		} else if (typeof define === "function" && define.amd) {
-			return define([ "jquery" ], pivotModule);
-		} else {
-			return pivotModule(jQuery);
-		}
-	};
-
-	callWithJQuery(function($) {
-
-		/*
-		 * Utilities
-		 */
-		var PivotData, addSeparators, aggregatorTemplates, aggregators, dayNamesEn, derivers, getSort, locales, mthNamesEn, naturalSort, numberFormat, pivotTableRenderer, renderers, sortAs, usFmt, usFmtInt, usFmtPct, zeroPad;
-		addSeparators = function(nStr, thousandsSep, decimalSep) {
-			var rgx, x, x1, x2;
-			nStr += '';
-			x = nStr.split('.');
-			x1 = x[0];
-			x2 = x.length > 1 ? decimalSep + x[1] : '';
-			rgx = /(\d+)(\d{3})/;
-			while (rgx.test(x1)) {
-				x1 = x1.replace(rgx, '$1' + thousandsSep + '$2');
-			}
-			return x1 + x2;
-		};
-		numberFormat = function(opts) {
-			var defaults;
-			defaults = {
-			digitsAfterDecimal : 3,
-			scaler : 1,
-			thousandsSep : ",",
-			decimalSep : ".",
-			prefix : "",
-			suffix : "",
-			showZero : false
-			};
-			opts = $.extend(defaults, opts);
-			return function(x) {
-				var result;
-				if (isNaN(x) || !isFinite(x)) {
-					return "";
-				}
-				if (x === 0 && !opts.showZero) {
-					return "";
-				}
-				result = addSeparators((opts.scaler * x).toFixed(opts.digitsAfterDecimal), opts.thousandsSep, opts.decimalSep);
-				return "" + opts.prefix + result + opts.suffix;
[…]

### 2015-11-30 src/pdbf/tests/capturePages.js

@@ -60,9 +60,9 @@ var getPage = function (i, pagesCount) {
 	window.setTimeout(function () {
 		waitFor(
 			function() {
-				return page.evaluate(function() {
-					return !PDFViewer.updateInProgress;
-				});
+                return page.evaluate(function(i) {
+                	return pdbfReady(i);
+                }, i);
 			}, 
 			function() {
 				window.setTimeout(function () {

### 2015-12-23 pdbf-doc.tex
pdbf compiler now correctly supports compressed xref tables (PDF version >= 1.5)
replaced phantom-js for mac with a working version

Version 1.1.2
Converted windows to unix line-endings
Version 1.2.1 - Bugfixes - renamed signatureplot to compareToBest

@@
 }
-\date{\today{}, Saarbrücken}
+\date{}
 
@@
 \def\a{5cm}
-\def\b{10.5cm}
+\def\b{11.2cm}
 
@@
 \begin{itemize}
-\item Make sure you have a Java Runtime (version $>=$ 1.7) installed
+\item Make sure you have a Java Runtime (version $>=$ 1.7) and a LaTeX distribution installed
 \item \href{https://github.com/uds-datalab/PDBF/archive/gh-pages.zip}{Download the latest version}
@@
 \item Adjust config.cfg
-\item Try to compile pdbf-doc.tex file with this command: \texttt{java -jar pdbf.jar pdbf-doc.tex}
-\item Open pdbf-doc.html, this is the final output of the compilation process, if you rename it to ".pdf" it is also a valid pdf-document
-\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --vm pdbf-doc.html vldb-Invaders.ova}.
-\item Open pdbf-doc.ova, this is the final output with the attached ova file. Its still a valid pdf and html at the same time.
-\item You can play around with the minimal.tex file which contains a minimal example of PDBF usage
+\item Try to compile minimal.tex file with this command: \texttt{java -jar pdbf.jar minimal.tex}
+\item Open minimal.html, this is the final output of the compilation process, if you rename it to ".pdf" it is also a valid pdf-document
+\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --vm minimal.html vldb-Invaders.ova}.
+\item Open minimal.ova, this is the final output with the attached ova file. Its still a valid pdf and html at the same time.
+\item You can play around with the minimal.tex file. It contains a small example on how to specify PDBF elements in LaTeX
 \item For further information take a look at the \href{http://uds-datalab.github.io/PDBF/}{documentation}
@@
     \begin{minipage}{1.0\textwidth}
-    \chart[width=\textwidth, height=0.36\textwidth, chartType=signatureplot]{SELECT b FROM test;}
-    \caption{Signatureplot Chart}
+    \chart[width=\textwidth, height=0.36\textwidth, chartType=compareToBest]{SELECT b FROM test;}
+    \caption{compareToBest Chart}
     \end{minipage}
@@
 {Sets the type of the chart.}
-{\texttt{line} or \texttt{bar} or \texttt{signatureplot}}
+{\texttt{line} or \texttt{bar} or \texttt{compareToBest}}
 {\texttt{line}}
+
+\noindent\textbf{Note:} A chart or multiplot chart with option "chartType" set to "compareToBest" expects either an array of numbers or an array of an array of numbers as input.
+It plots the slowdown compared to the best value in the array.
+In the second case the inner array of numbers contains values for repetitions of the same experiment.
+\\[6pt]
 

### 2015-12-23 external-tools/capture.js

@@ -89,7 +89,7 @@ page.open(system.args[1], function() {
 			window.setTimeout(function () {
 				console.log("Finished " + data.name);
 				fs.write('../' + data.name + '.json', data.result, 'w');
-				page.render('../' + data.name +'.pdf', {format: 'pdf', quality: '100'});
+				page.render('../' + data.name +'.pdf', {format: 'pdf', quality: '0'});
 				phantom.exit();
 			}, 1000);
 		},

### 2015-12-23 src/pdbf/latex/LaTeX_Compiler.java

@@ -7,9 +7,9 @@ import java.io.IOException;
 import java.io.InputStreamReader;
 import java.sql.Connection;
 import java.sql.DriverManager;
-import java.sql.PreparedStatement;
 import java.sql.ResultSet;
 import java.sql.ResultSetMetaData;
+import java.sql.Statement;
 import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.Date;
@@ -20,6 +20,7 @@ import org.apache.commons.csv.CSVFormat;
 import org.apache.commons.csv.CSVParser;
 import org.apache.commons.csv.CSVRecord;
 import org.apache.commons.io.FileUtils;
+import org.postgresql.util.PSQLException;
 
 import com.google.gson.Gson;
 import com.google.gson.GsonBuilder;
@@ -82,8 +83,7 @@ public class LaTeX_Compiler {
 		    StringTokenizer stok = new StringTokenizer(strLine, ":");
 		    stok.nextToken();
 		    dpiScalingFactor = Float.parseFloat(stok.nextToken());
-		}
-		else if (!strLine.startsWith("#") && !strLine.equals("")) {
+		} else if (!strLine.startsWith("#") && !strLine.equals("")) {
 		    tmp.add(strLine);
 		}
 	    }
@@ -149,6 +149,11 @@ public class LaTeX_Compiler {
 		e3.printStackTrace();
 		System.exit(-1);
 	    }
+	}
+
+	for (int i = 0; i < 100; ++i) {
+	    new File(baseDir + "Overlay" + i + ".pdf").delete(); //HACK: replace this by a scan for all OverlayX.pdf files where X is a number
+	    new File(latexFolder + File.separator + "Overlay" + i + ".pdf").delete();
 	}
 
 	System.out.println("Compiling LaTeX (1/2)...");
@@ -432,25 +437,29 @@ public class LaTeX_Compiler {
 			System.exit(-1);
 		    }
 
-		    PreparedStatement stmt = conn.prepareStatement("SELECT * FROM ?");
-		    stmt.setString(1, curTable);
-		    ResultSet rs = stmt.executeQuery();
-		    ResultSetMetaData rsmd = rs.getMetaData();
-		    int cols = rsmd.getColumnCount();
-		    for (int i = 1; i <= cols; ++i) {
-			table.columns.add(new Column(rsmd.getColumnName(i), rsmd.getColumnTypeName(i).toUpperCase()));
-		    }
-		    if (cols > 0) {
[…]

### 2016-02-01 pdbf-doc.tex
major restructuring - tar support

@@
 \section{How to use the PDBF compiler}
-Right now there are only two modes: \\[4pt]
+Right now there are only three modes: \\[4pt]
 -- The standart mode which compiles a tex file into the janiform PDBF format:\\
 \verb|java -jar PDBF.jar LaTeX_file| \\[2pt]
-\textbf{Note:} The compiled document can be found in the same folder as the tex file and has the same name as the tex file but ".html" filename extension.\\[8pt]
--- The VM mode which attaches a VirtualBox .ova file to an existing PDBF:\\
+\textbf{Note:} The compiled document can be found in the same folder as the tex file and has the same name as the tex file but ".html" as filename extension.\\[8pt]
+-- The VM mode which attaches a VirtualBox .ova file to an existing PDBF document:\\
 \verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|\\[2pt]
-\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".ova" filename extension.
+\textbf{Note:} Currently there is a hard size limit of 1.9GB for .ova files!
+\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".ova" as filename extension.
+-- The TAR mode which attaches a tar archive file to an existing PDBF document:\\
+\verb|java -jar PDBF.jar --tar PDBF_File.html TAR_File.tar|\\[2pt]
+\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".tar" as filename extension.
+\textbf{Note:} Currently there is a hard size limit of 1.9GB for .tar files!
 

### 2016-02-01 src/pdbf/PDBF_Compiler.java

@@ -11,12 +11,12 @@ import org.apache.log4j.BasicConfigurator;
 import org.apache.log4j.Level;
 import org.apache.log4j.Logger;
 
-import pdbf.compilers.HTML_Compiler;
-import pdbf.compilers.LaTeX_Compiler;
+import pdbf.compilers.HTML_PDF_Compiler;
+import pdbf.compilers.Pre_Compiler;
 import pdbf.compilers.TAR_Compiler;
 import pdbf.compilers.VM_Compiler;
-import pdbf.tools.MinifyResources;
-import pdbf.tools.Tools;
+import pdbf.misc.MinifyResources;
+import pdbf.misc.Tools;
 
 public class PDBF_Compiler {
 
@@ -31,6 +31,8 @@ public class PDBF_Compiler {
 			+ "--no-pdf-protection : Disables the protection of the pdf part of the PDBF file\n"
 			+ "OR\n"
 			+ "java -jar pdbf.jar --vm  PDBF_File.html VM_File.ova\n"
+			+ "OR\n"
+			+ "java -jar pdbf.jar --tar  PDBF_File.html TAR_File.tar\n"
 			+ "For further help visit: https://github.com/uds-datalab/PDBF");
 	System.exit(0);
     }
@@ -39,8 +41,8 @@ public class PDBF_Compiler {
 	if (includeRes) {
 	    MinifyResources.main(args);
 	}
-	LaTeX_Compiler.main(args);
-	HTML_Compiler.main(args);
+	Pre_Compiler.main(args);
+	HTML_PDF_Compiler.main(args);
     }
 
     public static void main(String[] args) {
@@ -49,7 +51,7 @@ public class PDBF_Compiler {
 	Logger.getRootLogger().setLevel(Level.ERROR);
 
 	//
-	LaTeX_Compiler.suffix = Tools.getOS();
+	Pre_Compiler.suffix = Tools.getOS();
 
 	String version = null;
 	try {
@@ -99,11 +101,11 @@ public class PDBF_Compiler {
 	    VM_Compiler.main(args);
 	    System.exit(0);
 	} else if (args.length == 1) {
-	    LaTeX_Compiler.pdfProtect = true;
+	    Pre_Compiler.pdfProtect = true;
 	    compile(args);
 	} else if (args.length == 2
 		&& args[1].equalsIgnoreCase("--no-pdf-protection")) {
-	    LaTeX_Compiler.pdfProtect = false;
+	    Pre_Compiler.pdfProtect = false;
 	    compile(args);
 	}
 

### 2016-02-01 src/pdbf/compilers/LaTeX_Compiler.java

@@ -1,880 +1,0 @@
-package pdbf.compilers;
-
-import java.io.BufferedReader;
-import java.io.File;
-import java.io.FileInputStream;
-import java.io.FileOutputStream;
-import java.io.IOException;
-import java.io.InputStream;
-import java.io.InputStreamReader;
-import java.io.OutputStream;
-import java.sql.Connection;
-import java.sql.DriverManager;
-import java.sql.ResultSet;
-import java.sql.ResultSetMetaData;
-import java.sql.Statement;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.Date;
-import java.util.List;
-import java.util.StringTokenizer;
-import java.util.UUID;
-import java.util.regex.Pattern;
-
-import org.apache.commons.csv.CSVFormat;
-import org.apache.commons.csv.CSVParser;
-import org.apache.commons.csv.CSVRecord;
-import org.apache.commons.io.FileUtils;
-import org.apache.pdfbox.pdmodel.PDDocument;
-import org.apache.pdfbox.pdmodel.encryption.AccessPermission;
-import org.apache.pdfbox.pdmodel.encryption.StandardProtectionPolicy;
-import org.postgresql.util.PSQLException;
-
-import com.google.gson.Gson;
-import com.google.gson.GsonBuilder;
-
-import pdbf.PDBF_Compiler;
-import pdbf.json.Dimension;
-import pdbf.json.Overlay;
-import pdbf.json.Visualization;
-import pdbf.json.VisualizationTypeAdapter;
-import pdbf.json.alasql.Alasql;
-import pdbf.json.alasql.Column;
-import pdbf.json.alasql.Data;
-import pdbf.json.alasql.DatabaseContainer;
-import pdbf.json.alasql.Table;
-import pdbf.json.pdbf.Chart;
-import pdbf.json.pdbf.DataTable;
-import pdbf.json.pdbf.DataText;
-import pdbf.json.pdbf.Database;
-import pdbf.json.pdbf.Text;
-import pdbf.tools.Tools;
-
-public class LaTeX_Compiler {
-
-    private static String[] pathToLaTeXScript = new String[0];
-
-    private static ArrayList<Process> processes = new ArrayList<Process>();
-    private static ArrayList<String> cleanupfiles = new ArrayList<String>();
-    private static ArrayList<String> copyfiles = new ArrayList<String>();
-    private static ArrayList<String> preloadfiles = new ArrayList<String>();
[…]

### 2016-02-01 src/pdbf/compilers/Pre_Compiler.java

@@ -1,0 +1,798 @@
+package pdbf.compilers;
+
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.FileOutputStream;
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.InputStreamReader;
+import java.io.OutputStream;
+import java.sql.Connection;
+import java.sql.DriverManager;
+import java.sql.ResultSet;
+import java.sql.ResultSetMetaData;
+import java.sql.Statement;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.Date;
+import java.util.List;
+import java.util.StringTokenizer;
+import java.util.UUID;
+import java.util.regex.Pattern;
+
+import org.apache.commons.csv.CSVFormat;
+import org.apache.commons.csv.CSVParser;
+import org.apache.commons.csv.CSVRecord;
+import org.apache.commons.io.FileUtils;
+import org.apache.pdfbox.pdmodel.PDDocument;
+import org.apache.pdfbox.pdmodel.encryption.AccessPermission;
+import org.apache.pdfbox.pdmodel.encryption.StandardProtectionPolicy;
+import org.postgresql.util.PSQLException;
+
+import com.google.gson.Gson;
+import com.google.gson.GsonBuilder;
+
+import pdbf.PDBF_Compiler;
+import pdbf.json.Database;
+import pdbf.json.Dimension;
+import pdbf.json.PDBFelement;
+import pdbf.json.PDBFelementContainer;
+import pdbf.json.PDBFelementTypeAdapter;
+import pdbf.json.Text;
+import pdbf.json.TextualPDBFelement;
+import pdbf.json.VisualPDBFelement;
+import pdbf.json.alasql.Alasql;
+import pdbf.json.alasql.Column;
+import pdbf.json.alasql.Data;
+import pdbf.json.alasql.DatabaseContainer;
+import pdbf.json.alasql.Table;
+import pdbf.misc.Tools;
+
+/*
+ * This class produces all necessary auxiliary files that are needed to produce the final PDBF file
+ */
+public class Pre_Compiler {
+
+    private static String[] pathToLaTeXScript = new String[0];
+
+    private static ArrayList<Process> processes = new ArrayList<Process>();
+    private static ArrayList<String> cleanupfiles = new ArrayList<String>();
+    private static ArrayList<String> copyfiles = new ArrayList<String>();
[…]

### 2016-02-01 src/pdbf/json/Pivot.java

@@ -1,0 +1,9 @@
+package pdbf.json;
+
+/*
+ * Defines the properties that are exclusive to Pivot
+ * Common properties are defined in PDBFelement.java
+ */
+public class Pivot extends VisualPDBFelement {
+
+}

### 2016-02-01 src/pdbf/json/pdbf/Chart.java

@@ -1,12 +1,0 @@
-package pdbf.json.pdbf;
-
-import pdbf.json.Visualization;
-
-/*
- * Defines the properties that are exclusive to Charts
- * Common properties are defined in Visualization.java
- */
-
-public class Chart extends Visualization {
-
-}

### 2016-02-01 src/pdbf/json/pdbf/Database.java

@@ -1,23 +1,0 @@
-package pdbf.json.pdbf;
-
-import pdbf.json.Visualization;
-
-/*
- * Defines the properties that are exclusive to Database
- * For simplicity Database elements are also considered to be Visualizations
- */
-
-public class Database extends Visualization {
-    public int type;
-    public String value1;
-    public String value2;
-    public String value3;
-    public String value4;
-    public String[] headers;
-    public char quote;
-    public char seperator;
-
-    public Database() {
-    }
-
-}

### 2016-02-01 src/pdbf/json/pdbf/Pivot.java

@@ -1,12 +1,0 @@
-package pdbf.json.pdbf;
-
-import pdbf.json.Visualization;
-
-/*
- * Defines the properties that are exclusive to Pivot
- * Common properties are defined in Visualization.java
- */
-
-public class Pivot extends Visualization {
-
-}

### 2016-02-01 src/pdbf/json/pdbf/Text.java

@@ -1,12 +1,0 @@
-package pdbf.json.pdbf;
-
-import pdbf.json.Visualization;
-
-/*
- * Defines the properties that are exclusive to Text
- * Common properties are defined in Visualization.java
- */
-
-public class Text extends Visualization {
-    
-}

### 2016-02-01 src/pdbf/misc/Tools.java

@@ -1,0 +1,269 @@
+package pdbf.misc;
+
+import java.io.ByteArrayInputStream;
+import java.io.ByteArrayOutputStream;
+import java.io.File;
+import java.io.IOException;
+import java.io.UnsupportedEncodingException;
+import java.net.URISyntaxException;
+import java.nio.ByteBuffer;
+import java.nio.charset.Charset;
+import java.nio.charset.StandardCharsets;
+import java.security.CodeSource;
+import java.text.DecimalFormat;
+import java.util.ArrayList;
+import java.util.regex.Matcher;
+import java.util.regex.Pattern;
+import java.util.zip.Deflater;
+import java.util.zip.DeflaterOutputStream;
+import java.util.zip.InflaterOutputStream;
+
+import org.apache.commons.codec.binary.Base64;
+import org.apache.commons.io.FileUtils;
+
+import pdbf.PDBF_Compiler;
+
+/*
+ * Useful functions
+ */
+public class Tools {
+
+    public static DecimalFormat df = new DecimalFormat("0000000000");
+
+    public static Charset utf8 = Charset.forName("UTF-8");
+
+    public static String baseDir = new File(getBaseDir()).getParent() + File.separator;
+    public static String refDir = baseDir + "src" + File.separator + "pdbf" + File.separator + "referenceImages" + File.separator;
+    public static String testDir = baseDir + "src" + File.separator + "pdbf" + File.separator + "tests" + File.separator;
+    public static String suffix = getOS();
+
+    public static ArrayList<File> deleteList = new ArrayList<File>();
+    public static ArrayList<Process> processes = new ArrayList<Process>();
+
+    /*
+     * Helper function of fixXref. Skips all \n and \r characters in a
+     * stringbuilder either in forward or backward direction
+     */
+    public static int skipNewline(StringBuilder sb, int i, boolean forward) {
+	if (i == -1) {
+	    return -1;
+	}
+	if (!forward) {
+	    --i;
+	}
+	while (sb.charAt(i) == '\n' || sb.charAt(i) == '\r') {
+	    i = forward ? i + 1 : i - 1;
+	}
+	return forward ? i : i + 1;
+    }
+
+    /*
+     * Check if operating system is supported and return corresponding short
+     * name of the OS
+     */
[…]

### 2016-02-04 pdbf-doc.tex
sync README and pdbf-doc

@@
 \section{Getting started}
+\subsection{Normal usage}
 \begin{itemize}
@@
 \item Open minimal.html, this is the final output of the compilation process, if you rename it to ".pdf" it is also a valid pdf-document
+\end{itemize}
+\subsection{Attach a VirtualBox image}
+\begin{itemize}
 \item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --vm minimal.html vldb-Invaders.ova}.
-\item Open minimal.ova, this is the final output with the attached ova file. Its still a valid pdf and html at the same time.
+\item Open minimal.ova, this is the final output with the attached ova file. Its still a valid pdf and html file at the same time.
+\end{itemize}
+\subsection{Attach a tar archive}
+\begin{itemize}
+\item Optionally you can attach a tar archive file to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --tar minimal.html TAR\_file.tar}
+\item Open minimal.tar, this is the final output with the attached tar file. Its still a valid pdf and html file at the same time.
+\end{itemize}
+\subsection{Learn how to use the PDBF framework}
+\begin{itemize}
 \item You can play around with the minimal.tex file. It contains a small example on how to specify PDBF elements in LaTeX
@@
 \textbf{Note:} Currently there is a hard size limit of 1.9GB for .ova files!
-\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".ova" as filename extension.
+\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".ova" as filename extension.\\[8pt]
 -- The TAR mode which attaches a tar archive file to an existing PDBF document:\\

### 2016-02-10 minimal.tex
Converted windows to unix line-endings
Version 1.2.1 - Bugfixes - renamed signatureplot to compareToBest
- fixed a bug that sometimes made the xref fixing fail - minified all js and css files - updated lot of javascript libraries to up to date versions

@@
 
-\dbSQLText{CREATE TABLE test(a int, b int, c int); INSERT INTO test VALUES(1, 1, 1); INSERT INTO test VALUES(2, 2, 4); INSERT INTO test VALUES(3, 3, 8); INSERT INTO test VALUES(4, 4, 16);}
+\dbSQLText{CREATE TABLE test(a int, b int, [attribute with space] int); INSERT INTO test VALUES(1, 1, 1); INSERT INTO test VALUES(2, 2, 4); INSERT INTO test VALUES(3, 3, 8); INSERT INTO test VALUES(4, 4, 16);}
 %\dbSQLFile{somedb.sql}
@@
 The following points are linked to the chart on the left side below: \dataText[linkTo=testname, linkSelector=2, linkLabel=Value at 2]{SELECT * FROM test WHERE a=2;}, \dataText[linkTo=testname, linkSelector=3, linkLabel=Value at 3]{SELECT * FROM test WHERE a=3;}.\\[3pt]
-\chart[width=0.45\textwidth, height=0.3\textwidth, logscale=true, chartType=line, name=testname]{SELECT * FROM test;}
-\chart[width=0.45\textwidth, height=0.3\textwidth, chartType=bar]{SELECT * FROM test;}\\[3pt]
+\chart[width=0.45\textwidth, height=0.3\textwidth, logscale=true, chartType=line, name=testname]{SELECT a, b, [attribute with space] FROM test;}
+\chart[width=0.45\textwidth, height=0.3\textwidth, chartType=bar]{SELECT a, b, [attribute with space] FROM test;}\\[3pt]
 {\fontfamily{phv} \tiny \selectfont
@@
 Change the font before this line to Helvetica and fontsize to tiny.\\[3pt]
-\chart[width=0.6\textwidth, height=0.25\textwidth, chartType=signatureplot]{SELECT b FROM test;}\\[3pt]
+\chart[width=0.6\textwidth, height=0.25\textwidth, chartType=compareToBest]{SELECT b FROM test;}\\[3pt]
 \multiplotChart[width=0.6\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, topArr={["a smaller or equal 2", "a greater 2"]}, yFirst]{SELECT a,?

### 2016-02-10 pdbf-doc.tex
- fixed a bug that sometimes made the xref fixing fail - minified all js and css files - updated lot of javascript libraries to up to date versions

@@
 lz-string (\url{https://github.com/pieroxy/lz-string})\\
+YUI Compressor (\url{https://github.com/yui/yuicompressor})\\
 

### 2016-02-10 src/pdbf/misc/MinifyResources.java

@@ -14,13 +14,16 @@ public class MinifyResources {
     public static void main(String[] args) {
 	String baseDirData = Tools.getBaseDirData();
 
-	String command[] = { "java", "-jar", baseDirData + "compiler.jar", baseDirData + "d3.js", baseDirData + "c3.js", baseDirData + "excanvas.compiled.js",
-		baseDirData + "diff_match_patch.js", baseDirData + "jquery-1.11.2.min.js", baseDirData + "pivot.js",
+	String command[] = { "java", "-jar", baseDirData + "google-closure-compiler-20160208.jar", baseDirData + "alasql.js", baseDirData + "base64.js",
+		baseDirData + "lz-string.js", baseDirData + "d3.js", baseDirData + "c3.js", baseDirData + "excanvas.compiled.js",
+		baseDirData + "diff_match_patch.js", baseDirData + "jquery-3.0.0-beta1.min.js", baseDirData + "pivot.js",
 		baseDirData + "jquery-ui-1.9.2.custom.min.js", baseDirData + "l10n.js", baseDirData + "viewer.js", baseDirData + "main.js",
-		baseDirData + "preMain.js", baseDirData + "compatibility.js", baseDirData + "jstat.js", baseDirData + "pdf.js", "--js_output_file",
-		baseDirData + "all", "--language_in", "ECMASCRIPT5", "--compilation_level", "WHITESPACE_ONLY", "--charset", "UTF-8" };
+		baseDirData + "preMain.js", baseDirData + "compatibility.js", baseDirData + "jstat.js", baseDirData + "pdf.js",
+		baseDirData + "jquery.dataTables.js", baseDirData + "codemirror-compressed.js", "--js_output_file", baseDirData + "all.js", "--language_in",
+		"ECMASCRIPT5", "--compilation_level", "WHITESPACE_ONLY", "--charset", "UTF-8" };
 
 	try {
+	    new File(baseDirData + "all.js").delete();
 	    ProcessBuilder pb = new ProcessBuilder(Arrays.asList(command));
 	    pb.inheritIO();
 	    Process p = pb.start();
@@ -33,29 +36,48 @@ public class MinifyResources {
 	    e.printStackTrace();
 	}
 
-	try {
-	    // TODO: for unknown reasons some files break when we minify them.
-	    // Therefore we include them without minification
-	    String all = FileUtils.readFileToString(new File(baseDirData + "all"), Tools.utf8);
[…]

### 2016-02-29 pdbf-doc.tex
Version 1.2.4 - Lots of new commandline options - You can now use arbitrary images for the in-page view of a PDBF Element

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{listings}
 \usepackage[a4paper,vmargin={20mm,40mm},hmargin={20mm,20mm}]{geometry}
@@
 These are not fully compatible with PDBF.
-If you really have to use the utf8 package then you cant use special characters inside options and queries of the PDBF package.
+If you really have to use the utf8 package then you cant use special characters inside options and queries of the PDBF package.\\
+\item Q.: I get strange errors when running "maven verify".\\
+A.: If you are using TexLive 2014 or earlier and get the following errors when running "maven verify":
+\begin{lstlisting}[breaklines]
+The file is not valid, error(s) :
+1.2.1 : Body Syntax error, Single space expected [offset=2786901; key=2786901; line=5 0 obj <<; object=COSObject{5, 0}]
+1.2.1 : Body Syntax error, Single space expected [offset=2786635; key=2786635; line=11 0 obj <<; object=COSObject{11, 0}]
+1.2.1 : Body Syntax error, EOL expected before the 'endobj' keyword at offset 2786894
+1.2.1 : Body Syntax error, Single space expected [offset=2761071; key=2761071; line=3 0 obj <<; object=COSObject{3, 0}]
+1.2.1 : Body Syntax error, Single space expected [offset=2761433; key=2761433; line=8 0 obj <<; object=COSObject{8, 0}]
+1.2.1 : Body Syntax error, Single space expected [offset=2787764; key=2787764; line=12 0 obj <<; object=COSObject{12, 0}]
+\end{lstlisting}
+Then you can safely ignore them. They are non critical errors. You can get rid of these errors by upgrading to TexLive 2015 or later.
 \end{itemize}
-
 \end{document}

### 2016-02-29 src/pdbf/json/PDBFelement.java

@@ -53,6 +53,7 @@ public abstract class PDBFelement {
     public double quality;
     public double fontsize;
     public String name;
+    public String customImage;
 
     // DataText
     public String linkTo;

### 2016-03-23 src/pdbf/compilers/VM_Compiler.java
Finished streaming implementation for ova/tar attachements

@@ -5,10 +5,12 @@ import java.io.RandomAccessFile;
 import java.nio.charset.StandardCharsets;
 import java.text.DecimalFormat;
 import java.text.NumberFormat;
+import java.util.Arrays;
 import java.util.regex.Matcher;
 import java.util.regex.Pattern;
 
 import pdbf.PDBF_Compiler;
+import pdbf.misc.Tools;
 import pdbf.tests.CheckAttached;
 
 public class VM_Compiler {
@@ -23,14 +25,14 @@ public class VM_Compiler {
 	String a = new File(args[0]).getName();
 	// TODO: sanity checks. compare beginning of html/pdbf file with
 	// "removeFromHTML".
-	if (!a.toUpperCase().endsWith(".HTML")) {
-	    System.err.println("Error: Only .HTML files are supported as first argument!");
+	if (!a.toLowerCase().endsWith(".html")) {
+	    System.err.println("Error: Only .html files are supported as first argument!");
 	    System.exit(1);
 	}
 	// TODO: sanity checks. use "testTar" to check if ova/tar file is valid.
 	// for ova we should even test if name of the first file in tar ends
 	// with ".ovf"
-	if (!args[1].toUpperCase().endsWith("." + fileType)) {
+	if (!args[1].toLowerCase().endsWith("." + fileType)) {
 	    System.err.println("Error: Only ." + fileType + " files are supported as second argument!");
 	    System.exit(1);
 	}
@@ -48,17 +50,18 @@ public class VM_Compiler {
 	    outFile.setLength(0);
 
 	    // For keeping track of progress
-	    String removeFromHTML = "%PDF-1.5\n%ª«¬­<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>" + "<head><meta charset=\"utf-8\"><!--\n1337 0 obj\nstream";
-	    String addToHTML = "</script>";
+	    String removeFromPdbf = "%PDF-1.5\n%ª«¬­<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>"
+		    + "<head><meta charset=\"utf-8\"><script>\n1337 0 obj\nstream\n</script>\n";
+	    String addToPdbf = "</script>";
 	    RandomAccessFile ovaOrTarFile = new RandomAccessFile(new File(args[1]), "rw");
[…]

### 2016-04-06 pdbf-doc.tex
Finished streaming implementation for ova/tar attachements
fixed tar and ova attaching

@@
 \usepackage[a4paper,vmargin={20mm,40mm},hmargin={20mm,20mm}]{geometry}
+
+\pdfobjcompresslevel=2
 
@@
 They are a valid PDF and a valid HTML page at the same time.
-You can now optionally add an VirtualBox OVA file with a complete operating system to the PDBF document.
+You can now optionally add an Open Virtual Appliance (OVA) file with a complete operating system to the PDBF document.
 Yes, this means that the resulting file is a valid PDF, HTML, and OVA file at the same time.
@@
 \end{itemize}
-\subsection{Attach a VirtualBox image}
+\subsection{Attach a Open Virtual Appliance (OVA) file}
 \begin{itemize}
-\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) VirtualBox image and attach it to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --vm minimal.html vldb-Invaders.ova}.
+\item Optionally you can attach the included vldb-Invaders.ova (Space invaders clone) or \href{https://github.com/uds-datalab/PDBF/releases/download/1.0.1/dsl.ova}{download the dsl.ova} (Damn small linux) Open Virtual Appliance (OVA) file and attach it to the compiled PDBF file with this command: \texttt{java -jar pdbf.jar --vm minimal.html vldb-Invaders.ova}.
 \item Open minimal.ova, this is the final output with the attached ova file. Its still a valid pdf and html file at the same time.
@@
 \subsection{Attach a tar archive}
+\textbf{Note:} compressed tar.gz archives are not supported. You can only wrap your tar.gz file in another tar and then attach it.
 \begin{itemize}
@@
 \section{How to use the PDBF compiler}
-Right now there are only three modes: \\[4pt]
 -- The standart mode which compiles a tex file into the janiform PDBF format:\\
-\verb|java -jar PDBF.jar LaTeX_file| \\[2pt]
+\verb|java -jar PDBF.jar LaTeX_file| \\[4pt]
 \textbf{Note:} The compiled document can be found in the same folder as the tex file and has the same name as the tex file but ".html" as filename extension.\\[8pt]
--- The VM mode which attaches a VirtualBox .ova file to an existing PDBF document:\\
-\verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|\\[2pt]
-\textbf{Note:} Currently there is a hard size limit of 1.9GB for .ova files!
+-- The VM mode which attaches a Open Virtual Appliance (OVA) file to an existing PDBF document:\\
+\verb|java -jar PDBF.jar --vm PDBF_File.html VM_File.ova|\\[4pt]
 \textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".ova" as filename extension.\\[8pt]
 -- The TAR mode which attaches a tar archive file to an existing PDBF document:\\
-\verb|java -jar PDBF.jar --tar PDBF_File.html TAR_File.tar|\\[2pt]
-\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".tar" as filename extension.
-\textbf{Note:} Currently there is a hard size limit of 1.9GB for .tar files!
+\verb|java -jar PDBF.jar --tar PDBF_File.html TAR_File.tar|\\[4pt]
+\textbf{Note:} The compiled document can be found in the same folder as the html file and has the same name as the html file but ".tar" as filename extension.\\[2pt]
+\textbf{Note:} compressed tar.gz archives are not supported. You can only wrap your tar.gz file in another tar and then attach it.
 

### 2016-04-06 src/pdbf/PDBF_Compiler.java

@@ -25,8 +25,7 @@ import org.apache.log4j.Logger;
 
 import pdbf.compilers.HTML_PDF_Compiler;
 import pdbf.compilers.Pre_Compiler;
-import pdbf.compilers.TAR_Compiler;
-import pdbf.compilers.VM_Compiler;
+import pdbf.compilers.VM_TAR_Compiler;
 import pdbf.misc.MinifyResources;
 import pdbf.misc.Tools;
 import pdbf.tests.CheckAttached;
@@ -34,7 +33,7 @@ import pdbf.tests.CreateReferencePictures;
 
 public class PDBF_Compiler {
 
-    private static int bufferSizeInMB = 50; //TODO: make option to change this. Add F.A.Q entry for java.lang.OutOfMemoryError: Java heap space 
+    private static int bufferSizeInMB = 2; //TODO: make option to change this. Add F.A.Q entry for java.lang.OutOfMemoryError: Java heap space 
     public static byte[] bytearray = new byte[1024*1024*PDBF_Compiler.bufferSizeInMB];
     
     // This flag indicates if all js and css files from data folder are included
@@ -100,11 +99,9 @@ public class PDBF_Compiler {
 	optionsGroup.addOption(Option.builder("t").longOpt("tar").numberOfArgs(2).argName("PDBF_File.html> <TAR_File.tar")
 		.desc("Appends a tar archive to an existing PDBF document. Only exactly one ova or tar file can be appended to an PDBF document.").build());
 	optionsGroup.addOption(Option.builder("v").longOpt("vm").numberOfArgs(2).argName("PDBF_File.html> <OVA_File.ova")
-		.desc("Appends a VirtualBox ova file to an existing PDBF document. Only exactly one ova or tar file can be appended to an PDBF document.")
+		.desc("Appends a Open Virtual Appliance (OVA) file to an existing PDBF document. Only exactly one ova or tar file can be appended to an PDBF document.")
 		.build());
-	optionsGroup.addOption(Option.builder("cri").longOpt("create-reference-images").numberOfArgs(0)
-		.desc("Appends a VirtualBox ova file to an existing PDBF document. Only exactly one ova or tar file can be appended to an PDBF document.")
-		.build());
+	optionsGroup.addOption(Option.builder("cri").longOpt("create-reference-images").numberOfArgs(0).build());
[…]

### 2016-04-06 src/pdbf/compilers/HTML_PDF_Compiler.java

@@ -14,7 +14,7 @@ import pdbf.misc.Tools;
 public class HTML_PDF_Compiler {
 
     public static String pdfHTML_Header = "%ª«¬­<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>"
-	    + "<head><meta charset=\"utf-8\"><script>\n" + "1337 0 obj\n" + "stream\n";
+	    + "<head><meta charset=\"utf-8\"><script>\n" + "1337 0 obj\n" + "stream\n" + "</script>\n";
 
     public static void main(String[] args) {
 	String baseDir = Tools.getBaseDir();
@@ -53,7 +53,7 @@ public class HTML_PDF_Compiler {
 		    + "\";\r\n" + "db_base64 = \"\";\r\n" + "json_base64 = \"" + Tools.encodeFileToBase64Binary(new File(baseDir + "pdbf-config.json"))
 		    + "\";\r\n" + "dbjson_base64 = \"" + Tools.escapeSpecialChars(new File(baseDir + "pdbf-db.json")) + "\";\r\n" + preload + "\r\n"
 		    + (PDBF_Compiler.includeRes ? (all + "\r\n") : ("")) + add + viewerTAIL;
-	    String insert = pdfHTML_Header + "</script>\n" + viewer + "<script>\n" + "endstream\n" + "endobj\n";
+	    String insert = pdfHTML_Header + viewer + "<script>\n" + "endstream\n" + "endobj\n";
 	    String pdfcontent = FileUtils.readFileToString(new File(pdfname), StandardCharsets.ISO_8859_1);
 	    if (pdfcontent.toLowerCase().contains("</script>")) {
 		System.err.println("The generated pdf cannot be used to generate a pdbf document!"

### 2016-04-06 src/pdbf/compilers/VM_TAR_Compiler.java

@@ -13,14 +13,22 @@ import pdbf.PDBF_Compiler;
 import pdbf.misc.Tools;
 import pdbf.tests.CheckAttached;
 
-public class VM_Compiler {
+public class VM_TAR_Compiler {
 
     public static DecimalFormat df = new DecimalFormat("0000000000");
 
     public static String fileType = "ova";
 
+    public static String ova_Header = "%PDF-1.5\n%ª«¬­.ovf\0\n1337 0 obj\nstream\n<head><meta charset=UTF-8><script>";
+    public static String tar_Header = "%PDF-1.5\n%ª«¬­DO_NOT_DELETE\0\n1337 0 obj\nstream\n<head><meta charset=UTF-8><script>";
+
     public static void main(String[] args) {
-	System.out.println("Compiling VM...");
+	boolean isVM = fileType.equals("ova");
+	if (isVM) {
+	    System.out.println("Compiling VM...");
+	} else {
+	    System.out.println("Compiling TAR...");
+	}
 
 	String a = new File(args[0]).getName();
 	// TODO: sanity checks. compare beginning of html/pdbf file with
@@ -50,8 +58,7 @@ public class VM_Compiler {
 	    outFile.setLength(0);
 
 	    // For keeping track of progress
-	    String removeFromPdbf = "%PDF-1.5\n%ª«¬­<!DOCTYPE html><html dir=\"ltr\" mozdisallowselectionprint moznomarginboxes>"
-		    + "<head><meta charset=\"utf-8\"><script>\n1337 0 obj\nstream\n</script>\n";
+	    String removeFromPdbf = "%PDF-1.5\n" + HTML_PDF_Compiler.pdfHTML_Header;
 	    String addToPdbf = "</script>";
 	    RandomAccessFile ovaOrTarFile = new RandomAccessFile(new File(args[1]), "rw");
 	    RandomAccessFile pdbfFile = new RandomAccessFile(new File(args[0]), "rw");
@@ -61,15 +68,27 @@ public class VM_Compiler {
 
 	    // Replace beginning of the ovaOrTarFile file with our
 	    // (ova/tar)/pdf/html file header
-	    String replaceInOvaOrTar = "%PDF-1.5\n%ª«¬­.ovf\0\n1337 0 obj\nstream\n<head><meta charset=UTF-8><script>";
-	    ovaOrTarFile.seek(replaceInOvaOrTar.length());
[…]

### 2016-04-06 src/pdbf/tests/CheckAttached.java

@@ -2,13 +2,32 @@ package pdbf.tests;
 
 import java.io.File;
 import java.io.FileInputStream;
+import java.nio.charset.StandardCharsets;
+
 import org.apache.commons.compress.archivers.tar.TarArchiveInputStream;
+
+import pdbf.compilers.VM_TAR_Compiler;
 
 public class CheckAttached {
 
     public static void main(String[] args) {
 	if (checkAttached(new File(args[0]))) {
-	    System.out.println("This file has an ova or tar file attached");
+	    String start = "";
+	    try {
+		FileInputStream fis = new FileInputStream(new File(args[0]));
+		byte[] b = new byte[512];
+		fis.read(b, 0, b.length);
+		start = new String(b, StandardCharsets.ISO_8859_1);
+	    } catch (Exception e) {
+		e.printStackTrace();
+		System.exit(1);
+	    }
+	    int len = "%PDF-1.5\n%ª«¬­.ovf\0".length();
+	    if (start.substring(0, len).equals(VM_TAR_Compiler.ova_Header.substring(0, len))) {
+		System.out.println("This file has an ova file attached");
+	    } else {
+		System.out.println("This file has an tar file attached");
+	    }
 	} else {
 	    System.out.println("This file doesnt have an ova or tar file attached");
 	}

### 2016-07-18 minimal.tex
restored minimal

@@
 \noindent
+
 \sql{SELECT * FROM test;}\\[3pt]
 The table ``test'' contains \dataText{SELECT COUNT(*) FROM test;} tuples.\\[3pt]
+
 The following points are linked to the chart on the left side below: \dataText[linkTo=testname, linkSelector=2, linkLabel=Value at 2]{SELECT * FROM test WHERE a=2;}, \dataText[linkTo=testname, linkSelector=3, linkLabel=Value at 3]{SELECT * FROM test WHERE a=3;}.\\[3pt]
@@
 \chart[width=0.45\textwidth, height=0.3\textwidth, chartType=bar]{SELECT a, b, [attribute with space] FROM test;}\\[3pt]
+
 {\fontfamily{phv} \tiny \selectfont
@@
 }
+
 \end{document}

### 2016-07-19 minimal.tex
Version 1.2.6

@@
 %\dbCSVFile{result.csv}{test}
+
+\jsonDirectory{examples/json}
+\jsonFile{examples/json/nested/cake.json}{cake}
 
@@
 
+The following Chart uses information from JSON-files. \\[3pt]
+\chart[width=0.45\textwidth, height=0.3\textwidth, logscale=true, chartType=line]{SELECT a, b, c FROM json_files WHERE json_filename LIKE 'run\%';}
+\chart[width=0.45\textwidth, height=0.3\textwidth, logscale=true, chartType=line, name=testname]{SELECT a, b, c FROM json_files WHERE c > 5 AND json_filename LIKE 'run\%';} \\[3pt]
+
+To check if the results from JSON-files contain correct data, one can insert his own data for comparision. Examples to drag and drop can be found in the examples folder.
+\begin{itemize}
+\item This JSON has correct values \json{run1}
+\item Here some values are changed \json{run4}
+\item And here are many nested values \json{cake}
+\end{itemize}
+
 \end{document}

### 2016-07-19 pdbf-doc.tex
Version 1.2.6

@@
 \author{
- Patrick Bender\\
-  \texttt{s9pabend@stud.uni-saarland.de}
+ Patrick Bender \hspace{2cm} Lukas Lange \hspace{2cm} Shoaib Malik\\
+  \texttt{\{s9pabend, s9lslang, s9mumali\}@stud.uni-saarland.de}
 }
@@
 
+\subsection{Enable the comparison of your results to other users}
+To make your own results even more transparent, you can let other users input their own result files.
+If there is a difference, the PDBF will recognize it, and will automaticly create a highlighted view of changes.
+However right now, this option is only avaible for JSON-files as datasource.
+
 \subsection{Compile LaTeX to single HTML file}
@@
 google-diff-match-patch (\url{https://code.google.com/p/google-diff-match-patch/})\\
+jsondiffpatch (\url{https://github.com/benjamine/jsondiffpatch}) \\
 explorercanvas (\url{https://code.google.com/p/explorercanvas/})\\
@@
 \noindent Macros: \\[3pt]
-\verb|\dbSQLText{sqlQueryString}| \\
-\verb|\dbSQLFile{fileWithSQLQueries}| \\
+\verb|\jsonFile{path/to/file}{filename} | \\
+\verb|\jsonDirectory{path/to/directory}| \\
 \textbf{Note:} Relative filepaths are relative to the tex file.\\
@@
 \textbf{Note:} jdbcConnectionURL, user, and password are not stored in the output documents.\\[4pt]
+
+\subsection{JSON}
+\noindent Macros: \\[3pt]
+\verb|\dbSQLText{sqlQueryString}| \\
+\verb|\dbSQLFile{fileWithSQLQueries}| \\
+\textbf{Note:} Relative filepaths are relative to the tex file.\\
+\textbf{Note:} The \verb|\dbSQLFile{fileWithSQLQueries}|-command will search the whole directory (it will not scan deeper file structures), and chooses the filename according to existing filename. \\
+\textbf{Note:} You can insert as many JSON-commands as you like. \\[3pt]
 
@@
 %
+
+\noindent\textbf{Note:} To choose JSON-files as your datasource for charts, add to your SQL-FROM clause \texttt{json-files} and to your SQL-WHERE clause \texttt{filename LIKE \%What You Want\%) \\[3pt]
 

### 2016-07-19 pdbf-doc.tex
corrected wrong ) -> }

@@
 \author{
- Patrick Bender \hspace{2cm} Lukas Lange \hspace{2cm} Shoaib Malik\\
+ Patrick Bender \hspace{1cm} Lukas Lange \hspace{1cm} Shoaib Malik\\
   \texttt{\{s9pabend, s9lslang, s9mumali\}@stud.uni-saarland.de}
@@
 
-\noindent\textbf{Note:} To choose JSON-files as your datasource for charts, add to your SQL-FROM clause \texttt{json-files} and to your SQL-WHERE clause \texttt{filename LIKE \%What You Want\%) \\[3pt]
+\noindent\textbf{Note:} To choose JSON-files as your datasource for charts, add to your SQL-FROM clause \texttt{json-files} and to your SQL-WHERE clause \texttt{filename LIKE \%What You Want\%} \\[3pt]
 

### 2016-07-20 pdbf-doc.tex

@@
 \dbSQLText{CREATE TABLE test(a int, b int); INSERT INTO test VALUES(1, 10); INSERT INTO test VALUES(2, 100); INSERT INTO test VALUES(3, 120); INSERT INTO test VALUES(4, 100);}
+
+\jsonFile{examples/json/nested/cake.json}{cake}
 
@@
 \noindent Macros: \\[3pt]
-\verb|\jsonFile{path/to/file}{filename} | \\
-\verb|\jsonDirectory{path/to/directory}| \\
+\verb|\dbSQLText{sqlQueryString}| \\
+\verb|\dbSQLFile{fileWithSQLQueries}| \\
 \textbf{Note:} Relative filepaths are relative to the tex file.\\
@@
 \noindent Macros: \\[3pt]
-\verb|\dbSQLText{sqlQueryString}| \\
-\verb|\dbSQLFile{fileWithSQLQueries}| \\
+\verb|\jsonFile{path/to/file}{filename} | \\
+\verb|\jsonDirectory{path/to/directory}| \\
 \textbf{Note:} Relative filepaths are relative to the tex file.\\
-\textbf{Note:} The \verb|\dbSQLFile{fileWithSQLQueries}|-command will search the whole directory (it will not scan deeper file structures), and chooses the filename according to existing filename. \\
+\textbf{Note:} The \verb|\jsonDirectory|-command will search the whole directory (it will not scan deeper file structures), and chooses the filename according to existing filename. \\
 \textbf{Note:} You can insert as many JSON-commands as you like. \\[3pt]
@@
 
+\subsection{JSON}
+Clickable JSON-statement, that will open a seperated view for JSON files.
+Within this view, other user can input their own JSON files and a highlighted verison, showing differences between the files, will be automatically generated.
+\\[3pt]
+\textbf{Note:} Currently JSON is limited to one line of text. No line breaks are possible.\\
+Example:\\[3pt]
+You can compare the the cake.json in \textit{examples/comparison/} to this json: \json{cake}\\[3pt]
+%
+\noindent Macros: \\[3pt]
+\verb|\json{filename}| \\
+
+\noindent Options: \\[3pt]
+
+\option
+{color}
+{The textcolor of the displayed text.}
+{\texttt{white}, \texttt{black}, \texttt{red}, \texttt{green}, \texttt{blue}, \texttt{cyan}, \texttt{magenta}, \texttt{yellow} or xcolor syntax (see \href{http://mirror.unicorncloud.org/CTAN/macros/latex/contrib/xcolor/xcolor.pdf}{here}).
+Can also be redefined globally (\texttt{pdbfDynamicTextColor})}
+{\texttt{blue}}
+
 \subsection{DataTable}

### 2016-07-29 minimal.tex
versioning 1.2.6

@@
 \begin{document}
-\noindent
-
-\sql{SELECT * FROM test;}\\[3pt]
-The table ``test'' contains \dataText{SELECT COUNT(*) FROM test;} tuples.\\[3pt]
-
-The following points are linked to the chart on the left side below: \dataText[linkTo=testname, linkSelector=2, linkLabel=Value at 2]{SELECT * FROM test WHERE a=2;}, \dataText[linkTo=testname, linkSelector=3, linkLabel=Value at 3]{SELECT * FROM test WHERE a=3;}.\\[3pt]
-\chart[width=0.45\textwidth, height=0.3\textwidth, logscale=true, chartType=line, name=testname]{SELECT a, b, [attribute with space] FROM test;}
-\chart[width=0.45\textwidth, height=0.3\textwidth, chartType=bar]{SELECT a, b, [attribute with space] FROM test;}\\[3pt]
-
-{\fontfamily{phv} \tiny \selectfont
-	\noindent
-	Change the font before this line to Helvetica and fontsize to tiny.\\[3pt]
-	\chart[width=0.6\textwidth, height=0.25\textwidth, chartType=compareToBest]{SELECT b FROM test;}\\[3pt]
-	\multiplotChart[width=0.6\textwidth, height=0.5\textwidth, xCount=2, yCount=2, yValues={["b", "b*2"]}, xValues={["a <= 2", "a > 2"]}, topArr={["a smaller or equal 2", "a greater 2"]}, yFirst]{SELECT a,?
-FROM test WHERE ?;}\\[3pt]
-	\pivotTable[width=0.10\textwidth, height=0.15\textwidth, rows={["a"]}, aggregationattribute=b]{SELECT * FROM test;}
-	\dataTable[verticalLines=i, horizontalLinesHeader=b hh, horizontalLinesBody=i]{SELECT * FROM test;}\\[3pt]
-}
 
@@
 
+\attachment[click me][view][local]{examples/attachment/pdf\_reference.pdf} to view pdf\_reference.pdf embedded from local storage. \\
+\attachment[click me][download][local]{examples/attachment/pdf\_reference.pdf} to download pdf\_reference.pdf embedded from local storage. \\
+\attachment[click me][view][local]{examples/attachment/big\_mountain.png} to view big\_mountain.png embedded from local storage. \\
+%\attachment[click me][view][web]{https://www.google.de/} to view Google.html embedded from web storage. \\
+\attachment[click me][download][local]{examples/attachment/Google.html} to download Google.html embedded from local storage. \\
+%\attachment[click me][view][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to view pdf\_reference.pdf embedded from web storage. \\
+%\attachment[click me][download][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to download pdf\_reference.pdf embedded from web storage. \\
+
+\graph[width=0.75\textwidth, height=0.5\textwidth]{examples/graphs/graph.txt} \\
+\graph[width=0.75\textwidth, height=0.5\textwidth]{examples/graphs/graph2.txt} \\
+\graph[width=0.75\textwidth, height=0.75\textwidth, customImage=examples/graphs/customGraph.png]{} \\
+
 \end{document}

### 2016-07-29 src/pdbf/tests/attachment.tex
versioning 1.2.6

\documentclass[10pt]{article}
\usepackage{pdbf}
\usepackage[a4paper]{geometry}

\begin{document}

\attachment[click me][view][local]{attachment/pdf\_reference.pdf} to view pdf\_reference.pdf embedded from local storage. \\
\attachment[click me][download][local]{attachment/pdf\_reference.pdf} to download pdf\_reference.pdf embedded from local storage. \\
\attachment[click me][view][local]{attachment/big\_mountain.png} to view big\_mountain.png embedded from local storage. \\
\attachment[click me][view][web]{https://www.google.de/} to view Google.html embedded from web storage. \\
\attachment[click me][download][local]{attachment/Google.html} to download Google.html embedded from local storage. \\
\attachment[click me][view][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to view pdf\_reference.pdf embedded from web storage. \\
\attachment[click me][download][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to download pdf\_reference.pdf embedded from web storage. \\

\end{document}

### 2016-07-29 data/dracula_graph.js

@@ -1,0 +1,529 @@
+/*
+ *  Dracula Graph Layout and Drawing Framework 0.0.3alpha
+ *  (c) 2010 Philipp Strathausen <strathausen@gmail.com>, http://strathausen.eu
+ *  Contributions by Jake Stothard <stothardj@gmail.com>.
+ *
+ *  based on the Graph JavaScript framework, version 0.0.1
+ *  (c) 2006 Aslak Hellesoy <aslak.hellesoy@gmail.com>
+ *  (c) 2006 Dave Hoover <dave.hoover@gmail.com>
+ *
+ *  Ported from Graph::Layouter::Spring in
+ *    http://search.cpan.org/~pasky/Graph-Layderer-0.02/
+ *  The algorithm is based on a spring-style layouter of a Java-based social
+ *  network tracker PieSpy written by Paul Mutton <paul@jibble.org>.
+ *
+ *  This code is freely distributable under the MIT license. Commercial use is
+ *  hereby granted without any cost or restriction.
+ *
+ *  Links:
+ *
+ *  Graph Dracula JavaScript Framework:
+ *      http://graphdracula.net
+ *
+ /*--------------------------------------------------------------------------*/
+
+/*
+ * Edge Factory
+ */
+var AbstractEdge = function() {
+}
+AbstractEdge.prototype = {
+    hide: function() {
+        this.connection.fg.hide();
+        this.connection.bg && this.bg.connection.hide();
+    }
+};
+var EdgeFactory = function() {
+    this.template = new AbstractEdge();
+    this.template.style = new Object();
+    this.template.style.directed = false;
+    this.template.weight = 1;
+};
+EdgeFactory.prototype = {
+    build: function(source, target) {
+        var e = jQuery.extend(true, {}, this.template);
+        e.source = source;
+        e.target = target;
+        return e;
+    }
+};
+
+/*
+ * Graph
+ */
+var Graph = function() {
+    this.nodes = {};
+    this.edges = [];
+    this.snapshots = []; // previous graph states TODO to be implemented
+    this.edgeFactory = new EdgeFactory();
+};
+Graph.prototype = {
+    /* 
+     * add a node
+     * @id          the node's ID (string or number)
+     * @content     (optional, dictionary) can contain any information that is
+     *              being interpreted by the layout algorithm or the graph
+     *              representation
+     */
+    addNode: function(id, content) {
+        /* testing if node is already existing in the graph */
+        if(this.nodes[id] == undefined) {
[…]

### 2016-07-29 data/main.js

@@ -72,6 +72,133 @@ function getCheckbox(labelname, containerControl) {
 	return checkbox;
 }
 
+var graphs = {}
+
+function createGraph(json, edgeList, canvasId, phantomJS, display) {
+	var edges = edgeList.split(",");
+	var degreeList = {};
+	var g = new Graph();
+	for(var i = 0; i < edges.length; i++) {
+		var edge = edges[i];
+		var node1 = edge.split("_")[0];
+		var node2 = edge.split("_")[1];
+		if (node1 in degreeList) {
+			degreeList[node1].push(node2);
+		}
+		else {
+			degreeList[node1] = [node2];
+		}
+		if (node2 in degreeList) {
+			degreeList[node2].push(node1);
+		}
+		else {
+			degreeList[node2] = [node1];
+		}
+		if(display) {
+		    g.addEdge(node1, node2);
+		}
+	}
+	if(display) {
+        var layouter = new Graph.Layout.Spring(g);
+        layouter.layout();
+        var renderer;
+        if(canvasId == "canvas"+json.type.I.graphID) {
+                renderer = new Graph.Renderer.Raphael(canvasId, g, (json.type.I.x2 - json.type.I.x1 + 0.001) * 1000,
+                        (json.type.I.y1 - json.type.I.y2 + 0.001) * 1000 * 1.40, canvasId, false);
+                graphs[canvasId] = renderer;
+        }
+        else {
+            var width;
+            var height;
+            if(canvasId == "canvasRight"+json.type.I.graphID || canvasId == "canvasLeft"+json.type.I.graphID+"Comapre" || canvasId == "canvasLeft"+json.type.I.graphID+"SavedComparison"){
+                width = document.getElementById(document.getElementById(canvasId).parentElement.getAttribute("id")).offsetWidth;
+                height = document.getElementById(document.getElementById(canvasId).parentElement.getAttribute("id")).offsetHeight;
+                width = width - width/10;
+                height = height - height/10;
+            }
+            else{
+                width = document.getElementById(json.name+"SliderWrapper").offsetWidth;
+                height = document.getElementById(json.name+"SliderWrapper").offsetHeight;
+            }
+
+            renderer = new Graph.Renderer.Raphael(canvasId, g, width, height, canvasId, true);
+        }
+        renderer.draw();
+
+        if(phantomJS) {
+            var svg = document.getElementById(canvasId).firstChild;
+            svg.setAttribute("width", "100%");
[…]

### 2016-07-29 src/pdbf/json/Graph.java

@@ -1,0 +1,5 @@
+package pdbf.json;
+
+public class Graph extends VisualPDBFelement{
+
+}

### 2016-07-29 minimal.tex
versioning 1.2.6

--- 2016-07-19 minimal.tex
@@
 
+\attachment[click me][view][local]{examples/attachment/pdf\_reference.pdf} to view pdf\_reference.pdf embedded from local storage. \\
+\attachment[click me][download][local]{examples/attachment/pdf\_reference.pdf} to download pdf\_reference.pdf embedded from local storage. \\
+\attachment[click me][view][local]{examples/attachment/big\_mountain.png} to view big\_mountain.png embedded from local storage. \\
+\attachment[click me][view][web]{https://www.google.de/} to view Google.html embedded from web storage. \\
+\attachment[click me][download][local]{examples/attachment/Google.html} to download Google.html embedded from local storage. \\
+\attachment[click me][view][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to view pdf\_reference.pdf embedded from web storage. \\
+\attachment[click me][download][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to download pdf\_reference.pdf embedded from web storage. \\
+
+\graph[width=0.75\textwidth, height=0.5\textwidth]{examples/graphs/graph.txt} \\
+
 \end{document}

### 2016-07-29 data/main.js

@@ -104,7 +104,7 @@ function createGraph(json, edgeList, canvasId, phantomJS, display) {
         var renderer;
         if(canvasId == "canvas"+json.type.I.graphID) {
                 renderer = new Graph.Renderer.Raphael(canvasId, g, (json.type.I.x2 - json.type.I.x1 + 0.001) * 1000,
-                        (json.type.I.y1 - json.type.I.y2 + 0.001) * 1000 * 1.40, canvasId, false);
+                        (json.type.I.y1 - json.type.I.y2 + 0.001) * 1000 * 1.40, canvasId, true);
                 graphs[canvasId] = renderer;
         }
         else {
@@ -4106,6 +4106,7 @@ function getPivotTableData(json, isBig) {
 	var aggrName = (isBig ? json.type.I.aggregationBig : json.type.I.aggregation);
 	try {
 		var aggr = $.pivotUtilities.aggregators[aggrName]([ aggrAttribute ]);
+		console.log(aggr);
 	} catch (e) {
 		return {
 			error : 'Aggregation function ' + aggrName + ' does not exist!'
@@ -4471,7 +4472,7 @@ function prettifySQL(sql) {
 	// pretty printing of alasql breaks quoted literals
 }
 
-
+/*
 document.addEventListener("mousewheel", mouseWheelEvent);
 
 function mouseWheelEvent(event) {
@@ -4486,7 +4487,7 @@ function mouseWheelEvent(event) {
 
 	}
 }
-
+*/
 
 function zoomGraphs() {
 	for(var i = 0; i < Object.keys(edgeLists).length; i++) {

### 2016-07-29 src/pdbf/compilers/Pre_Compiler.java

@@ -313,7 +313,7 @@ public class Pre_Compiler {
         // Produce necessary data and auxiliary files
         System.out.println("Generating images, latex tables, and text...\n" + count + " PDBF elements to process");
         for (int i = 0; i < pdbfElementContainers.length; ++i) {
-            if (pdbfElementContainers[i].type instanceof VisualPDBFelement && pdbfElementContainers[i].type.customImage == null) {
+        	if (pdbfElementContainers[i].type instanceof VisualPDBFelement && pdbfElementContainers[i].type.customImage == null) {
                 processVisual(pdbfElementContainers[i]);
             } else if (pdbfElementContainers[i].type instanceof Text) {
                 System.out.println("Finished " + pdbfElementContainers[i].name);
@@ -1070,16 +1070,15 @@ public class Pre_Compiler {
         return true;
     }
 
-
     private static void processVisual(PDBFelementContainer o) {
-		if (o.type instanceof Graph) {
+   		if (o.type instanceof Graph) {
 			processGraph(o);
 		}
         if (PDBF_Compiler.includeRes) {
             VisualPDBFelement c = (VisualPDBFelement) o.type;
             cleanupfiles.add(baseDirData + o.name + ".html");
             cleanupfiles.add(baseDir + o.name + ".json");
-			if (!(o.type instanceof Graph)) {
+			if (!(o.type instanceof Graph) ) {
 				preloadfiles.add(baseDir + o.name + ".json");
 			}
 			copyfiles.add(baseDir + o.name + ".pdf");

### 2016-07-29 pdbf-doc.tex
versioning 1.2.6

@@
 \usepackage{listings}
+\usepackage{tabto}
 \usepackage[a4paper,vmargin={20mm,40mm},hmargin={20mm,20mm}]{geometry}
@@
 \section{Features}
-\subsection{Automatic generation of Charts, Multiplot Charts, Pivot tables}
+\subsection{Automatic generation of Charts, Multiplot Charts, Pivot tables, Graphs}
 With PDBF you don't need to manually generate these kinds of elements.
@@
 \subsection{Generate your document directly from the results of your experiment}
-The idea of PDBF is to store the results of the experiment directly in the document and to make it more transparent how this chart, pivot table, etc.
+The idea of PDBF is to store the results of the experiment directly in the document and to make it more transparent how this chart, pivot table, graphs etc.
 was generated from the result data.
-Therefore we currently support CSV files, SQL files, and SQL servers as data sources and use SQL as description language for the transformation of the raw result data to the final representation in the document.
+Excluding graphs, we currently support CSV files, SQL files, and SQL servers as data sources and use SQL as description language for the transformation of the raw result data to the final representation in the document.
+For graphs, we use .txt files are the data source.
 
@@
 If there is a difference, the PDBF will recognize it, and will automaticly create a highlighted view of changes.
-However right now, this option is only avaible for JSON-files as datasource.
+However right now, this option is only avaible for JSON-files as datasource for charts and TXT-files as datasource for graphs.
 
@@
 YUI Compressor (\url{https://github.com/yui/yuicompressor})\\
-
+Dracula.js (\url{https://www.graphdracula.net/})\\
 \newpage
@@
 
+\subsection{Undirected Graph}
+
+\begin{figure}[h!]%
+\hspace{-218pt}Example:\\
+    \centering
+    \begin{minipage}{.75\textwidth}
+    \graph[width=0.75\textwidth, height=0.5\textwidth]{examples/graphs/graph.txt} \\
+    \caption{An undirected graph}
+    \end{minipage}
+\end{figure}
+
+\noindent Macros: \\[3pt]
+\verb|\graph[options]{filePath}| \\
+\textbf{Note:} \verb|filePath| is the path to the .txt graph file with tab seperated nodeID pair on each line representing an edge as follows;\\
+nodeA \space\space\space\space\space nodeB \\
+nodeB \space\space\space\space\space nodeC \\
+nodeA \space\space\space\space\space nodeC \\ \\
+\noindent Options:\\[3pt]
+%
+
+\option
+{width}
+{Sets the width of the graph.}
+{\LaTeX~length}
+{No default value. Value must always be set!}
+
+\option
+{height}
+{Sets the height of the graph.}
+{\LaTeX~length}
+{No default value. Value must always be set!}
+
+\subsection{Attachment}
+\hspace{-218pt}Example:\\
+\attachment[click me][view][local]{examples/attachment/big\_mountain.png} to view big\_mountain.png embedded from local storage. \\
+\attachment[click me][view][web]{http://www.khazanah.com.my/khazanah/files/20/200f21f3-07ff-4903-ab99-7c0cb557eb51.pdf} to view pdf\_reference.pdf embedded from web storage. \\
+
+\noindent Macros: \\[3pt]
+\verb|\attachment[fileText][target][storage]{filePath}| \\
+\textbf{Note:} \verb|filePath| is the path to the file which you want to attach. \\
+
+\noindent Options:\\[3pt]
+%
+\newpage
+\option
+{fileText}
+{Sets the text that is appeared on the link in the PDBF document.}
+{Unquoted string}
+{No default value. Value must always be set!}
+
+\option
+{target}
+{Sets whether the attachment be viewed in a new tab or downloaded on local machine on link click in HTML.}
+{\textit{view} / \textit{download}. \textit{view} is only supported for a certain file formats such as png, jpg, jpeg, txt, pdf, html. \textit{download} is supported for all file formats.}
+{No default value. Value must always be set!}
+
+\option
+{target}
+{Sets whether the file to be embedded is present on local machine or on a web server.}
+{\textit{local} / \textit{web}.
+If storage is \textit{web}, then filePath is a URL to the file that you want to embed, else if storage is \textit{local}, filePath is simply a path to the file where it is located.}
+{No default value. Value must always be set!}
+
 \section{F.A.Q.}

### 2016-07-30 minimal.tex
versioning 1.2.6
version 1.2.7

@@
 
-\attachment[click me][view][local]{examples/attachment/pdf\_reference.pdf} to view pdf\_reference.pdf embedded from local storage. \\
-\attachment[click me][download][local]{examples/attachment/pdf\_reference.pdf} to download pdf\_reference.pdf embedded from local storage. \\
 \attachment[click me][view][local]{examples/attachment/big\_mountain.png} to view big\_mountain.png embedded from local storage. \\
-\attachment[click me][view][web]{https://www.google.de/} to view Google.html embedded from web storage. \\
+\attachment[click me][view][web]{https://www.google.de} to view Google.html embedded from web storage. \\
 \attachment[click me][download][local]{examples/attachment/Google.html} to download Google.html embedded from local storage. \\

### 2016-07-31 pdbf-doc.tex
version 1.2.7
versioning 1.2.7

@@
 \item Adjust config.cfg
-\item Try to compile minimal.tex file with this command: \texttt{java -jar pdbf.jar minimal.tex}
+\item Try to compile minimal.tex (with working internet connection) file with this command: \texttt{java -jar pdbf.jar minimal.tex}
 \item Open minimal.html, this is the final output of the compilation process, if you rename it to ".pdf" it is also a valid pdf-document
@@
 Excluding graphs, we currently support CSV files, SQL files, and SQL servers as data sources and use SQL as description language for the transformation of the raw result data to the final representation in the document.
-For graphs, we use .txt files are the data source.
+For graphs, we use .txt files as the data source.
 
@@
 \item Run "mvn package" if you only want to compile pdbf.jar
-\item Run "mvn verify" if you want to compile pdbf.jar and run integration tests.
+\item Run "mvn verify" (with working internet connection) if you want to compile pdbf.jar and run integration tests.
 \end{itemize}

### 2016-07-31 src/pdbf/PDBF_Compiler.java

@@ -62,7 +62,7 @@ public class PDBF_Compiler {
 	    URL url = new URL("https://raw.githubusercontent.com/uds-datalab/PDBF/gh-pages/VERSION.md");
 	    String currentVersion = IOUtils.toString(url.openStream(), StandardCharsets.UTF_8);
 	    if (!version.equals(currentVersion)) {
-		System.err.println(currentVersion + " of PDBF compiler is available.\n"
+		System.err.println(currentVersion.replace(System.lineSeparator(), "") + " of PDBF compiler is available.\n"
 			+ "Please visit https://github.com/uds-datalab/PDBF and download the latest version\n");
 	    }
 	} catch (Throwable t) {
